---
output:
  pdf_document: default
  html_document: default
---
# ========================================================================================================= #
## IMPORTANT - To get started:
Because this script is an R Markdown script, the default directory is the location of this .Rmd file.
You must open this script from within the .Rproj file associated with your wiosym database for it to work.
If you haven't used R notebooks before (.Rmd), each of the code "chunks" can be executed by clicking the green play
button at the top right of the chunk.

To run the script from the project directory, in R Studio, you need to go to: Tools > Global Options > R Markdown
and set "Evaluate chunks in directory" to "Project", though for latest version of R/Rstudio it may not be needed...
# ========================================================================================================= #
# ABOUT
# ========================================================================================================= #
## Brief description: Coral components, shallow and mesophotic
## Sripts/processes that needs to be updated prior?: Mangrove, depth
## Suggestions on future improvements: 
## Created: gk, sgu
## Updated: gk, sgu, 211021, adding new data from coral atlas (somalia region), and fixing problems with old version. Also converting to rmd
## Developer check: Initial, org, yymmdd, comments
## External check: Initial, org, yymmdd, comments
# ========================================================================================================= #
### INSTRUCTIONS
# ========================================================================================================= #    
# Style
  # Write code that is easy to follow and repeat.
  # Suggested R style is tidyverse. Style guide here https://style.tidyverse.org/index.html

# Metadata
  # All indata from "data_raw" must have a unique ID with *_metasym.txt file to go along, check before you add data.
  # All indata from and outputs to "data" must have a *_sourcesym.txt file with all (accumulated) sourceIDs used to create the layer.
  # The metasym files and tracking of those in the sourcesym files ensures correct sources for products and ISO metadata. This template helps you get it right..
  
# Naming and folders
  # Predefined standard names for main/sub folders and final products in two text files that can be added to in need.
  # ..\shiny_data_upload\modify_txt_files_v01.2.xlsx and ..\process\templates\modify_txt_files_data_v01.0.xlsx
  # Main version (e.g. v01 for 2021) follows dev cycles for all WIOSYm. For version control within v01 use V01.1, V01.2 etc..
  # More code/instructions to standardize folders and names under "set destinations" section
  # Each product dir have a "proc" folder. It's the temp space where all workfiles are stored. No final products here.  
  # All final products are stored under the root folder (e.g. ..\data\reg\eco\bh_coral\v01) together with 
  # a nameofmyfile_sourcesym.txt file with all source IDs used. No other files here. move products to _archive\ dir if defunct 
# ========================================================================================================= # 
# PREPARATIONS
# ========================================================================================================= #

```{r}
x <- c("tidyverse", "sf", "raster", "rgdal", "fasterize", "labelled", "gdalUtils", "foreign")
#install.packages(x, dependencies=TRUE)
                                            
library(tidyverse)
library(sf)
library(raster)
library(rgdal)
library(labelled)
library(gdalUtils)
library(foreign) # write dbf 
```

## Set version
```{r}
version = "v01" # relates to major iteration of WIOSym (v01 for grid 2021, v02 for grid 2022)
d_version = ".1"  # detailed version (you can add another level e.g. 0.0  -> v01.0.0)
seq = "s01" # sequential order - change accordingly if process is run in several scripts
```

## Set geographic scope
```{r}
read_tsv("./shiny_data_upload/locations.txt")  # prints available locations to choose from and their abbreviations
```

```{r}
location = "reg" # choose applicable 3 letter abbreviation from "location_val" -  reg for whole WIO area
```

## Set main theme
```{r}
folders_raw <- read_tsv("./shiny_data_upload/folders.txt")
folders_data <- read_tsv("./shiny_data_upload/folders_data.txt") # additional folders for data not applicable for data_raw 
folders <- folders_raw %>% add_row(folders_data)
(unique(folders["theme_folder"]))
```

```{r}
theme <- "eco" # Copy from "theme_folder" (e.g. "eco")
```

## Set Subtheme
```{r}
folders %>%
  filter(theme_folder == theme) %>% 
  dplyr::select(subtheme_folder, subtheme_name) %>% 
  print(n = Inf)
```

```{r}
subtheme <- "bh_coral" # copy from "subtheme_folder"
```

## Set paths
```{r}
(dest_path <- paste("./data", location, theme, subtheme, version, "", sep="/")) # path to final product
(work_path <- paste(dest_path, "proc/", sep=""))
(proc_path <- paste(dest_path, "proc_log/", sep=""))
(archive_path <- paste(dest_path, "_archive/", sep=""))
(script_path <- "./process/r/")
```

## Create directories
```{r}
dir.create(dest_path, recursive = TRUE)
dir.create(work_path, recursive = TRUE)
dir.create(proc_path, recursive = TRUE)
dir.create(archive_path, recursive = TRUE)
dir.create(script_path, recursive = TRUE)
```

## Save R Script
```{r}
name1 <- "" # "REPLACE_" if script need additional name... don't use unless needed to avoid duplication.
(script_path_windows <- paste(getwd(), "/process/r", sep=""))
(script_name <- paste(theme, "_", subtheme,"_", location, "_", name1, seq, "_", version, ".Rmd", sep=""))
```
 # copy path and name and use File/Save As in R studio to save rmd file
 
#==========================================================================================================================#
# INDATA
#==========================================================================================================================#

# Set DATA_RAW sources
IMPORTANT all data harvested from data_raw needs to have a metasym.txt file in its root directory (not file specific)
##DIR1 Allen Coral
```{r}
# data1: coral data from allen atlas, full cover of wio area (except somalia)
source1_dir <- source_dir <-  "./data_raw/reg/eco/bhab/allen/gk2104281131/"
source1_id <- source_id <- "gk2104281131" # cut and past your source id here
source1_metasym <- read_tsv(paste(source_dir, source_id, "_metasym.txt", sep="")) # reads the associated metasym file
print(source1_metasym) # check that metasym file is populated, and fix if not
# check what data is in data_raw directory and read in all files to use from the directory:
```

```{r}
#File1: Allen coral data
dir(source_dir) # check what data is in data_raw directory
source_file <- "Benthic-Map/benthic.gpkg" # your data file(s)
source1_path1 <- source_path <- paste(source_dir, source_file, sep="")
source_path
coral_allen <- st_read(source_path)
#coral_sf_full
```
## DIR2 wcmc
```{r}
# data2: wcmc
source2_dir <- source_dir <-  "./data_raw/glo/eco/bh_coral/wcmc/gk2105272129/"
source2_id <- source_id <- "gk2105272129" # cut and past your source id here
source2_metasym <- read_tsv(paste(source_dir, source_id, "_metasym.txt", sep="")) # reads the associated metasym file
source2_metasym # check that metasym file is populated, and fix if not
# check what data is in data_raw directory and read in all files to use from the directory:
dir(source_dir, recursive=T) 
#File1: wcmc coral data
source_file <- "01_Data/WCMC008_CoralReef2018_Py_v4_1.shp"  # your data file(s)
source2_path1 <- source_path <- paste(source_dir, source_file, sep="")
source_path
coral_wcmc <- st_read(source_path)
coral_wcmc
```


## DIR3 Allen data 2
```{r}
# data3: updated coral data from allen atlas, somalia region
source3_dir <- source_dir <-  "./data_raw/reg/eco/bhab/allen/gk2110071206/"
source3_id <- source_id <- "gk2110071206" # cut and past your source id here
source3_metasym <- read_tsv(paste(source_dir, source_id, "_metasym.txt", sep="")) # reads the associated metasym file
print(source3_metasym) # check that metasym file is populated, and fix if not
# check what data is in data_raw directory and read in all files to use from the directory:
```

```{r}
#File1: Allen coral data
dir(source_dir) # check what data is in data_raw directory
source_file <- "Benthic-Map/benthic.geojson" # your data file(s)
source3_path1 <- source_path <- paste(source_dir, source_file, sep="")
source_path
coral_allen_2 <- st_read(source_path)
#coral_sf_full
```

## Number of "data_raw" directorys?
```{r}
data_raw_metasym_num <- 3 # set total number of data_raw dir used (e.g. <- 1), to help in export section
```

# DATA sources
IMPORTANT Products in data are based on information in data_raw. Check that *_sourcesym.txt exists for each file to track acccumulated IDs. 

## DATA1: Grid 1km - loading all flavours (watermask 1, 0, NA) at once since same sourcesym apply
```{r}
data_dir <-  "./data/reg/grid/grid/v01/"  # input your data directory
dir(data_dir) # check what data is in this directory

data1_file <- "grid_1km_v01.1.tif" 
data1_file2 <- "grid_1km_0_v01.1.tif" 
data1_file3 <- "grid_1km_na_v01.1.tif" 

(grid_1km <- raster(paste(data_dir, data1_file, sep=""))) # read and check your data
(grid_1km_0 <- raster(paste(data_dir, data1_file2, sep=""))) # read and check your data
(grid_1km_na <- raster(paste(data_dir, data1_file3, sep=""))) # read and check your data

data1_sourcesym <- read_tsv(paste(data_dir, data1_file, "_sourcesym.txt", sep=""))
# check that source IDs exists and make sense

plot(grid_1km)
```

## DATA2: Grid 250m - loading all flavours (watermask 1, 0, NA) at once since same sourcesym apply
```{r}
data2_dir <- data_dir <-  "./data/reg/grid/grid/v01/"  # input your data directory

data2_file <- "grid_250m_v01.1.tif" # your data file
data2_file2 <- "grid_250m_0_v01.1.tif" # your data file
data2_file3 <- "grid_250m_na_v01.1.tif" # your data file

(grid_250m <- raster(paste(data_dir, data2_file, sep=""))) # read and check your data
(grid_250m_0 <- raster(paste(data_dir, data2_file2, sep=""))) # read and check your data
(grid_250m_na <- raster(paste(data_dir, data2_file3, sep=""))) # read and check your data

(data2_sourcesym <- read_tsv(paste(data_dir, data2_file, "_sourcesym.txt", sep="")))
# check so source IDs exists and make sense
# check that source IDs exists and make sense
#plot(grid_250m)
```


## DATA 3: bounding box shapefile
```{r}
data3_dir <- data_dir <-  "./data/reg/grid/scope/v01/"  # input your data directory

data3_file <- "wiosym_grid_bounding_box_v01.shp" # your data file

data3_sourcesym <- read_tsv(paste(data_dir, data3_file, "_sourcesym.txt", sep=""))
#data3_sourcesym # check so source IDs exists and make sense - ok for these
data3_path <- data_path <- paste(data_dir, data3_file, sep="")

grid_poly <- st_read(data_path) # read in your data object if appropriate and perhaps change name to more informative...
#grid_poly
#plot(grid_poly)
```


```{r}
# data6: mangrove
#data6_dir <- data_dir <-  "./data/reg/eco/ch_mangrove/v01/"  # input your data directory
#dir(data_dir) # check what data is in this directory
#data_file <- "eco_ch_mangrove_1km_combined_presence_v01.0.tif" # your data file
#data6_sourcesym <- read_tsv(paste(data_dir, data_file, "_sourcesym.txt", sep=""))
#data6_sourcesym # check so source IDs exists and make sense
#data6_path <- data_path <- paste(data_dir, data_file, sep="")
#data_path
#Mangrove_1km <- raster(data_path) # read in your data object if appropriate and perhaps change name to more informative...
#plot(Mangrove_1km)
```

```{r}
# data7: depth
#data7_dir <- data_dir <-  "./data/reg/env/topo/v01/"  # input your data directory
#dir(data_dir) # check what data is in this directory
#data_file <- "bathymetry_mean_m_1km_v01.1.tif" # your data file
#data7_sourcesym <- read_tsv(paste(data_dir, data_file, "_sourcesym.txt", sep=""))
#data7_sourcesym # check so source IDs exists and make sense
#data7_path <- data_path <- paste(data_dir, data_file, sep="")
#data_path
#depth_1km <- raster(data_path) # read in your data object if appropriate and perhaps change name to more informative...
#plot(depth_1km)
```


# declare number of "data" files
```{r}
data_sourcesym_num <- 3 # REPLACE - set total number of data files (e.g. <- 2), to help in export section
```

#==========================================================================================================================#
### PROCESSING
#==========================================================================================================================#
# Allen coral FILE1 from shape to raster grid 
```{r}
coral_sf <- coral_allen
glimpse(coral_sf)
#unique(coral_sf$class)
#class(coral_sf)
```

## Allen 1. Coral only - Select coral polygons and convert to raster (next chunk)
```{r}
# combine into one file

# select coral class
selected_sf <- coral_sf %>%
  filter(class=="Coral/Algae") %>% 
  #  mutate(class_code=class) %>% 
  mutate(class_code = recode(class, "Coral/Algae" = "1")) %>% 
  mutate(class_code = as.numeric(class_code))
#  recode(class_code, "Coral/Algae" = 1)

glimpse(selected_sf)
```

```{r}
#write coral file to work directory for gdalutil
input_path <- paste(work_path, "temp.shp", sep="")
input_path

st_write(selected_sf,input_path, append=FALSE) 

allen1_coral_250m_path <- output_path <- paste(work_path, "grid_250m_na_allen1_coral.tif", sep= "")
output_path

writeRaster(grid_250m_na, output_path,  overwrite=T, COMPRESS=LZW)

# 250m grid mapping
coral_warp <- gdalUtils::gdal_rasterize(src_datasource = input_path,
                                        dst_filename = output_path,
                                        b = 1,
                                        at = T,
                                        a = "class_code",
                                        output_Raster = TRUE,
)

r <- raster(output_path)
allen1_coral_250m_path <- output_path <- paste(work_path, "grid_250m_na_allen1_coral_lzw.tif", sep= "")
writeRaster(r, output_path, overwrite=T, datatype = 'FLT4S', COMPRESS=LZW)
allen1_coral_250m_r <- raster(output_path)

```

## Allen 1. Seagrass  - not needed for the coral layer, only here to enable faster work with seagrass layer later on...
```{r}
# combine into one file

# select coral class
selected_sf <- coral_sf %>%
  filter(class=="Seagrass") %>% 
  #  mutate(class_code=class) %>% 
  mutate(class_code = recode(class, "Seagrass" = "1")) %>% 
  mutate(class_code = as.numeric(class_code))
#  recode(class_code, "Coral/Algae" = 1)

glimpse(selected_sf)
```

```{r}
#write coral file to work directory for gdalutil
input_path <- paste(work_path, "temp.shp", sep="")
input_path

st_write(selected_sf,input_path, append=FALSE) 

allen1_seagrass_250m_path <- output_path <- paste(work_path, "grid_250m_na_allen1_seagrass.tif", sep= "")
output_path

writeRaster(grid_250m_na, output_path,  overwrite=T, COMPRESS=LZW)

# 250m grid mapping
coral_warp <- gdalUtils::gdal_rasterize(src_datasource = input_path,
                                        dst_filename = output_path,
                                        b = 1,
                                        at = T,
                                        a = "class_code",
                                        output_Raster = TRUE,
)

rm(coral_warp)
r <- raster(output_path)
allen1_seagrass_250m_path <- output_path <- paste(work_path, "grid_250m_na_allen1_seagrass_lzw.tif", sep= "")
writeRaster(r, output_path, overwrite=T, datatype = 'FLT4S', COMPRESS=LZW)
allen1_seagrass_250m_r <- raster(output_path)
```

## Allen 1. Map footprint - shape to raster grid 
```{r}
coral_sf <- coral_allen
#glimpse(coral_sf)
#unique(coral_sf$class)
#class(coral_sf)

#Select only coral polygons

selected_sf <- coral_sf %>%
  filter(class!="Unknown") %>% 
  #  mutate(class_code=class) %>% 
  mutate(class_code = recode(class, "Coral/Algae" = "0", "Rubble" = "0", "Microalgal Mats" = "0", "Seagrass" = "0", "Rock"="0", "Sand"="0")) %>% 
  mutate(class_code = as.numeric(class_code))
#  recode(class_code, "Coral/Algae" = 1)

(glimpse(selected_sf))

# convertion from coral shape to raster 

#write habitat file to work directory for gdalutil
input_path <- paste(work_path, "temp.shp", sep= "")

st_write(selected_sf,input_path, append=FALSE) 

# write empty grid for gdalutil work
allen1_habitat_250m_path <- output_path <- paste(work_path, "grid_250m_na_allen1_habitat.tif", sep= "")

writeRaster(grid_250m_na, output_path,  overwrite=T, COMPRESS=LZW)

# 250m grid mapping
coral_warp <- gdalUtils::gdal_rasterize(src_datasource = input_path,
                                        dst_filename = output_path,
                                        b = 1,
                                        at = T,
                                        a = "class_code",
                                        output_Raster = TRUE,
)

r <- raster(output_path)
allen1_habitat_250m_path <- output_path <- paste(work_path, "grid_250m_na_allen1_habitat_lzw.tif", sep= "")
writeRaster(r, output_path, overwrite=T, datatype = 'FLT4S', COMPRESS=LZW)
allen1_habitat_250m_r <- raster(output_path)
```

# Allen 2. Coral only  - shape to raster grid 
```{r}
coral_sf <- coral_allen_2
glimpse(coral_sf)
#unique(coral_sf$class)
#class(coral_sf)
```

```{r}
# combine into one file

# select coral class
selected_sf <- coral_sf %>%
  filter(class=="Coral/Algae") %>% 
  #  mutate(class_code=class) %>% 
  mutate(class_code = recode(class, "Coral/Algae" = "1")) %>% 
  mutate(class_code = as.numeric(class_code))
#  recode(class_code, "Coral/Algae" = 1)

glimpse(selected_sf)
```

```{r}
# Shape to raster
# write coral file to work directory for gdalutil
input_path <- paste(work_path, "temp.shp", sep="")
input_path

st_write(selected_sf,input_path, append=FALSE) 

allen2_coral_250m_path <- output_path <- paste(work_path, "grid_250m_na_allen2_coral.tif", sep= "")
output_path

writeRaster(grid_250m_na, output_path,  overwrite=T, COMPRESS=LZW)

# 250m grid mapping
coral_warp <- gdalUtils::gdal_rasterize(src_datasource = input_path,
                                        dst_filename = output_path,
                                        b = 1,
                                        at = T,
                                        a = "class_code",
                                        output_Raster = TRUE,
)

r <- raster(output_path)
allen2_coral_250m_path <- output_path <- paste(work_path, "grid_250m_na_allen2_coral_lzw.tif", sep= "")
writeRaster(r, output_path, overwrite=T, datatype = 'FLT4S', COMPRESS=LZW)
allen2_coral_250m_r <- raster(output_path)

```

## Allen 2. Seagrass - shape to raster
```{r}
# combine into one file

# select coral class
selected_sf <- coral_sf %>%
  filter(class=="Seagrass") %>% 
  #  mutate(class_code=class) %>% 
  mutate(class_code = recode(class, "Seagrass" = "1")) %>% 
  mutate(class_code = as.numeric(class_code))
#  recode(class_code, "Coral/Algae" = 1)

glimpse(selected_sf)
```

```{r}
# shape to raster - 
# write file to work directory for gdalutil
input_path <- paste(work_path, "temp.shp", sep="")
input_path

st_write(selected_sf,input_path, append=FALSE) 

allen2_seagrass_250m_path <- output_path <- paste(work_path, "grid_250m_na_allen2_seagrass.tif", sep= "")
output_path

writeRaster(grid_250m_na, output_path,  overwrite=T, COMPRESS=LZW)

# 250m grid mapping
coral_warp <- gdalUtils::gdal_rasterize(src_datasource = input_path,
                                        dst_filename = output_path,
                                        b = 1,
                                        at = T,
                                        a = "class_code",
                                        output_Raster = TRUE,
)

rm(coral_warp)
r <- raster(output_path)
allen2_seagrass_250m_path <- output_path <- paste(work_path, "grid_250m_na_allen2_seagrass_lzw.tif", sep= "")
writeRaster(r, output_path, overwrite=T, datatype = 'FLT4S', COMPRESS=LZW)
allen2_seagrass_250m_r <- raster(output_path)
```

## Allen 2. Map footprint 
```{r}
coral_sf <- coral_allen_2
#glimpse(coral_sf)
#unique(coral_sf$class)
#class(coral_sf)

#Select only coral polygons

selected_sf <- coral_sf %>%
  filter(class!="Unknown") %>% 
  #  mutate(class_code=class) %>% 
  mutate(class_code = recode(class, "Coral/Algae" = "0", "Rubble" = "0", "Microalgal Mats" = "0", "Seagrass" = "0", "Rock"="0", "Sand"="0")) %>% 
  mutate(class_code = as.numeric(class_code))
#  recode(class_code, "Coral/Algae" = 1)

(glimpse(selected_sf))

# convertion from coral shape to raster 

#write habitat file to work directory for gdalutil
input_path <- paste(work_path, "temp.shp", sep= "")

st_write(selected_sf,input_path, append=FALSE) 

# write empty grid for gdalutil work
allen2_habitat_250m_path <- output_path <- paste(work_path, "grid_250m_na_allen2_habitat.tif", sep= "")

writeRaster(grid_250m_na, output_path,  overwrite=T, COMPRESS=LZW)

# 250m grid mapping
coral_warp <- gdalUtils::gdal_rasterize(src_datasource = input_path,
                                        dst_filename = output_path,
                                        b = 1,
                                        at = T,
                                        a = "class_code",
                                        output_Raster = TRUE,
)

r <- raster(output_path)
allen2_habitat_250m_path <- output_path <- paste(work_path, "grid_250m_na_allen2_habitat_lzw.tif", sep= "")
writeRaster(r, output_path, overwrite=T, datatype = 'FLT4S', COMPRESS=LZW)
allen2_habitat_250m_r <- raster(output_path)
```

# COMBINE Allen 1,2 Coral
```{r}
# Write empty raster
allen_coral_250m_path <- output_path <- paste(work_path, "grid_250m_na_allen_coral.tif", sep= "")

writeRaster(grid_250m_na, output_path,  overwrite=T, COMPRESS=LZW)

coral_warp <- gdalUtils::mosaic_rasters(
  gdalfile = c(allen1_coral_250m_path, allen2_coral_250m_path),
  dst_dataset = output_path,
  output.vrt = NULL,
  output_Raster = FALSE,
  separate = FALSE,
  trim_margins = NULL,
  gdalwarp_index = 1,
  gdalwarp_params = list(r = "near"),
  #force_ot = "Int16",
  verbose = FALSE,
  COMPRESS=LZW,
)
(r <- raster(output_path))
crs(r) <- crs(grid_250m)

allen_coral_250m_path <- output_path <- paste(work_path, "grid_250m_na_allen_coral_lzw.tif", sep= "")
writeRaster(r, output_path, overwrite=T, datatype = 'FLT4S', COMPRESS=LZW)
(allen_coral_250m_r <- raster(output_path))
```


## Combine Allen Seagrass
```{r}
# Write empty raster
allen_seagrass_250m_path <- output_path <- paste(work_path, "grid_250m_na_allen_seagrass.tif", sep= "")

writeRaster(grid_250m_na, output_path,  overwrite=T, COMPRESS=LZW)

coral_warp <- gdalUtils::mosaic_rasters(
  gdalfile = c(allen1_seagrass_250m_path, allen2_seagrass_250m_path),
  dst_dataset = output_path,
  output.vrt = NULL,
  output_Raster = FALSE,
  separate = FALSE,
  trim_margins = NULL,
  gdalwarp_index = 1,
  gdalwarp_params = list(r = "near"),
  #force_ot = "Int16",
  verbose = FALSE,
  COMPRESS=LZW,
)
(r <- raster(output_path))
crs(r) <- crs(grid_250m)

allen_seagrass_250m_path <- output_path <- paste(work_path, "grid_250m_na_allen_seagrass_lzw.tif", sep= "")
writeRaster(r, output_path, overwrite=T, datatype = 'FLT4S', COMPRESS=LZW)
(allen_seagrass_250m_r <- raster(output_path))
```

## Combine Allen 1,2, Map footprint
```{r}
# Write empty raster
allen_habitat_250m_path <- output_path <- paste(work_path, "grid_250m_na_allen_habitat.tif", sep= "")

writeRaster(grid_250m_na, output_path,  overwrite=T, COMPRESS=LZW)

coral_mosaic <- gdalUtils::mosaic_rasters(
  gdalfile = c(allen1_habitat_250m_path, allen2_habitat_250m_path),
  dst_dataset = output_path,
  output.vrt = NULL,
  output_Raster = FALSE,
  separate = FALSE,
  trim_margins = NULL,
  gdalwarp_index = 1,
  gdalwarp_params = list(r = "near"),
  #force_ot = "Int16",
  verbose = FALSE,
  COMPRESS=LZW,
)

r <- raster(output_path)
crs(r) <- crs(grid_250m)

allen_habitat_250m_path <- output_path <- paste(work_path, "grid_250m_allen_habitat_lzw.tif", sep= "")
writeRaster(r, output_path,  overwrite=T, datatype = 'FLT4S', COMPRESS=LZW)
allen_habitat_250m_r <- raster(output_path)

allen_habitat_250m_va11_r <- reclassify(allen_habitat_250m_r, c(-1,1,1))
allen_habitat_250m_val1_path <- output_path <- paste(work_path, "grid_250m_allen_habitat_val1_lzw.tif", sep= "")
writeRaster(allen_habitat_250m_va11_r , output_path,  overwrite=T, datatype = 'FLT4S', COMPRESS=LZW)



```


# WCMC coral - shape to raster 250m grid 
```{r}

coral_sf <- coral_wcmc
#glimpse(coral_sf)
#unique(coral_sf$METADATA_I)
#class(coral_sf)

coral_sf_sel1 <- coral_sf %>%
  #  filter(class=="Coral/Algae") %>% 
  #  mutate(class_code=class) %>% 
  #  mutate(class_code = recode(class, "Coral/Algae" = "1")) %>% 
  mutate(class_code = 1)
#  recode(class_code, "Coral/Algae" = 1)

#glimpse(coral_sf_sel1)


# convertion from coral shape to raster 

#write coral file to work directory for gdalutil
input_path <- paste(work_path, "temp.shp", sep= "")
st_write(coral_sf_sel1, input_path, append=FALSE)


# write empty grid for gdalutil work
wcmc_coral_250m_path <- output_path <- paste(work_path, "grid_250m_na_coral_wcmc.tif", sep= "")
writeRaster(grid_250m_na, output_path,  overwrite=T, COMPRESS=LZW)

# 250m grid mapping
coral_warp <- gdalUtils::gdal_rasterize(src_datasource = input_path,
                                        dst_filename = output_path,
                                        b = 1,
                                        at = T,
                                        a = "class_code",
                                        output_Raster = TRUE,
)

r <- raster(output_path)
wcmc_coral_250m_path <- output_path <- paste(work_path, "grid_250m_na_coral_wcmc_lzw.tif", sep= "")
writeRaster(r, output_path, overwrite=T, datatype = 'FLT4S', COMPRESS=LZW)
wcmc_coral_250m_r <- raster(output_path)
```



# COMBINE All 
## COMBINED CORAL MAP 250m 
reclassify wcmc cells which are not mapped as coral in Allen atlas, due to lower spatial precision in wcmc data
```{r}
# write empty grid for gdalutil output

writeRaster(grid_250m_na, paste(work_path, "grid_250m_na_coral_combined_2.tif", sep= ""), overwrite=T, COMPRESS=LZW)

coral_250m_combined <- paste(work_path, "grid_250m_na_coral_combined_2.tif", sep= "")

coral_250m_combined_r <- gdalUtils::mosaic_rasters(
  gdalfile = c(wcmc_coral_250m_path, allen_habitat_250m_path, allen_coral_250m_path),
  dst_dataset = coral_250m_combined,
  output.vrt = NULL,
  output_Raster = FALSE,
  separate = FALSE,
  trim_margins = NULL,
  gdalwarp_index = 1,
  gdalwarp_params = list(r = "near"),
  #force_ot = "Int16",
  verbose = FALSE,
  COMPRESS=LZW,
)

r <- raster(coral_250m_combined) # export as product
crs(r) <- crs(grid_250m)
path <- paste(work_path, "grid_250m_na_coral_combined_2_lzw.tif", sep= "")
writeRaster(r, path,  overwrite=T, COMPRESS=LZW)
coral_250m_combined_r <- raster(path)

#writeRaster(coral_250m_combined, paste(work_path, "grid_250m_na_coral_combined_2_lzw.tif", sep= ""), overwrite=T, COMPRESS=LZW, #datatype = 'INT4S')

```

## COMBINED Footprint 250m 
combine all footprints from allen and wcmc in order to balance proportions in 1km agg with number of cells with values, and for uncertainty later
```{r}
# write empty grid for gdalutil output

habitat_footprint_250m_path <- output_path <- paste(work_path, "grid_250m_na_habitat_footprint_combined.tif", sep= "")
writeRaster(grid_250m_na, output_path, overwrite=T, COMPRESS=LZW)


mosaic_r <- gdalUtils::mosaic_rasters(
  gdalfile = c(wcmc_coral_250m_path, allen_habitat_250m_val1_path),
  dst_dataset = output_path,
  output.vrt = NULL,
  output_Raster = FALSE,
  separate = FALSE,
  trim_margins = NULL,
  gdalwarp_index = 1,
  gdalwarp_params = list(r = "near"),
  #force_ot = "Int16",
  verbose = FALSE,
  COMPRESS=LZW,
)

r <- raster(output_path) # export as product
crs(r) <- crs(grid_250m)
habitat_footprint_250m_path <- path <- paste(work_path, "grid_250m_na_habitat_footprint_combined_lzw.tif", sep= "")
writeRaster(r, path,  overwrite=T, COMPRESS=LZW)
habitat_footprint_250m_r <- raster(path)

#writeRaster(coral_250m_combined, paste(work_path, "grid_250m_na_coral_combined_2_lzw.tif", sep= ""), overwrite=T, COMPRESS=LZW, #datatype = 'INT4S')

```


## AGGREGATE 1km all corals
```{r}
combined_coral_250m_agg1km <- aggregate(coral_250m_combined_r, fact=4, fun=sum) / 16 #4*4=16, i.e. divide by the maximum cell number to get proportions - export as product
 plot(combined_coral_250m_agg1k)

writeRaster(combined_coral_250m_agg1km, paste(work_path, "grid_1km_na_coral_combined_2_proportion250m.tif", sep= ""),  overwrite=T, COMPRESS=LZW)

```

## AGGREGATE 1km all footprint
```{r}
combined_footprint_250m_agg1km <- aggregate(habitat_footprint_250m_r, fact=4, fun=sum) / 16 #4*4=16, i.e. divide by the maximum cell number to get proportions - export as product
plot(combined_footprint_250m_agg1km)
writeRaster(combined_footprint_250m_agg1km, paste(work_path, "grid_1km_na_mapfootprint_proportion250m.tif", sep= ""),  overwrite=T, COMPRESS=LZW)

```

## 1km Relative foot print - Coral proportion 
```{r}
r1 <- combined_coral_250m_agg1km
r2 <- combined_footprint_250m_agg1km

summary(r1)
summary(r2)

r3 <- overlay(r1, r2, fun=function(r1, r2){return(r1/r2)})  # adjusting cells that have only a part mapped
r3
plot(r3)

#writeRaster(r3, paste(work_path, "grid_1km_na_coral_combined_2_proportion250m_footprintadjusted.tif", sep= ""),  overwrite=T, COMPRESS=LZW)

```


# Map final layers to 1km standard grid with 0 values
```{r}

combined_coral_250m_agg1km_zero <- merge(combined_coral_250m_agg1km, grid_1km_0)

combined_coral_250m_agg1km_relativefootprint_zero <- merge(r3, grid_1km_0)

  
```


#==========================================================================================================================#
### EXPORT
#==========================================================================================================================#
Write products to root directory when checked and ready to use elsewhere, accompanied by yourfilename.ext_sourcesym.txt 
for each file to track source IDs as they accumulate between processes / data


# SOURCESYM
Create sourcesym file with all source IDs used (data and data_raw) 
data raw sources (dir)
```{r}
print(paste("data_raw metasym files used = ", data_raw_metasym_num, sep=""))
(data_raw_sources <- tibble(id = c(source1_id, source2_id, source3_id))) # REPLACE / add all data_raw sources
```

## data sources (files)
```{r}
print(paste("data_sourcym files = ", data_sourcesym_num, sep=""))
data_sources <- data1_sourcesym %>% # add any "sourcesym" files applicable to this product (check indata section), keep adding sourcesym files as needed
  add_row(data2_sourcesym)%>%  
  #add_row(data3_sourcesym)%>% # bounding box do not have ID since we did it ourselves...
  #add_row(data4_sourcesym)%>% 
  #add_row(data5_sourcesym)%>% 
  #add_row(data6_sourcesym)%>% 
  unique() %>% 
  print()
```
# Sources combined
```{r}
(all_sources <- data_raw_sources %>% add_row(data_sources))
```
# if a product only use a subset of the total sources, copy the section above and create individual files below (e.g. product1_sources <- ...)


# PRODUCT 1: Coral 1km
### read objects
```{r}
product_orig <- combined_coral_250m_agg1km_zero # input and check your object to be written to file
product_orig2 <-combined_coral_250m_agg1km_relativefootprint_zero # input and check your object to be written to file
plot(product_orig)
```
### set names
```{r}
scale <- "1km"
unit <- "proportion"  # add unit for product with original values (e.g. m, km, perc, presence, proportion
unit2 <- "proportion_mapnorm"
component_names <- read_tsv("./shiny_data_upload/component_names.txt") 
component_names_selected <- component_names %>%
  dplyr::filter(theme_folder == theme) %>% 
  dplyr::select(group, file_name, description) %>% 
  print(n = Inf)
```

```{r}
descriptive_name <- "coral_reef"  # name of product, for pressure/ecosystem components select from name list above
```
### paths
```{r}
(product_orig_path <- paste(dest_path, descriptive_name, "_", unit, "_", scale, "_", version, d_version, ".tif", sep="" ))
(product_orig2_path <- paste(dest_path, descriptive_name, "_", unit2, "_", scale, "_", version, d_version, ".tif", sep="" ))
```
### Write to file
```{r}
# check datatype to be appropriate for values (flt or int, 'INT4S' option to..)
writeRaster(product_orig, product_orig_path, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
writeRaster(product_orig2, product_orig2_path, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
write_tsv(all_sources, paste(product_orig_path, "_sourcesym.txt", sep=""))  
write_tsv(all_sources, paste(product_orig2_path, "_sourcesym.txt", sep=""))  
```


# SAVE SCRIPT
## 1. save your current R script (File/Save)
## 2. Save a carbon copy of the R script to proc_log (identical script to what is used to create the products exported

```{r}
# 2. run code below and go to File/Save As and paste link and name:
dest_path_script <- paste(getwd(), proc_path)
print(dest_path_script <- gsub(" .", "", dest_path_script)) # path
print(script_name_copy <- paste(gsub(".Rmd", "", script_name), d_version, ".Rmd", sep="")) # script name
```

#==========================================================================================================================#
### FINAL CHECK
#==========================================================================================================================#
## Developer check
Do your own final check on the products and double check sourcesym file exists and are complete for all files in root 
 -> Sign at top of script in about section (# Approved by:)

## External check 
 To ensure repeatability and quality, make sure one colleage can run the script
 When script is proven repeatable/understandable and products/metadata look ok 
 -> Sign at top of script in about section (# Approved by:), and make any comments if needed
 There is also a work_log document in onedrive / gitHUB, check current arrangement with Swedish dev team

