---
output:
  pdf_document: default
  html_document: default
---
# ========================================================================================================= #
# IMPORTANT - To get started:
Because this script is an R Markdown script, the default directory is the location of this .Rmd file.
You must open this script from within the .Rproj file associated with your wiosym database for it to work.
If you haven't used R notebooks before (.Rmd), each of the code "chunks" can be executed by clicking the green play
button at the top right of the chunk.

To run the script from the project directory, in R Studio, you need to go to: Tools > Global Options > R Markdown
and set "Evaluate chunks in directory" to "Project", though for latest version of R/Rstudio it may not be needed...
# ========================================================================================================= #
# ABOUT
# ========================================================================================================= #
## Brief description:  Seagrass map first version. combininging multiple sources at 250m resolution similar to mangrove and coral process
###Needs updating with temperate/tropical to be completed

### Sripts/processes that needs to be updated prior?: depth, corals, (substrate?)
###
### Script by: gk, sgu, R 4.1.1
### Updated: Initial, org, Rver, comments
### Developer check: Initial, org, yymmdd, comments
### External check: Initial, org, yymmdd, comments
# ========================================================================================================= #
# INSTRUCTIONS
# ========================================================================================================= #    
## Style
  # Write code that is easy to follow and repeat.
  # Suggested R style is tidyverse. Style guide here https://style.tidyverse.org/index.html

## Metadata
  # All indata from "data_raw" must have a unique ID with *_metasym.txt file to go along, check before you add data.
  # All indata from and outputs to "data" must have a *_sourcesym.txt file with all (accumulated) sourceIDs used to create the layer.
  # The metasym files and tracking of those in the sourcesym files ensures correct sources for products and ISO metadata. This template helps you get it right..
  
## Naming and folders
  # Predefined standard names for main/sub folders and final products in two text files that can be added to in need.
  # ../shiny_data_upload/modify_txt_files_v01.2.xlsx and ..\process\templates\modify_txt_files_data_v01.0.xlsx
  # Main version (e.g. v01 for 2021) follows dev cycles for all WIOSYm. For version control within v01 use V01.1, V01.2 etc..
  # More code/instructions to standardize folders and names under "set destinations" section
  # Each product dir have a "proc" folder. It's the temp space where all workfiles are stored. No final products here.  
  # All final products are stored under the root folder (e.g. ..\data\reg\eco\bh_coral\v01) together with 
  # a nameofmyfile_sourcesym.txt file with all source IDs used. No other files here. move products to _archive\ dir if defunct 
# ========================================================================================================= # 
# PREPARATIONS
# ========================================================================================================= #

### Install packages
```{r, include = FALSE}
# list all packages that this script is dependent on, remove any redundant ones
x <- c("leaflet", "sf", "raster", "rgdal", "tidyverse")
#install.packages(x, dependencies=TRUE)
```

### Load packages
```{r, include = FALSE}
library(leaflet)
library(sf) 
library(raster)
library(rgdal)
library(tidyverse)
```

## Set version
```{r}
version = "v01" # relates to major iteration of WIOSym (v01 for grid 2021, v02 for grid 2022)
d_version = ".0"  # detailed version (you can add another level e.g. 0.0  -> v01.0.0)
seq = "s01" # sequential order - change accordingly if process is run in several scripts
```

## Set geographic scope
```{r}
read_tsv("./shiny_data_upload/locations.txt")  # prints available locations to choose from and their abbreviations
```

```{r}
location = "reg" # choose applicable 3 letter abbreviation from "location_val" -  reg for whole WIO area
```

## Set main theme
```{r}
folders_raw <- read_tsv("./shiny_data_upload/folders.txt")
folders_data <- read_tsv("./shiny_data_upload/folders_data.txt") # additional folders for data not applicable for data_raw 
folders <- folders_raw %>% add_row(folders_data)
(unique(folders["theme_folder"]))
```

```{r}
theme <- "eco" # Copy from "theme_folder" (e.g. "eco")
```

## Set Subtheme
```{r}
folders %>%
  filter(theme_folder == theme) %>% 
  dplyr::select(subtheme_folder, subtheme_name) %>% 
  print(n = Inf)
```

```{r}
subtheme <- "bh_veg_seagr" # copy from "subtheme_folder"
```

## Set paths
```{r}
(dest_path <- paste("./data", location, theme, subtheme, version, "", sep="/")) # path to final product
(work_path <- paste(dest_path, "proc/", sep=""))
(proc_path <- paste(dest_path, "proc_log/", sep=""))
(archive_path <- paste(dest_path, "_archive/", sep=""))
(script_path <- "./process/r/")
```

## Create directories
```{r}
dir.create(dest_path, recursive = TRUE)
dir.create(work_path, recursive = TRUE)
dir.create(proc_path, recursive = TRUE)
dir.create(archive_path, recursive = TRUE)
dir.create(script_path, recursive = TRUE)
```

## Save R Script
```{r}
name1 <- "" # "REPLACE_" if script need additional name... don't use unless needed to avoid duplication.
(script_path_windows <- paste(getwd(), "/process/r", sep=""))
(script_name <- paste(theme, "_", subtheme,"_", location, "_", name1, seq, "_", version, ".Rmd", sep=""))
```
 # copy path and name and use File/Save As in R studio to save rmd file
 
#==========================================================================================================================#
# INDATA
#==========================================================================================================================#
# DATA_RAW sources
IMPORTANT all data harvested from data_raw needs to have a metasym.txt file in its root directory (not file specific)

## DIR1: WCMC 
```{r}
source1_dir <- source_dir <-  "./data_raw/glo/eco/bh_veg_seagr/wcmc/gk2110231307/"  # REPLACE - internal path
source1 <- source_id <- "gk2110231307"
(read_tsv(paste(source_dir, source_id, "_metasym.txt", sep=""))) # check the metasym file
``` 

```{r}
dir(source_dir, recursive=T)
```

```{r}
# Files
#source1_path1 <- path <- paste(source_dir, "014_001_WCMC013-014_SeagrassPtPy2021_v7_1/01_Data/WCMC_013_014_SeagrassesPt_v7_1.shp", sep="")
source1_path1 <- path <- "./data/reg/eco/bh_veg_seagr/v01/proc/WCMC_013_014_SeagrassesPt_v7_1_buffer750m.shp" # reading in modified file (750m buffer around points using gis/GM21)
wcmc_sg_pt <- st_read(path)
(crs(wcmc_sg_pt))
glimpse(wcmc_sg_pt)
#source1_path2 <- path <- paste(source_dir, "014_001_WCMC013-014_SeagrassPtPy2021_v7_1/01_Data/WCMC013014-Seagrasses-Py-v7_1.shp", sep="")
source1_path2 <- path <- "./data/reg/eco/bh_veg_seagr/v01/proc/WCMC013014-Seagrasses-Py-v7_1_wiosymcrop.shp" # reading in cropped file gm21 due to prolems in r
wcmc_sg <- st_read(path) %>% st_transform((crs(wcmc_sg_pt)))
(crs(wcmc_sg))
wcmc_sg <- wcmc_sg %>% st_transform((crs(wcmc_sg_pt)))
glimpse(wcmc_sg)
# add files from dir as needed
```

### number of "data_raw" directorys?
```{r}
data_raw_metasym_num <- 1 # REPLACE - set total number of data_raw dir used (e.g. <- 1), to help in export section
```



# DATA sources
IMPORTANT Products in data are based on information in data_raw. Check that *_sourcesym.txt exists for each file to track acccumulated IDs. 

## DATA1: Grid 1km - loading all flavours (watermask 1, 0, NA) at once since same sourcesym apply
```{r}
data_dir <-  "./data/reg/grid/grid/v01/"  # input your data directory
dir(data_dir) # check what data is in this directory
```

```{r}
data1_file <- "grid_1km_v01.1.tif" 
data1_file2 <- "grid_1km_0_v01.1.tif" 
data1_file3 <- "grid_1km_na_v01.1.tif" 

(grid_1km <- raster(paste(data_dir, data1_file, sep=""))) # read and check your data
(grid_1km_0 <- raster(paste(data_dir, data1_file2, sep=""))) # read and check your data
(grid_1km_na <- raster(paste(data_dir, data1_file3, sep=""))) # read and check your data

(data1_sourcesym <- read_tsv(paste(data_dir, data1_file, "_sourcesym.txt", sep="")))
# check that source IDs exists and make sense

plot(grid_1km)
```

## DATA2: Grid 250m - loading all flavours (watermask 1, 0, NA) at once since same sourcesym apply
```{r}
data2_dir <- data_dir <-  "./data/reg/grid/grid/v01/"  # input your data directory
dir(data_dir) 
data2_file <- "grid_250m_v01.1.tif" # your data file
data2_file2 <- "grid_250m_0_v01.1.tif" # your data file
data2_file3 <- "grid_250m_na_v01.1.tif" # your data file

(grid_250m <- raster(paste(data_dir, data2_file, sep=""))) # read and check your data
(grid_250m_0 <- raster(paste(data_dir, data2_file2, sep=""))) # read and check your data
(grid_250m_na <- raster(paste(data_dir, data2_file3, sep=""))) # read and check your data

(data2_sourcesym <- read_tsv(paste(data_dir, data2_file, "_sourcesym.txt", sep="")))
# check so source IDs exists and make sense
# check that source IDs exists and make sense
#plot(grid_250m)
```


## DATA3: Allen seagrass - by-product of coral reef process... loading from data instead of redoing from source
```{r}
dir(data_dir <-  "./data/reg/eco/bh_coral/v01/")  # input your data directory
data_file <- "seagrass_allen_presence_250m_v01.1.tif" # your data file
(allen_sg_path <- path <- paste(data_dir, data_file, sep=""))
(allen_sg_250m_r <- raster(paste(data_dir, data_file, sep=""))) # read and check your data

(data3_sourcesym <- read_tsv(paste(path, "_sourcesym.txt", sep="")))
# check so source IDs exists and make sense
# check that source IDs exists and make sense
#plot(grid_250m)
```
## DATA 3: bounding box shapefile
```{r}
data4_dir <- data_dir <-  "./data/reg/grid/scope/v01/"  # input your data directory
data4_file <- "wiosym_grid_bounding_box_v01.shp" # your data file
data4_sourcesym <- read_tsv(paste(data_dir, data4_file, "_sourcesym.txt", sep=""))
data4_path <- data_path <- paste(data_dir, data4_file, sep="")
grid_poly <- st_read(data_path) # read in your data object if appropriate and perhaps change name to more informative...
#grid_poly
#plot(grid_poly)
```


## declare number of "data" files
```{r}
data_sourcesym_num <- 3 # REPLACE - set total number of data files (e.g. <- 2), to help in export section
```


#==========================================================================================================================#
# PROCESSING
#==========================================================================================================================#

# WCMC Polygons to 250m grid
```{r}
sf <- wcmc_sg
#glimpse(sf)
#unique(sf$pxlval)
sf <- sf %>% 
  mutate(pxlval = 1) %>% 
  dplyr::select(pxlval)


#problems with crop... doing in GIS and reloading ready file...
#sf <- sf %>% 
#  st_intersection(grid_poly)
#sf_crop <- st_crop(sf
#                    , xmin = 6, ymin = -42, xmax = 80, ymax = 18)
#glimpse(sf_crop)
#class(sf)
#sf <- sf_crop

input_path <- paste(work_path, "temp1.shp", sep="")
# write modified shapefile (value 1) to temp file)
st_write(sf,input_path, append = F) 

# write empty grid for gdalutil work:
wcmc_sg_250m_path <- output_path <- paste(work_path, "wcmc_sg_250m_na.tif", sep= "")
writeRaster(grid_250m_na, output_path,  overwrite=T, COMPRESS=LZW)

# 250m grid mapping
warp <- gdalUtils::gdal_rasterize(src_datasource = input_path,
                                        dst_filename = output_path,
                                        b = 1,
                                        at = T,
                                        a = "pxlval",
                                        output_Raster = TRUE,
)

r <- raster(output_path)
wcmc_sg_250m_path <- output_path <- paste(work_path, "wcmc_sg_250m_na_lzw.tif", sep= "")
writeRaster(r, output_path, overwrite=T, datatype = 'FLT4S', COMPRESS=LZW)
wcmc_sg_250m_r <- raster(output_path)

```

# WCMC Buffered Points (750m) 250m grid
```{r}
sf <- wcmc_sg_pt
#glimpse(sf)
#unique(sf$pxlval)
sf <- sf %>% 
  mutate(pxlval = 1) %>% 
  dplyr::select(pxlval)

glimpse(sf)
#class(sf)

(input_path <- paste(work_path, "temp2.shp", sep=""))
# write modified shapefile (value 1) to temp file)
st_write(sf,input_path, append = F) 

# write empty grid for gdalutil work:
wcmc_sg_pt_250m_path <- output_path <- paste(work_path, "wcmc_sg_pt_250m_na.tif", sep= "")
writeRaster(grid_250m_na, output_path,  overwrite=T, COMPRESS=LZW)

# 250m grid mapping
warp <- gdalUtils::gdal_rasterize(src_datasource = input_path,
                                        dst_filename = output_path,
                                        b = 1,
                                        at = T,
                                        a = "pxlval",
                                        output_Raster = TRUE,
)

r <- raster(output_path)
wcmc_sg_pt_250m_path <- output_path <- paste(work_path, "wcmc_sg_pt_250m_na_lzw.tif", sep= "")
writeRaster(r, output_path, overwrite=T, datatype = 'FLT4S', COMPRESS=LZW)
wcmc_sg_pt_250m_r <- raster(output_path)

```
# COMBINE seagrass 250m into one layer
```{r}

# set path to rasters to combine
(indata_paths <- c(wcmc_sg_250m_path, wcmc_sg_pt_250m_path, allen_sg_path))

# write empty grid for gdalutil output
sg_250m_path <- output_path <- paste(work_path, "sg_250m_na.tif", sep= "")
writeRaster(grid_250m_na, output_path,  overwrite=T, COMPRESS=LZW)


mosaic_r <- gdalUtils::mosaic_rasters(
  gdalfile = indata_paths,
  dst_dataset = output_path,
  output.vrt = NULL,
  output_Raster = FALSE,
  separate = FALSE,
  trim_margins = NULL,
  gdalwarp_index = 1,
  gdalwarp_params = list(r = "near"),
  #force_ot = "Int16",
  verbose = FALSE,
  #COMPRESS=LZW,
)

r <- raster(output_path)
crs(r) <- crs(grid_1km)
(r)
sg_250m_path <- output_path <- paste(work_path, "sg_250m_na_lzw.tif", sep= "")
writeRaster(r, output_path,  overwrite=T, datatype = 'FLT4S', COMPRESS=LZW)
sg_250m_r <- raster(output_path)

```

# Aggregate to 1km grid

```{r}
sg_1km_proportion <- aggregate(sg_250m_r, fact=4, fun=sum) / 16 #4*4=16, i.e. divide by the maximum cell number to get proportions
crs(sg_1km_proportion) <- crs(grid_1km)
(sg_1km_proportion)
plot(sg_1km_proportion) # ready to be exported if OK
```

# Map final layers to 1km standard grid with 0 values
```{r}

sg_1km_proportion_zero <- merge(sg_1km_proportion, grid_1km_0)

#sg_1km_proportion_relativefootprint_zero <- merge(.., grid_1km_o) # have not done this version yet... only for corals for now. 

plot(sg_1km_proportion_zero)
  
```

# Map final layers to 250m standard grid with 0 values
```{r}

sg_250m_zero <- merge(sg_250m_r, grid_250m_0)

plot(sg_250m_zero)
  
```




#==========================================================================================================================#
# EXPORT
#==========================================================================================================================#
# Write products to root directory when checked and ready to use elsewhere, accompanied by yourfilename.ext_sourcesym.txt 
# for each file to track source IDs as they accumulate between processes / data


# SOURCESYM
Create sourcesym file with all source IDs used (data and data_raw) 
### data raw sources (dir)
```{r}
print(paste("data_raw metasym files used = ", data_raw_metasym_num, sep=""))
(data_raw_sources <- tibble(id = c(source1))) # REPLACE / add all data_raw sources
```
### data sources (files)
```{r}
print(paste("data_sourcym files = ", data_sourcesym_num, sep=""))
data_sources <- data1_sourcesym %>% # add any "sourcesym" files applicable to this product (check indata section), keep adding sourcesym files as needed
  add_row(data2_sourcesym)%>%  
  add_row(data3_sourcesym)%>% 
  #add_row(data4_sourcesym)%>% 
  #add_row(data5_sourcesym)%>% 
  #add_row(data6_sourcesym)%>% 
  unique() %>% 
  print()
```
Sources combined
```{r}
(all_sources <- data_raw_sources %>% add_row(data_sources))
```
if a product only use a subset of the total sources, copy the section above and create individual files below (e.g. product1_sources <- ...)


# PRODUCT 1: SG 1km
read objects
```{r}
product_orig <- sg_1km_proportion_zero # input and check your object to be written to file
plot(product_orig)
```
set names
```{r}
scale <- "1km" # "1km" / "250m"
unit <- "proportion"  # add unit for product with original values (e.g. m, km, perc, presence, proportion
component_names <- read_tsv("./shiny_data_upload/component_names.txt") 
component_names_selected <- component_names %>%
  dplyr::filter(theme_folder == theme) %>% 
  dplyr::select(group, file_name, description) %>% 
  print(n = Inf)
```

```{r}
descriptive_name <- "seagrass_bed"  # name of product, for pressure/ecosystem components select from name list above
```
paths
```{r}
(product_orig_path <- paste(dest_path, descriptive_name, "_", unit, "_", scale, "_", version, d_version, ".tif", sep="" ))
```
Write to file
```{r}
# check datatype to be appropriate for values (flt or int, 'INT4S' option to..)
writeRaster(product_orig, product_orig_path, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
write_tsv(all_sources, paste(product_orig_path, "_sourcesym.txt", sep=""))  

```


# PRODUCT 2: SG 250m 
```{r}
product_orig <- sg_250m_zero # input and check your object to be written to file
plot(product_orig)

scale <- "250m" # "1km" / "250m"
unit <- "presence"  # add unit for product with original values (e.g. m, km, perc, presence, proportion

descriptive_name <- "seagrass_bed"  # name of product, for pressure/ecosystem components select from name list above

(product_orig_path <- paste(dest_path, descriptive_name, "_", unit, "_", scale, "_", version, d_version, ".tif", sep="" ))

```
Write to file
```{r}
# check datatype to be appropriate for values (flt or int, 'INT4S' option to..)
writeRaster(product_orig, product_orig_path, COMPRESS=LZW, datatype = 'INT4S', overwrite=TRUE)
write_tsv(all_sources, paste(product_orig_path, "_sourcesym.txt", sep=""))  
```




# SAVE SCRIPT
## 1. save your current R script (File/Save)
## 2. Save a carbon copy of the R script to proc_log (identical script to what is used to create the products exported
## Do not change the *.Rmd carbon copy in proc_log/ unless you change this version of the products (new versions are developed under process/r)

```{r}
# 2. run code below and go to File/Save As and paste link and name:
dest_path_script <- paste(getwd(), proc_path)
print(dest_path_script <- gsub(" .", "", dest_path_script)) # path
print(script_name_copy <- paste(gsub(".Rmd", "", script_name), d_version, ".Rmd", sep="")) # script name
```

#==========================================================================================================================#
# FINAL CHECK
#==========================================================================================================================#
## Developer check
 Do your own final check on the products and double check sourcesym file exists and are complete for all files in
 root -> Sign at top of script in about section (# Approved by:)

## External check 
 To ensure repeatability and quality, make sure one colleage can run the script When script is proven 
 repeatable/understandable and products/metadata look ok 
 -> Sign at top of script in about section (# Approved by:), and make any comments if needed
 There is also a work_log document in onedrive / gitHUB, check current arrangement with Swedish dev team

