### IMPORTANT: Because this script is an R Markdown script, the default directory is the location of this .Rmd file.
# To run the script from the project directory, in R Studio, you need to go to: Tools > Global Options > R Markdown
# and set "Evaluate chunks in directory" to "Project"
# -----------------------------------------------------------------------------------------------------------

### ABOUT
# ========================================================================================================= #
# Brief description: Mangrove map, adding additional data from madagascar to previous version (v1.0)
# Sripts/processes that needs to be updated prior?:
#
# Script by: gk, sgu, R 4.1.1
# Updated: Initial, org, Rver, comments
# Developer check: Initial, org, yymmdd, comments
# External check: Initial, org, yymmdd, comments
# ========================================================================================================= #
### PREPARATIONS
# ========================================================================================================= #

# Install packages
```{r, include = FALSE}
# list all packages that this script is dependent on, remove any redundant ones
x <- c("leaflet", "sf", "raster", "rgdal", "tidyverse", "gdalUtils")
#install.packages(x, dependencies=TRUE)
```

# Load packages
```{r, include = FALSE}
library(leaflet)
library(sf) 
library(raster)
library(rgdal)
library(tidyverse)
library(gdalUtils)
```

## Set version
```{r}
version = "v01" # relates to major iteration of WIOSym (v01 for grid 2021, v02 for grid 2022)
d_version = ".1"  # detailed version (you can add another level e.g. 0.0  -> v01.0.0)
seq = "s01" # sequential order - change accordingly if process is run in several scripts
```

## Set geographic scope
```{r}
read_tsv("./shiny_data_upload/locations.txt")  # prints available locations to choose from and their abbreviations
```
```{r}
location = "reg" # choose applicable 3 letter abbreviation from "location_val" -  reg for whole WIO area
```

## Set main theme
```{r}
folders_raw <- read_tsv("./shiny_data_upload/folders.txt")
folders_data <- read_tsv("./shiny_data_upload/folders_data.txt") # additional folders for data not applicable for data_raw 
folders <- folders_raw %>% add_row(folders_data)
(unique(folders["theme_folder"]))
```
```{r}
theme <- "eco" # Copy from "theme_folder" (e.g. "eco")
```

## Set Subtheme
```{r}
folders %>%
  filter(theme_folder == theme) %>% 
  dplyr::select(subtheme_folder, subtheme_name) %>% 
  print(n = Inf)
```
```{r}
subtheme <- "ch_mangrove" # copy from "subtheme_folder"
```

## Set paths
```{r}
(dest_path <- paste("./data", location, theme, subtheme, version, "", sep="/")) # path to final product
(work_path <- paste(dest_path, "proc/", sep=""))
(proc_path <- paste(dest_path, "proc_log/", sep=""))
(archive_path <- paste(dest_path, "_archive/", sep=""))
(script_path <- "./process/r/")
```

## Create directories
```{r}
dir.create(dest_path, recursive = TRUE)
dir.create(work_path, recursive = TRUE)
dir.create(proc_path, recursive = TRUE)
dir.create(archive_path, recursive = TRUE)
dir.create(script_path, recursive = TRUE)
```

## Save R Script
```{r}
name1 <- "" # "REPLACE_" if script need additional name... don't use unless needed to avoid duplication.
(script_path_windows <- paste(getwd(), "/process/r", sep=""))
(script_name <- paste(theme, "_", subtheme,"_", location, "_", name1, seq, "_", version, ".Rmd", sep=""))
```
 # copy path and name and use File/Save As in R studio to save rmd file
 
#==========================================================================================================================#
### INDATA
#==========================================================================================================================#
## Set DATA_RAW sources
# IMPORTANT all data harvested from data_raw needs to have a metasym.txt file in its root directory (not file specific)

# Data_raw DIR1: mangrove data from GMW, 2010 - 2016, downloaded from wcmc website
```{r}
source1_dir <- source_dir <-  "./data_raw/glo/eco/ch_mangrove/gmw/gk2105281014/"
source1 <- source_id <- "gk2105281014" # 
(read_tsv(paste(source_dir, source_id, "_metasym.txt", sep=""))) # check the metasym file
``` 
```{r}
dir(source_dir, recursive=F)
```
```{r}
# read file
source1_path <- path <- paste(source_dir, "GMW_001_GlobalMangroveWatch/01_Data/GMW_2016_v2.shp", sep="")
mangrove_2016 <- st_read(path) # REPLACE object_name
# add files from dir as needed
```
## Data_raw DIR2: TNC 2014
```{r}
source2_dir <- source_dir <-  "./data_raw/glo/eco/ch_mangrove/wcmc/gk2105281053/"
source2 <- source_id <- "gk2105281053" # 
(read_tsv(paste(source_dir, source_id, "_metasym.txt", sep=""))) # check the metasym file
``` 
```{r}
source2_path <- path <- paste(source_dir, "DataPack-14_001_TNC001_MangroveForestBiomass2014_v1/01_Data/14_001_TNC001_MangroveForestBiomass2014_v1.shp", sep="")
mangrove_2014 <- st_read(path) # REPLACE object_name
```
## Data_raw DIR3: wwf madagascar 1985-2018 mangrove data
```{r}
source3_dir <- source_dir <-  "./data_raw/reg/eco/ch_mangrove/wwf/gk2110061516/"
source3 <- source_id <- "gk2110061516" #  
(read_tsv(paste(source_dir, source_id, "_metasym.txt", sep=""))) # check the metasym file
``` 
```{r}
source3_path <- path <- paste(source_dir, "mada_mangrove_1985_2018/mada_mangrove_1985_2018.shp", sep="")
mangrove_2018 <- st_read(path) # REPLACE object_name
```

# declare number of "data_raw" directorys
```{r}
data_raw_metasym_num <- 3
```


## Set "DATA" sources
# IMPORTANT Products in data are based on information in data_raw. Check that *_sourcesym.txt exists for each file to track acccumulated IDs. 

# DATA 1: WIOSym standard grid 1km
```{r}
data_dir <-  "./data/reg/grid/grid/v01/"  # input your data directory
dir(data_dir) # check what data is in this directory
```
```{r}
data_file <- "grid_1km_v01.1.tif" # REPLACE - your data file
(data1_sourcesym <- read_tsv(paste(data_dir, data_file, "_sourcesym.txt", sep="")))
# check that source IDs exists and make sense
```
```{r}
(grid_1km <- raster(paste(data_dir, data_file, sep=""))) # read and check your data
```
```{r}
plot(grid_1km)
```

# DATA 2: grid 1km na  
```{r}
data_dir <-  "./data/reg/grid/grid/v01/"
dir(data_dir) # check what data is in this directory
```
```{r}
data_file <- "grid_1km_na_v01.1.tif" # your data file (must be under the root of data_dir, do not use files from proc/)
(data2_sourcesym <- read_tsv(paste(data_dir, data_file, "_sourcesym.txt", sep=""))) # REPLACE - check/change data number (e.g. data3_sourcesym)
# check that source IDs exists and make sense
```
```{r}
(grid_1km_na <- raster(paste(data_dir, data_file, sep="")))
```
```{r}
#plot(grid_1km_na) # REPLACE
```


# DATA 3: grid_250m
```{r}
data_dir <-  "./data/reg/grid/grid/v01/"
dir(data_dir) # check what data is in this directory
```
```{r}
data_file <- "grid_250m_v01.1.tif" # your data file (must be under the root of data_dir, do not use files from proc/)
(data3_sourcesym <- read_tsv(paste(data_dir, data_file, "_sourcesym.txt", sep=""))) # REPLACE - check/change data number (e.g. data3_sourcesym)
# check that source IDs exists and make sense
```
```{r}
(grid_250m <- raster(paste(data_dir, data_file, sep=""))) # REPLACE "object_name" - read in your data object
```
```{r}
plot(grid_250m) # REPLACE
```

# DATA 4: standard_grid 250m NA valu 
```{r}
data_file <- "grid_250m_na_v01.1.tif" # your data file (must be under the root of data_dir, do not use files from proc/)
(data4_sourcesym <- read_tsv(paste(data_dir, data_file, "_sourcesym.txt", sep=""))) # REPLACE - check/change data number (e.g. data3_sourcesym)
# check that source IDs exists and make sense
```
```{r}
(grid_250m_na <- raster(paste(data_dir, data_file, sep=""))) # REPLACE "object_name" - read in your data object
```
```{r}
plot(grid_250m_na)
```

# DATA 5: bounding box shapefile 
```{r}
data_dir <-  "./data/reg/grid/scope/v01/" 
dir(data_dir) # check what data is in this directory
```
```{r}
data_file <- "wiosym_grid_bounding_box_v01.shp" # your data file (must be under the root of data_dir, do not use files from proc/)
(data5_sourcesym <- read_tsv(paste(data_dir, data_file, "_sourcesym.txt", sep=""))) # REPLACE - check/change data number (e.g. data3_sourcesym)
# check that source IDs exists and make sense
```
```{r}
(grid_poly <- st_read(paste(data_dir, data_file, sep="")))
```
```{r}
plot(grid_poly)
```

# declare number of "data" files
```{r}
data_sourcesym_num <-  5 # REPLACE - set total number of data files (e.g. <- 2), to help in export section
```

#==========================================================================================================================#
### PROCESSING
#==========================================================================================================================#

# Map GMW mangrove polygons to 250m raster -----------------------------------------------
```{r}
indata_sf <- mangrove_2016
glimpse(indata_sf)
unique(indata_sf$pxlval)
class(indata_sf)
```
```{r}
# converting to old v00 script namings..
path_work <- work_path
indata_path <- source1_path

# convertion from shape to raster 

# write empty grid for gdalutil work
writeRaster(grid_250m_na, paste(path_work, "grid_250m_na_mangrove_gmw2016.tif", sep= ""),  overwrite=T, COMPRESS=LZW)
```
```{r}
# shapefile to 250m grid
sf_warp <- gdalUtils::gdal_rasterize(src_datasource = indata_path,
                                     dst_filename = paste(path_work, "grid_250m_na_mangrove_gmw2016.tif", sep= ""),
                                     b = 1,
                                     at = T,
                                     a = "pxlval",
                                     output_Raster = TRUE,
)


summary(sf_warp)
```

# Map TNC mangrove polygons to 250m raster
```{r}
indata_sf <- mangrove_2014
glimpse(indata_sf)
unique(indata_sf$pxlval)
class(indata_sf)

indata_sf <- indata_sf %>% 
  mutate(pxlval = 1)
```
```{r}
indata_path <- paste(path_work, "mangrove_tnc_value1.shp", sep= "")
st_write(indata_sf,indata_path) 

# write empty grid for gdalutil work
writeRaster(grid_250m_na, paste(path_work, "grid_250m_na_mangrove_tnc2014.tif", sep= ""),  overwrite=T, COMPRESS=LZW)
```
```{r}
# shape to 250m raster
sf_warp <- gdalUtils::gdal_rasterize(src_datasource = indata_path,
                                     dst_filename = paste(path_work, "grid_250m_na_mangrove_tnc2014.tif", sep= ""),
                                     b = 1,
                                     at = T,
                                     a = "pxlval",
                                     output_Raster = TRUE,
)


summary(sf_warp)
```

# Map wwf 2018 mangrove polygons to 250m raster
```{r}
indata_sf <- mangrove_2018
glimpse(indata_sf)
unique(indata_sf$pxlval)
class(indata_sf)

indata_sf <- indata_sf %>% 
  mutate(pxlval = 1)
```
```{r}
indata_path <- paste(path_work, "mangrove_2018_value1.shp", sep= "")
st_write(indata_sf,indata_path) 

# write empty grid for gdalutil work
writeRaster(grid_250m_na, paste(path_work, "grid_250m_na_mangrove_wwf2018.tif", sep= ""),  overwrite=T, COMPRESS=LZW)
```
```{r}
# shape to 250m raster
sf_warp <- gdalUtils::gdal_rasterize(src_datasource = indata_path,
                                     dst_filename = paste(path_work, "grid_250m_na_mangrove_wwf2018.tif", sep= ""),
                                     b = 1,
                                     at = T,
                                     a = "pxlval",
                                     output_Raster = TRUE,
)


summary(sf_warp)
```


# Combine all mangrove into one layer
```{r}
# paths to input rasters
gmw_250m <- paste(path_work, "grid_250m_na_mangrove_gmw2016.tif", sep= "")
tnc_250m <- paste(path_work, "grid_250m_na_mangrove_tnc2014.tif", sep= "")
wwf_250m <- paste(path_work, "grid_250m_na_mangrove_wwf2018.tif", sep= "")

# write empty grid for gdalutil output
outraster <- paste(path_work, "grid_250m_na_mangrove_combined.tif", sep= "")

writeRaster(grid_250m_na, outraster, overwrite=T, COMPRESS=LZW)

mangrove_250m_combined <- gdalUtils::mosaic_rasters(
  gdalfile = c(gmw_250m, tnc_250m, wwf_250m),
  dst_dataset = outraster,
  output.vrt = NULL,
  output_Raster = FALSE,
  separate = FALSE,
  trim_margins = NULL,
  gdalwarp_index = 1,
  gdalwarp_params = list(r = "near"),
  #force_ot = "Int16",
  verbose = FALSE,
  #COMPRESS=LZW,
)

mangrove_250m_combined <- raster(outraster)
crs(mangrove_250m_combined) <- crs(grid_1km)
(mangrove_250m_combined)
#plot(mangrove_250m_combined) # ready for export if ok
```

# Aggregate 250m mangrove polygons to 1km grid

```{r}
mangrove_1km_proportion <- aggregate(mangrove_250m_combined, fact=4, fun=sum) / 16 #4*4=16, i.e. divide by the maximum cell number to get proportions
crs(mangrove_1km_proportion) <- crs(grid_1km)
(mangrove_1km_proportion)
plot(mangrove_1km_proportion) # ready to be exported if OK
```
```{r}
#combined_mangrove_250m_agg1km_binary <- aggregate(mangrove_250m_combined, fact=4, fun=max) 
#writeRaster(combined_mangrove_250m_agg1km_binary, paste(path_work, "grid_1km_na_mangrove_combined.tif", sep= ""),  overwrite=T, COMPRESS=LZW)
```

# UNCERTAINTY MAP: Combine known range with shore connected pixels, and observed precenses to differentiate between "out of range, very good/good model (make seperate script? s02...)

#==========================================================================================================================#
### EXPORT
#==========================================================================================================================#
# Write products to root directory when checked and ready to use elsewhere, accompanied by yourfilename.ext_sourcesym.txt 
# for each file to track source IDs as they accumulate between processes / data


# SOURCESYM
# Create sourcesym file with all source IDs used (data and data_raw) 
# data raw sources (dir)
```{r}
print(paste("data_raw metasym files used = ", data_raw_metasym_num, sep=""))
(data_raw_sources <- tibble(id = c(source1, source2, source3))) 
```

# data sources (files)
```{r}
print(paste("data_sourcym files = ", data_sourcesym_num, sep=""))
data_sources <- data1_sourcesym %>% # add any "sourcesym" files applicable to this product (check indata section), keep adding sourcesym files as needed
  add_row(data2_sourcesym)%>%  
  add_row(data3_sourcesym)%>% 
  add_row(data4_sourcesym)%>% 
  add_row(data5_sourcesym)%>% 
  #add_row(data6_sourcesym)%>% 
  unique() %>% 
  print()
```
# Sources combined
```{r}
(all_sources <- data_raw_sources %>% add_row(data_sources))
```
# if a product only use a subset of the total sources, copy the section above and create individual files below (e.g. product1_sources <- ...)


# PRODUCT 1: 
# read objects

#outraster_lzw <- paste(path_work, "grid_250m_na_mangrove_combined_lzw.tif", sep="")
writeRaster(mangrove_250m_combined, outraster_lzw, overwrite=T, COMPRESS=LZW, datatype = 'INT4S')

writeRaster(mangrove_1km_proportion, paste(path_work, "grid_1km_na_mangrove_combined_proportion250m.tif", sep= ""),  overwrite=T, COMPRESS=LZW)


```{r}
product_orig <- mangrove_250m_combined # input and check your object to be written to file
#product_norm01 <- REPLACE # input and check your object to be written to file
#product_uncertainty <- REPLACE # input and check your object to be written to file
#plot(c(product_orig, product_norm01, product_uncertainty)) # use leaflet here instead?
```
# set names
```{r}
scale <- "250m" # "1km" / "250m"
unit <- "presence"  # add SI unit for product with original values (use short abbreviation e.g. meter = m, kilometer = km, percent = perc) 
component_names <- read_tsv("./shiny_data_upload/component_names.txt") 
component_names_selected <- component_names %>%
  dplyr::filter(theme_folder == theme) %>% 
  dplyr::select(group, file_name, description) %>% 
  print(n = Inf)
```
```{r}
descriptive_name <- "mangrove"  # name of product, for pressure/ecosystem components select from name list above
```
# paths
```{r}
(product_orig_path <- paste(dest_path, descriptive_name, "_", unit, "_", scale, "_", version, d_version, ".tif", sep="" ))
#(product_norm01_path <- paste(dest_path, descriptive_name, "_norm01_", scale, "_", version, d_version, ".tif", sep="" ))
#(product_uncertainty_path <- paste(dest_path, descriptive_name, "_uncertainty_", scale, "_", version, d_version, ".tif", sep="" ))
```
# Write to file
```{r}
writeRaster(product_orig, product_orig_path, COMPRESS=LZW, datatype = 'INT1U', overwrite=TRUE)
#writeRaster(product_norm01, product_norm01_path, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
#writeRaster(product_uncertainty, product_uncertainty_path, COMPRESS=LZW, datatype = 'INT1U', overwrite=TRUE)
# write sourcesym file, change "all_sources" to "selected_sources" if subset only for specific file
write_tsv(all_sources, paste(product_orig_path, "_sourcesym.txt", sep=""))  
#write_tsv(all_sources, paste(product_norm01_path, "_sourcesym.txt", sep=""))
#write_tsv(all_sources, paste(product_uncertainty_path, "_sourcesym.txt", sep=""))
```

# Export 2: 1km raster
# set names
```{r}
scale <- "1km" 
unit <- "proportion"  # add unit for product with original values (e.g. m, km, perc, presence, proportion

# paths 1km raster
```{r}
(product_orig_path <- paste(dest_path, descriptive_name, "_", unit, "_", scale, "_", version, d_version, ".tif", sep="" ))
#(product_norm01_path <- paste(dest_path, descriptive_name, "_norm01_", scale, "_", version, d_version, ".tif", sep="" ))
#(product_uncertainty_path <- paste(dest_path, descriptive_name, "_uncertainty_", scale, "_", version, d_version, ".tif", sep="" ))
```
# Write to file
```{r}
writeRaster(product_orig, product_orig_path, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
#writeRaster(product_norm01, product_norm01_path, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
#writeRaster(product_uncertainty, product_uncertainty_path, COMPRESS=LZW, datatype = 'INT1U', overwrite=TRUE)
# write sourcesym file, change "all_sources" to "selected_sources" if subset only for specific file
write_tsv(all_sources, paste(product_orig_path, "_sourcesym.txt", sep=""))  
#write_tsv(all_sources, paste(product_norm01_path, "_sourcesym.txt", sep=""))
#write_tsv(all_sources, paste(product_uncertainty_path, "_sourcesym.txt", sep=""))
```

### SAVE SCRIPT
# 1. save your current R script (File/Save)
# 2. Save a carbon copy of the R script to proc_log (identical script to what is used to create the products exported
# Do not change the *.Rmd carbon copy in proc_log/ unless you change this version of the products (new versions are developed under process/r)

```{r}
# 2. run code below and go to File/Save As and paste link and name:
dest_path_script <- paste(getwd(), proc_path)
print(dest_path_script <- gsub(" .", "", dest_path_script)) # path
print(script_name_copy <- paste(gsub(".Rmd", "", script_name), d_version, ".Rmd", sep="")) # script name
```

#==========================================================================================================================#
### FINAL CHECK
#==========================================================================================================================#
# Developer check
# Do your own final check on the products and double check sourcesym file exists and are complete for all files in root 
# -> Sign at top of script in about section (# Approved by:)

# External check 
# To ensure repeatability and quality, make sure one colleage can run the script
# When script is proven repeatable/understandable and products/metadata look ok 
# -> Sign at top of script in about section (# Approved by:), and make any comments if needed
# There is also a work_log document in onedrive / gitHUB, check current arrangement with Swedish dev team

