# ========================================================================================================= #
# IMPORTANT - To get started:
Because this script is an R Markdown script, the default directory is the location of this .Rmd file.
You must open this script from within the .Rproj file associated with your wiosym database for it to work.
If you haven't used R notebooks before (.Rmd), each of the code "chunks" can be executed by clicking the green play
button at the top right of the chunk.

To run the script from the project directory, in R Studio, you need to go to: Tools > Global Options > R Markdown
and set "Evaluate chunks in directory" to "Project", though for latest version of R/Rstudio it may not be needed...
# ========================================================================================================= #
# ABOUT
# ========================================================================================================= #
## Brief description: 
### Sripts/processes that needs to be updated prior?:
###
### Script by: Initial, org, Rver:
### Updated: Initial, org, Rver, comments
### Developer check: Initial, org, yymmdd, comments
### External check: Initial, org, yymmdd, comments
# ========================================================================================================= #
# INSTRUCTIONS
# ========================================================================================================= #    
## Style
  # Write code that is easy to follow and repeat.
  # Suggested R style is tidyverse. Style guide here https://style.tidyverse.org/index.html

## Metadata
  # All indata from "data_raw" must have a unique ID with *_metasym.txt file to go along, check before you add data.
  # All indata from and outputs to "data" must have a *_sourcesym.txt file with all (accumulated) sourceIDs used to create the layer.
  # The metasym files and tracking of those in the sourcesym files ensures correct sources for products and ISO metadata. This template helps you get it right..
  
## Naming and folders
  # Predefined standard names for main/sub folders and final products in two text files that can be added to in need.
  # ../shiny_data_upload/modify_txt_files_v01.2.xlsx and ../process/templates/modify_txt_files_data_v01.0.xlsx
  # Main version (e.g. v01 for 2021) follows dev cycles for all WIOSYm. For version control within v01 use V01.1, V01.2 etc..
  # More code/instructions to standardize folders and names under "set destinations" section
  # Each product dir have a "proc" folder. It's the temp space where all workfiles are stored. No final products here.  
  # All final products are stored under the root folder (e.g. ../data/reg/eco/bh_coral/v01) together with 
  # a nameofmyfile_sourcesym.txt file with all source IDs used. No other files here. move products to _archive/ dir if defunct 
# ========================================================================================================= # 
# PREPARATIONS
# ========================================================================================================= #

### Install packages
```{r, include = FALSE}
# list all packages that this script is dependent on, remove any redundant ones
x <- c("sf", "raster", "rgdal", "tidyverse")
#install.packages(x, dependencies=TRUE)
```

### Load packages
```{r, include = FALSE}
library(sf) 
library(raster)
library(rgdal)
library(tidyverse)
```

## Set version
```{r}
version = "v01" # relates to major iteration of WIOSym (v01 for grid 2021, v02 for grid 2022)
d_version = ".0"  # detailed version (you can add another level e.g. 0.0  -> v01.0.0)
seq = "s01" # sequential order - change accordingly if process is run in several scripts
```

## Set geographic scope
```{r}
read_tsv("./shiny_data_upload/locations.txt")  # prints available locations to choose from and their abbreviations
```

```{r}
location = "reg" # choose applicable 3 letter abbreviation from "location_val" -  reg for whole WIO area
```

## Set main theme
```{r}
folders_raw <- read_tsv("./shiny_data_upload/folders.txt")
folders_data <- read_tsv("./shiny_data_upload/folders_data.txt") # additional folders for data not applicable for data_raw 
folders <- folders_raw %>% add_row(folders_data)
(unique(folders["theme_folder"]))
```

```{r}
theme <- "eco" # Copy from "theme_folder" (e.g. "eco")
```

## Set Subtheme
```{r}
folders %>%
  filter(theme_folder == theme) %>% 
  dplyr::select(subtheme_folder, subtheme_name) %>% 
  print(n = Inf)
```

```{r}
subtheme <- "bh_hardsoft" # copy from "subtheme_folder"
```

## Set paths
```{r}
(dest_path <- paste("./data", location, theme, subtheme, version, "", sep="/")) # path to final product
(work_path <- paste(dest_path, "proc/", sep=""))
(proc_path <- paste(dest_path, "proc_log/", sep=""))
(archive_path <- paste(dest_path, "_archive/", sep=""))
(script_path <- "./process/r/")
```

## Create directories
```{r}
dir.create(dest_path, recursive = TRUE)
dir.create(work_path, recursive = TRUE)
dir.create(proc_path, recursive = TRUE)
dir.create(archive_path, recursive = TRUE)
dir.create(script_path, recursive = TRUE)
```

## Save R Script
```{r}
name1 <- "" # "REPLACE_" if script need additional name... don't use unless needed to avoid duplication.
(script_path_windows <- paste(getwd(), "/process/r", sep=""))
(script_name <- paste(theme, "_", subtheme,"_", location, "_", name1, seq, "_", version, ".Rmd", sep=""))
```
 # copy path and name and use File/Save As in R studio to save rmd file
 
#==========================================================================================================================#
# INDATA
#==========================================================================================================================#

### declare number of "data_raw" directorys above
```{r}
data_raw_metasym_num <-  0 # set total number of data_raw dir used (e.g. <- 1), to help in export section
```


# DATA sources
IMPORTANT Products in data are based on information in data_raw. Check that *_sourcesym.txt exists for each file to track acccumulated IDs. 

## DATA1: Grid 1km - loading all flavours (watermask 1, 0, NA) at once since identical sourcesym apply
```{r}
data_dir <-  "./data/reg/grid/grid/v01/"  # input your data directory
dir(data_dir) # check what data is in this directory

data1_file <- "grid_1km_v01.1.tif" 
data1_file2 <- "grid_1km_0_v01.1.tif" 
data1_file3 <- "grid_1km_na_v01.1.tif" 

(grid_1km_path <- path <- paste(data_dir, data1_file, sep=""))
(grid_1km <- raster(path)) # read and check your data
(grid_1km_0_path <- path <- paste(data_dir, data1_file2, sep=""))
(grid_1km_0 <- path <- raster(path)) # read and check your data
(grid_1km_na_path <- path <- paste(data_dir, data1_file3, sep=""))
(grid_1km_na <- path <- raster(path)) # read and check your data

data1_sourcesym <- read_tsv(paste(data_dir, data1_file, "_sourcesym.txt", sep=""))
# check that source IDs exists and make sense

plot(grid_1km)
```

## DATA2: Grid 250m - loading all flavours (watermask 1, 0, NA) at once since same sourcesym apply
```{r}
data2_dir <- data_dir <-  "./data/reg/grid/grid/v01/"  # input your data directory

data2_file <- "grid_250m_v01.1.tif" # your data file
data2_file2 <- "grid_250m_0_v01.1.tif" # your data file
data2_file3 <- "grid_250m_na_v01.1.tif" # your data file

(grid_250m_path <- path <- paste(data_dir, data2_file, sep=""))
(grid_250m <- raster(path)) # read and check your data
(grid_250m_0_path <- path <- paste(data_dir, data2_file2, sep=""))
(grid_250m_0 <- path <- raster(path)) # read and check your data
(grid_250m_na_path <- path <- paste(data_dir, data2_file3, sep=""))
(grid_250m_na <- path <- raster(path)) # read and check your data

(data2_sourcesym <- read_tsv(paste(data_dir, data2_file, "_sourcesym.txt", sep="")))
# check so source IDs exists and make sense
# check that source IDs exists and make sense
#plot(grid_250m)
```


## DATA 3: bounding box shapefile
```{r}
# "./data_raw/reg/env/geo_sub/dbseabed/gk2110151456/"

data3_dir <- data_dir <-  "./data/reg/env/geo_sub/v01/"  # input your data directory
dir(data_dir, recursive=F)

sub_stack_names <- list.files(data_dir, pattern = "proportion_1km_v01.0.tif$", recursive=F)
data3_sourcesym <- read_tsv(paste(data_dir, sub_stack_names[[2]], "_sourcesym.txt", sep=""))
sub_stack <- stack(paste(data_dir, list.files(data_dir, pattern = "proportion_1km_v01.0.tif$", recursive=F), sep=''))

plot(sub_stack[[1]])
```
## DATA4: depth
```{r}
data4_dir <- data_dir <-  "./data/reg/env/topo/v01/"  # input your data directory
dir(data_dir) # check what data is in this directory
data_file <- "bathymetry_mean_m_1km_v01.1.tif" # your data file
data4_sourcesym <- read_tsv(paste(data_dir, data_file, "_sourcesym.txt", sep=""))
data4_path <- data_path <- paste(data_dir, data_file, sep="")

depth_1km <- raster(data_path) # read in your data object if appropriate and perhaps change name to more informative...
plot(depth_1km)
```




## declare number of "data" files
```{r}
data_sourcesym_num <-  4 # set total number of data files (e.g. <- 2), to help in export section
```

#==========================================================================================================================#
# PROCESSING
#==========================================================================================================================#

## write function to classify into habitats..
```{r}

# linear rescale 0-1
soft <- sub_stack[[6]]
hard <- sub_stack[[4]]

r <- hard
r.min=cellStats(r,"min")
r.max=cellStats(r,"max")
hard <- (r-r.min)/(r.max-r.min)


r <- soft
r.min=cellStats(r,"min")
r.max=cellStats(r,"max")
soft <- (r-r.min)/(r.max-r.min)

# calculate habitat
shallow_soft <- (reclassify(depth_1km, c(-40, 1, 1,-7000, -40, 0)) * soft)
shallow_hard <- (reclassify(depth_1km, c(-40, 1, 1,-7000, -40, 0)) * hard)

shelf_soft <- reclassify(depth_1km, c(-200, -40, 1,-7000, -200, 0, -40, 1, 0)) * soft
shelf_hard <- reclassify(depth_1km, c(-200, -40, 1,-7000, -200, 0, -40, 1, 0)) * hard

continental_slope_soft <- reclassify(depth_1km, c(-1000, -200, 1,-7000, -1000, 0, -200, 1, 0)) * soft
continental_slope_hard <- reclassify(depth_1km, c(-1000, -200, 1,-7000, -1000, 0, -200, 1, 0)) * hard

abyssal_1000m_soft <- reclassify(depth_1km, c(-3000, -1000, 1,-7000, -3000, 0, -1000, 1, 0)) * soft
abyssal_1000m_hard <- reclassify(depth_1km, c(-3000, -1000, 1,-7000, -3000, 0, -1000, 1, 0)) * hard

abyssal_3000m_soft <- reclassify(depth_1km, c(-7000, -3000, 1,-3000, 1, 0)) * soft
abyssal_3000m_hard <- reclassify(depth_1km, c(-7000, -3000, 1,-3000, 1, 0)) * hard


```



#==========================================================================================================================#
# EXPORT
#==========================================================================================================================#
# Write products to root directory when checked and ready to use elsewhere, accompanied by yourfilename.ext_sourcesym.txt 
# for each file to track source IDs as they accumulate between processes / data


# SOURCESYM
Create sourcesym file with all source IDs used (data and data_raw) 
# data raw sources (dir)
```{r}
print(paste("data_raw metasym files used = ", data_raw_metasym_num, sep=""))
(data_raw_sources <- tibble(id = c())) # REPLACE / add all data_raw sources
```

## data sources (files)
```{r}
print(paste("data_sourcym files = ", data_sourcesym_num, sep=""))
data_sources <- data1_sourcesym %>% # add any "sourcesym" files applicable to this product (check indata section), keep adding sourcesym files as needed
  add_row(data2_sourcesym)%>%  
  add_row(data3_sourcesym)%>% 
  add_row(data4_sourcesym)%>% 
  #add_row(data5_sourcesym)%>% 
  #add_row(data6_sourcesym)%>% 
  unique() %>% 
  #print()
```
## Sources combined
```{r}
(all_sources <- (data_sources))
```
if a product only use a subset of the total sources, copy the section above and create individual files below (e.g. product1_sources <- ...)

# PRODUCT 1: 
## read objects
```{r}
product_orig <- shallow_soft # input and check your object to be written to file
product_orig2 <- shallow_hard # input and check your object to be written to file
product_orig3 <- shelf_soft # input and check your object to be written to file
product_orig4 <- shelf_hard # input and check your object to be written to file
product_orig5 <- continental_slope_soft # input and check your object to be written to file
product_orig6 <- continental_slope_hard # input and check your object to be written to file
product_orig7 <- abyssal_1000m_soft # input and check your object to be written to file
product_orig8 <- abyssal_1000m_hard # input and check your object to be written to file
product_orig9 <- abyssal_3000m_soft # input and check your object to be written to file
product_orig10 <- abyssal_3000m_hard # input and check your object to be written to file
plot(product_orig10)
```
## set names
```{r}
scale <- "1km" # "1km" / "250m"
unit <- "proportion"  # add unit for product with original values (e.g. m, km, perc, presence, proportion
component_names <- read_tsv("./shiny_data_upload/component_names.txt") 
component_names_selected <- component_names %>%
  dplyr::filter(theme_folder == theme) %>% 
  dplyr::select(group, file_name, description) %>% 
  print(n = Inf)
```

```{r}

descriptive_name <- "shallow_soft"  # name of product, for pressure/ecosystem components select from name list above
descriptive_name2 <- "shallow_hard"  # name of product, for pressure/ecosystem components select from name list above
descriptive_name3 <- "shelf_soft"  # name of product, for pressure/ecosystem components select from name list above
descriptive_name4 <- "shelf_hard"  # name of product, for pressure/ecosystem components select from name list above
descriptive_name5 <- "continental_slope_soft"  # name of product, for pressure/ecosystem components select from name list above
descriptive_name6 <- "continental_slope_hard"  # name of product, for pressure/ecosystem components select from name list above
descriptive_name7 <- "abyssal_1000m_soft"  # name of product, for pressure/ecosystem components select from name list above
descriptive_name8 <- "abyssal_1000m_hard"  # name of product, for pressure/ecosystem components select from name list above
descriptive_name9 <- "abyssal_3000m_soft"  # name of product, for pressure/ecosystem components select from name list above
descriptive_name10 <- "abyssal_3000m_hard"  # name of product, for pressure/ecosystem components select from name list above


```
## paths
```{r}
(product_orig_path <- paste(dest_path, descriptive_name, "_", unit, "_", scale, "_", version, d_version, ".tif", sep="" ))
(product_orig_path2 <- paste(dest_path, descriptive_name2, "_", unit, "_", scale, "_", version, d_version, ".tif", sep="" ))
(product_orig_path3 <- paste(dest_path, descriptive_name3, "_", unit, "_", scale, "_", version, d_version, ".tif", sep="" ))
(product_orig_path4 <- paste(dest_path, descriptive_name4, "_", unit, "_", scale, "_", version, d_version, ".tif", sep="" ))
(product_orig_path5 <- paste(dest_path, descriptive_name5, "_", unit, "_", scale, "_", version, d_version, ".tif", sep="" ))
(product_orig_path6 <- paste(dest_path, descriptive_name6, "_", unit, "_", scale, "_", version, d_version, ".tif", sep="" ))
(product_orig_path7 <- paste(dest_path, descriptive_name7, "_", unit, "_", scale, "_", version, d_version, ".tif", sep="" ))
(product_orig_path8 <- paste(dest_path, descriptive_name8, "_", unit, "_", scale, "_", version, d_version, ".tif", sep="" ))
(product_orig_path9 <- paste(dest_path, descriptive_name9, "_", unit, "_", scale, "_", version, d_version, ".tif", sep="" ))
(product_orig_path10 <- paste(dest_path, descriptive_name10, "_", unit, "_", scale, "_", version, d_version, ".tif", sep="" ))

```
## Write to file
```{r}
# check datatype to be appropriate for values (flt or int, 'INT4S' option to..)
writeRaster(product_orig, product_orig_path, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
writeRaster(product_orig2, product_orig_path2, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
writeRaster(product_orig3, product_orig_path3, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
writeRaster(product_orig4, product_orig_path4, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
writeRaster(product_orig5, product_orig_path5, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
writeRaster(product_orig6, product_orig_path6, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
writeRaster(product_orig7, product_orig_path7, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
writeRaster(product_orig8, product_orig_path8, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
writeRaster(product_orig9, product_orig_path9, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
writeRaster(product_orig10, product_orig_path10, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)


# write sourcesym file, change "all_sources" to "selected_sources" if subset only for specific file
write_tsv(all_sources, paste(product_orig_path, "_sourcesym.txt", sep=""))  
write_tsv(all_sources, paste(product_orig_path2, "_sourcesym.txt", sep=""))  
write_tsv(all_sources, paste(product_orig_path3, "_sourcesym.txt", sep=""))  
write_tsv(all_sources, paste(product_orig_path4, "_sourcesym.txt", sep=""))  
write_tsv(all_sources, paste(product_orig_path5, "_sourcesym.txt", sep=""))  
write_tsv(all_sources, paste(product_orig_path6, "_sourcesym.txt", sep=""))  
write_tsv(all_sources, paste(product_orig_path7, "_sourcesym.txt", sep=""))  
write_tsv(all_sources, paste(product_orig_path8, "_sourcesym.txt", sep=""))  
write_tsv(all_sources, paste(product_orig_path9, "_sourcesym.txt", sep=""))  
write_tsv(all_sources, paste(product_orig_path10, "_sourcesym.txt", sep=""))  
)  
```


# SAVE SCRIPT
## 1. save your current R script (File/Save)
## 2. Save a carbon copy of the R script to proc_log (identical script to what is used to create the products exported
## Do not change the *.Rmd carbon copy in proc_log/ unless you change this version of the products (new versions are developed under process/r)

```{r}
# 2. run code below and go to File/Save As and paste link and name:
dest_path_script <- paste(getwd(), proc_path)
print(dest_path_script <- gsub(" .", "", dest_path_script)) # path
print(script_name_copy <- paste(gsub(".Rmd", "", script_name), d_version, ".Rmd", sep="")) # script name
```

#==========================================================================================================================#
# FINAL CHECK
#==========================================================================================================================#
# Developer check
 Do your own final check on the products and double check sourcesym file exists and are complete for all files in root 
 -> Sign at top of script in about section (# Approved by:)

# External check 
 To ensure repeatability and quality, make sure one colleage can run the script
 When script is proven repeatable/understandable and products/metadata look ok 
 -> Sign at top of script in about section (# Approved by:), and make any comments if needed
 There is also a work_log document in onedrive / gitHUB, check current arrangement with Swedish dev team

