# =========================================================================================================
# Script by Edmond Sacre, SLU
# =========================================================================================================

# PREPARATIONS
```{r, include = FALSE}
# Install / load packages
#x <- c("sf", "raster", "tidyr", "dplyr")
#install.packages(x, dependencies=TRUE)
```

## Install packages
```{r, include = FALSE}
library(sp)
library(sf) 
library(raster)
library(fasterize)
library(tidyr)
library(dplyr)
library(rgdal)
library(tidyverse)
library(leaflet)
```

## View folders/themes
```{r}
folders_raw <- read.table("./shiny_data_upload/folders.txt", sep = "\t", header = TRUE)
folders_data <- read.table("./process/templates/folders_data.txt", sep = "\t", header = TRUE)
#folders <- folders_raw %>% add_row(folders_data)
```

## Define versions/directories
```{r}
version = "v01"
d_version = ".0"
location = "reg"
theme1 <- "act"
subtheme1 <- "nrg_oilgas"
name1 <- "oilgas_leakage"
(dest1_path <- paste("./data", location, theme1, subtheme1, name1, version, sep="/"))
(work1_path <- paste(dest1_path, "proc", sep="/"))
(proc1_path <- paste(dest1_path, "proc_log", sep="/"))
script_path <- "./process/r"
(script_path_windows <- paste(getwd(), "/process/r", sep=""))
(script_name <- paste(theme1, "_", subtheme1,"_", location, "_", name1,"_s01_",version, ".R", sep=""))
```

## Create directories
```{r}
dir.create(dest1_path, recursive = TRUE)
dir.create(work1_path, recursive = TRUE)
dir.create(proc1_path, recursive = TRUE)
```

## Read in WIOSym standard grid
```{r}
grid_1km <- raster("./data/reg/grid/grid/v01/grid_1km_v01.1.tif")
grid_1km_wgs <- projectRaster(grid_1km, crs = CRS("+proj=longlat +datum=WGS84 +no_defs"))
```

## Define sources
```{r}
num_sources <- 1  # write the number of sources you will use in this script (to be used later for exporting metadata)
  
source1_dir <-  "./data_raw/glo/act/nrg_oilgas/mapstand/es2110251315/"
source1_id <- "es2110251315"
source1_metasym <- read.table(paste0(source1_dir, source1_id, "_metasym.txt"), sep = "\t")
source1_metasym # check that metasym file is populated, and fix if not - before finalising the processing

source_list <- c(source1_id) # for multiple sources, should look like - c(source1_id, source2_id, source3_id) etc.
```

*****************************************************************
*****************************************************************
# PROCESSING

## NOTE: Some pre-processing steps were performed in ArcGIS for efficiency - Pipelines were converted to raster
## and Platform points were converted to raster, with cell values representing the number of platforms in each 1km cell

# Install processing specific packages
```{r, include = FALSE}
library(spatstat)
library(spatialEco)
library(mapview)
library(stars)
library(fasterize)
library(rgdal)
library(rgeos)
library(climateStability)
```

# Import data
```{r}
data_file <- "./data/reg/act/nrg_oilgas/oilgas_leakage/v01/proc/pipe_ras.tif"
pipe <- raster(data_file)
data_file <- "./data/reg/act/nrg_oilgas/oilgas_leakage/v01/proc/plat_ras.tif"
plat <- raster(data_file)
```

# Pipelines
```{r}
#pipe <- projectRaster(pipe, crs = CRS("+proj=longlat +datum=WGS84 +no_defs"))
#box <- c(xmin = 0, xmax = 88, ymin = -50,  ymax = 22)
pipe_sp <- rasterToPoints(pipe, spatial = T)
#pipe_sp <- raster::crop(pipe_sp, extent(box))
#mapview(pipe_sp)
```

```{r}
pipe_sp$value <- 1
pipe_kernel <- sp.kde(x = pipe_sp,
                        y = pipe_sp$value,
                        newdata = as.vector(extent(grid_1km)),
                        nr=2000,
                        nc=2000,
                        bw = 50000,
                        mask = FALSE,
                        standardize = T)
pipe_kernel <- resample(pipe_kernel, grid_1km)
pipe_kernel[is.na(grid_1km)==T] <- NA
```

# Set maximum value in map to 0.046 based on https://pubmed.ncbi.nlm.nih.gov/25057607/
# who estimate average accidental hydrocardbon discharges as 0.074 tonnes per pipeline per mile
# WIOSym grid is 1km, 1 mile = 1.61 km
# 0.074 * (1/1.61) = 0.046
```{r}
pipe_kernel <- pipe_kernel * 0.046
#mapview(pipe_kernel)
```

# Rigs/Platforms
```{r}
box <- c(xmin = -1000000, xmax = 8000000, ymin = -5000000,  ymax = 2300000)
plat_sp <- rasterToPoints(plat, spatial = T)
plat_sp <- raster::crop(plat_sp, extent(box))
#mapview(plat_sp)
```

```{r}
plat_sp$value <- 1
plat_kernel <- sp.kde(x = plat_sp,
                        y = plat_sp$value,
                        newdata = as.vector(extent(grid_1km)),
                        nr=2000,
                        nc=2000,
                        bw = 50000,
                        mask = FALSE,
                        standardize = T)
plat_kernel <- resample(plat_kernel, grid_1km)
plat_kernel[is.na(grid_1km)==T] <- NA
```

# Set maximum value in map to 0.045 based on https://pubmed.ncbi.nlm.nih.gov/25057607/
# who estimate average accidental hydrocardbon discharges as 0.045 tonnes per platform
```{r}
plat_kernel <- plat_kernel * 0.045
#mapview(plat_kernel)
```

# Create combine oil & gas leakage layer
```{r}
product1 <- pipe_kernel+plat_kernel
plot(product1)
```

*****************************************************************
*****************************************************************


# EXPORT

## Specify output directory and product file name

```{r}
(dest1_path_file1 <-paste0(dest1_path, "/", subtheme1, "_", name1, "_", version, d_version, ".tif"))
```

## Export metadata
```{r}
sourcesym <- data.frame(matrix(data = NA, nrow = num_sources, ncol = 1))
colnames(sourcesym) <- "id"

for(i in 1:num_sources){
  sourcesym[i,1] <- source_list[i]
}

sourcesym
```

```{r}
sourcesym_path <- paste0(dest1_path_file1, "_sourcesym.txt")
write.table(sourcesym, sourcesym_path, row.names = F, sep = "\t")
```

## Export product
```{r}
writeRaster(product1, dest1_path_file1, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
```

