---
title: "R Notebook"
output: html_notebook
---

```{r}
library(raster)
library(rio)
library(mapview)
library(RColorBrewer)
library(magick)

```


# SET VERSION
```{r}
# set version
version = "v01" # (1, 2, 3... relates to major releases for WIOSym, use v01.1, or v01.1.1 if you need additional version control use "detailed version" below)
d_version = ".5"  # detailed version (you can add another level e.g. 0.0  -> v01.0.0)

path_products <- "./products/"
path_cia <- paste0("./cia/reg/", version, d_version, "/")

```



```{r}
sens <- rio::import(paste0(path_cia, "matrix_june2022_shortnames.xlsx"))
rn <- sens[,"eco"]
sens <- dplyr::select(sens, -eco)
row.names(sens) <- rn
sens <- round(sens, digits = 2)
```

```{r}
grid <- raster("./data/reg/grid/grid/v01/grid_1km_v01.1.tif")*0
```

```{r}
l <- list.files("layers/layers1km")
l
```
# Combined ecocomponent map
```{r}
fulleco <- paste0(rownames(sens), ".tif")
ecofile <- l[l %in% fulleco]
econames <- tools::file_path_sans_ext(ecofile)
econames
```

```{r}
ecocomb <- grid
for(i in 1:length(ecofile)){
  temp <- raster(paste0("layers/layers1km/", ecofile[i]))
  ecocomb <- ecocomb + temp
  cat("\n", i, "min = ", minValue(temp), ", max = ", maxValue(temp))
}

writeRaster(ecocomb, filename = "./outputs/ecocomb.tif", datatype = 'INT4U', COMPRESS=LZW, overwrite = T)
m <- mapview(ecocomb, col.regions = brewer.pal(11, "Greens"), method = "ngb")
pngpath <- paste0("./outputs/ecocomponents_combined.png")
mapshot(m, 
        file = pngpath, 
        remove_controls = c("zoomControl", "layersControl", "homeButton", "scaleBar", "drawToolbar", "easyButton"))
img_crop <- image_crop(image_read(pngpath), "797x682+97+31")
image_write(img_crop, path = pngpath, format = "png")
```

```{r}
mapview(ecocomb, col.regions = brewer.pal(11, "Greens"), method = "ngb")
```

# Combined pressure map
```{r}
fullpres <- paste0(colnames(sens), ".tif")
presfile <- l[l %in% fullpres]
presnames <- tools::file_path_sans_ext(presfile)
presnames
```

```{r}
prescomb <- grid
for(i in 1:length(presfile)){
  temp <- raster(paste0("layers/layers1km/", presfile[i]))
  prescomb <- prescomb + temp
  cat("\n", i, "min = ", minValue(temp), ", max = ", maxValue(temp))
}

writeRaster(prescomb, filename = "./outputs/prescomb.tif", datatype = 'INT4U', COMPRESS=LZW, overwrite = T)
m <- mapview(prescomb, col.regions = brewer.pal(11, "Reds"), method = "ngb")
pngpath <- paste0("./outputs/pressures_combined.png")
mapshot(m, 
        file = pngpath, 
        remove_controls = c("zoomControl", "layersControl", "homeButton", "scaleBar", "drawToolbar", "easyButton"))
img_crop <- image_crop(image_read(pngpath), "797x682+97+31")
image_write(img_crop, path = pngpath, format = "png")
```

```{r}
mapview(prescomb, col.regions = brewer.pal(11, "Reds"), method = "ngb")
```

# Single component CIA
```{r}
ecocomponent <- "tuna"
cia <- grid
for (i in 1:length(presnames)){
  s <- sens[ecocomponent, presnames[i]]
  p <- raster(paste0("layers/layers1km/", presfile[i]))
  e <- raster(paste0("layers/layers1km/", ecocomponent, ".tif"))
  imp <- (s * p * e) / 100
  cia <- cia + imp
  cat("\n", "Iteration",i,"done")
}
```

```{r}
mapview(-cia, col.regions = brewer.pal(11, "RdYlGn"), method = "ngb")
```

# Full CIA
```{r, echo=FALSE, results='hide'}
cia <- grid

for(j in 1:length(econames)){
#for(j in 1:1){
  ecoimp <- grid
  ecocomponent <- econames[j]
for (i in 1:length(presnames)){
  s <- sens[ecocomponent, presnames[i]]
  p <- raster(paste0("layers/layers1km/", presfile[i]))
  e <- raster(paste0("layers/layers1km/", ecocomponent, ".tif"))
  imp <- (s * p * e) / 100
  ecoimp <- ecoimp + imp
  cia <- cia + imp
  cat("\n", "Iteration",j,"-",i,"done")
}
  writeRaster(ecoimp, filename = paste0("./outputs/raster/", econames[j], ".tif"), datatype = 'INT4U', COMPRESS=LZW, overwrite = T)
  m <- mapview(-ecoimp, col.regions = brewer.pal(11, "RdYlGn"), method = "ngb")
  pngpath <- paste0("./outputs/png/",econames[j], ".png")
  mapshot(m, 
          file = pngpath, 
          remove_controls = c("zoomControl", "layersControl", "homeButton", "scaleBar", "drawToolbar", "easyButton"))
  img_crop <- image_crop(image_read(pngpath), "797x682+97+31")
  image_write(img_crop, path = pngpath, format = "png")
}

writeRaster(cia, filename = "./outputs/cia.tif", datatype = 'INT4U', COMPRESS=LZW, overwrite = T)
m <- mapview(-cia, col.regions = brewer.pal(11, "RdYlGn"), method = "ngb")
pngpath <- paste0("./outputs/cia.png")
mapshot(m, 
        file = pngpath, 
        remove_controls = c("zoomControl", "layersControl", "homeButton", "scaleBar", "drawToolbar", "easyButton"))
img_crop <- image_crop(image_read(pngpath), "797x682+97+31")
image_write(img_crop, path = pngpath, format = "png")
```

```{r}
mapview(-cia, col.regions = brewer.pal(11, "RdYlGn"), method = "ngb")
```

# Rarity weighted CIA
```{r, echo=FALSE, results='hide'}
#cia <- grid
cia.stack <- stack(lapply(1:length(econames), function(i) grid))
names(cia.stack) <- econames

for(j in 1:length(econames)){
#for(j in 1:2){
  ecoimp <- grid
  ecocomponent <- econames[j]
  e <- raster(paste0("layers/layers1km/", ecocomponent, ".tif"))
for (i in 1:length(presnames)){
  s <- sens[ecocomponent, presnames[i]]
  p <- raster(paste0("layers/layers1km/", presfile[i]))
  imp <- (s * p * e) / 100
  ecoimp <- ecoimp + imp
  #cia <- cia + imp
  cat("\n", "Iteration",j,"-",i,"done")
}
  ecoimp[ecoimp > 0] <- (ecoimp[ecoimp > 0] / cellStats(e, stat = 'sum', na.rm = T)) # Add rarity weighting (and add multiplier so values are not so small?)
  cia.stack[[ecocomponent]] <- ecoimp
  m <- mapview(-ecoimp, col.regions = brewer.pal(11, "RdYlGn"), method = "ngb")
  pngpath <- paste0("./outputs/rarity/",econames[j], ".png")
  mapshot(m,
          file = pngpath,
          remove_controls = c("zoomControl", "layersControl", "homeButton", "scaleBar", "drawToolbar", "easyButton"))
  img_crop <- image_crop(image_read(pngpath), "797x682+97+31")
  image_write(img_crop, path = pngpath, format = "png")
}

cia <- calc(cia.stack, sum)
writeRaster(cia*10000000, filename = "./outputs/rarity_weighted_cia.tif", datatype = 'INT4U', COMPRESS=LZW, overwrite = T)
m <- mapview(-cia, col.regions = brewer.pal(11, "RdYlGn"), method = "ngb")
pngpath <- paste0("./outputs/rarity_weighted_cia.png")
mapshot(m,
        file = pngpath,
        remove_controls = c("zoomControl", "layersControl", "homeButton", "scaleBar", "drawToolbar", "easyButton"))
img_crop <- image_crop(image_read(pngpath), "797x682+97+31")
image_write(img_crop, path = pngpath, format = "png")
```

```{r}
mapview(-cia, col.regions = brewer.pal(11, "RdYlGn"), method = "ngb")
```

# Compress rasters
```{r}
r <- raster("outputs/cia.tif")
r <- (r / maxValue(r))*100
r <- round(r, digits = 0)
writeRaster(r, filename = paste0("outputs/cia_compress.tif"), datatype = 'INT2U', COMPRESS=LZW, overwrite = T)
```


```{r}
cellStats(cia, stat = 'sum', na.rm = T)
```

```{r}
r <- stack(lapply(1:length(econames), function(i) grid))
names(r) <- econames
#r <- brick(r)
```


```{r}
plot(cia.stack[["abyssalhard"]])
```

























# ECOSYSTEMT COMPONENTS - check layer projection, extent, resolution, and export correction
```{r}
layers <- rio::import("layer_list.xlsx") |> dplyr::filter(theme == "Ecosystem")
layers$file_name[is.na(layers$file_name)] <- "placeholder.tif"
layer_files <- layers$file_name
layer_files <- layer_files[c(1,2)] # run for only select layers

for(i in 1:length(layer_files)){
  temp <- raster(paste0("source_layers/", layer_files[i]))
  if(identical(crs(grid), crs(temp)) == F) {temp <- projectRaster(temp, grid)}
  x1 <- identical(extent(grid), extent(temp))
  x2 <- identical(crs(grid), crs(temp))
  x3 <- identical(res(grid), res(temp))
  cat("\nLayer", i ,"checked - ")
  if(x1&x2&x3){cat(layer_files[i], "passed checks")} else{cat(red("WARNING:", layer_files[i], "FAILED checks"))}
  writeRaster(temp, paste0("source_layers_fixed/", layer_files[i]), overwrite = T)
}
```

# PRESSURES - check layer projection, extent, resolution, and export correction
```{r}
layers <- rio::import("layer_list.xlsx") |> dplyr::filter(theme == "Pressure")
layers$file_name[is.na(layers$file_name)] <- "placeholder.tif"
layer_files <- layers$file_name
#layer_files <- layer_files[c(1,2)] # run for only select layers

for(i in 1:length(layer_files)){
  temp <- raster(paste0("source_layers/", layer_files[i]))
  if(identical(crs(grid), crs(temp)) == F) {temp <- projectRaster(temp, grid)}
  x1 <- identical(extent(grid), extent(temp))
  x2 <- identical(crs(grid), crs(temp))
  x3 <- identical(res(grid), res(temp))
  cat("\nLayer", i ,"checked - ")
  if(x1&x2&x3){cat(layer_files[i], "passed checks")} else{cat(red("WARNING:", layer_files[i], "FAILED checks"))}
  writeRaster(temp, paste0("source_layers_fixed/", layer_files[i]), overwrite = T)
}
```

# Reproject all rasters
```{r}
# Import grid
grid <- raster("grid/grid_1km_v01.1.tif")
l <- list.files("source_layers_fixed")
l <- l[c(1,50)] # run for only select layers

for(i in 1:length(l)){
  temp <- raster(paste0("source_layers_fixed/", l[i]))
  temp <- projectRaster(from = temp, to = grid_proj)
  writeRaster(temp, paste0("source_layers_reproject/", l[i]), overwrite = T)
  cat("\nIteration", i, "done")
}
```

# ECOSYSTEMT COMPONENTS - create geotif and images
```{r}
place <- raster("grid/grid_1km_v01.1.tif") * 0
layers <- rio::import("layer_list.xlsx") |> dplyr::filter(theme == "Ecosystem")
#layers <- filter(layers, title == "Deep pelagic" | title == "Midwater pelagic") # run for only select layers
layers$file_name[is.na(layers$file_name)] <- "placeholder.tif"
layer_files <- layers$file_name

# create folders / paths
path_tif <- dest_path <- paste0("./output_geotif/", paste0("_", version, d_version, "/eco/")) 
dest_path
if (!dir.exists(dest_path)){
  dir.create(dest_path, recursive = TRUE)
}

path_png <- dest_path <- paste0("./output_images/", paste0("_", version, d_version, "/eco/")) 
dest_path
if (!dir.exists(dest_path)){
  dir.create(dest_path, recursive = TRUE)
}



for(i in 1:length(layers$title)){
#i=1
  temp <- raster(paste0("source_layers_fixed/", layers$file_name[i]))
  temp[is.na(temp)] <- 0 #???
  temp[] <- scales::rescale(temp[], to = c(0, layers$max_value[i]))
  if(layers$file_name[i] == "placeholder.tif"){temp <- place}
  temp[is.na(temp)] <- 0
  temp <- round(temp, digits = 0)
  temp[is.na(grid)] <- NA
  
  #writeRaster(temp, paste0(path_tif, layers$title[i], " ",version, d_version,".tif"), datatype = 'INT2U', COMPRESS=LZW, overwrite=TRUE)
 
  png(paste0(path_png, layers$title[i], " ",version, d_version, ".png"), width = ncol(temp), height = nrow(temp))
  pal <- colorRampPalette(rev(c('#d53e4f','#f46d43','#fdae61','#fee08b','#ffffbf','#e6f598','#abdda4','#66c2a5','#3288bd')))
  plot(temp, maxpixels = ncell(temp), main = "", axes = FALSE, ylim=extent(temp)[3:4], 
       breaks = seq(0,100,1), col = pal(101), legend = F, box = F, colNA = "gray92")
  dev.off()
  cat("\nIteration", i, "done")
}

```

# PRESSURES - create geotif and images
```{r}
place <- raster("grid/grid_1km_v01.1.tif") * 0
layers <- rio::import("layer_list.xlsx") |> dplyr::filter(theme == "Pressure")
#layers <- filter(layers, title == "Temperature rise" | title == "Sea level rise") # run for only select layers
layers$file_name[is.na(layers$file_name)] <- "placeholder.tif"
layer_files <- layers$file_name


# create folders / paths
path_tif <- dest_path <- paste0("./output_geotif/", paste0("_", version, d_version, "/pres/")) 
dest_path
if (!dir.exists(dest_path)){
  dir.create(dest_path, recursive = TRUE)
}

path_png <- dest_path <- paste0("./output_images/", paste0("_", version, d_version, "/pres/")) 
dest_path
if (!dir.exists(dest_path)){
  dir.create(dest_path, recursive = TRUE)
}


for(i in 1:length(layers$title)){
  temp <- raster(paste0("source_layers_fixed/", layers$file_name[i]))
  temp[is.na(temp)] <- 0 #???
  temp[] <- scales::rescale(temp[], to = c(0, layers$max_value[i]))
  if(layers$file_name[i] == "placeholder.tif"){temp <- place}
  temp[is.na(temp)] <- 0
  temp <- round(temp, digits = 0)
  temp[is.na(grid)] <- NA
  
  writeRaster(temp, paste0(path_tif, layers$title[i], " ",version, d_version,".tif"), datatype = 'INT2U', COMPRESS=LZW, overwrite=TRUE)
 
  png(paste0(path_png, layers$title[i], " ",version, d_version, ".png"), width = ncol(temp), height = nrow(temp))
  pal <- colorRampPalette(rev(c('#d53e4f','#f46d43','#fdae61','#fee08b','#ffffbf','#e6f598','#abdda4','#66c2a5','#3288bd')))
  plot(temp, maxpixels = ncell(temp), main = "", axes = FALSE, ylim=extent(temp)[3:4], 
       breaks = seq(0,100,1), col = pal(101), legend = F, box = F, colNA = "gray92")
  dev.off()
  cat("\nIteration", i, "done")
}

```

# ECOSYSTEM COMPONENTS - produce metadata
```{r}
meta <- read.csv("metadata_template.csv", sep = ";")
layers <- rio::import("layer_list.xlsx") |> dplyr::filter(theme == "Ecosystem")
meta[1:nrow(layers),] <- NA

meta$Bandnumber <- 1:nrow(layers)
meta$Title <- layers$title
meta$Symphony.Category <- "Ecosystem"
meta$Symphony.Theme <- layers$subtheme

names(meta) <- gsub(x = names(meta), pattern = "\\.", replacement = " ")
write.csv2(meta, "output_multiband/ecosystems_metadata.csv", row.names = FALSE, quote = FALSE)
```

# PRESSURES - produce metadata
```{r}
meta <- read.csv("metadata_template.csv", sep = ";")
layers <- rio::import("layer_list.xlsx") |> dplyr::filter(theme == "Pressure")
meta[1:nrow(layers),] <- NA

meta$Bandnumber <- 1:nrow(layers)
meta$Title <- layers$title
meta$Symphony.Category <- "Pressure"
meta$Symphony.Theme <- layers$subtheme

names(meta) <- gsub(x = names(meta), pattern = "\\.", replacement = " ")
write.csv2(meta, "output_multiband/pressures_metadata.csv", row.names = FALSE, quote = FALSE)
```

# ECOSYSTEM COMPONENTS - create multiband tif
```{r}
place <- grid_proj * 0
layers <- rio::import("layer_list.xlsx") |> dplyr::filter(theme == "Ecosystem")
layers$file_name[is.na(layers$file_name)] <- "placeholder.tif"
layer_files <- layers$file_name

eco <- raster::stack()

for(i in 1:length(layer_files)){
  temp <- raster(paste0("source_layers_reproject/", layer_files[i]))
  temp[is.na(temp)] <- 0 # Add zeros before rescaling?
  temp[] <- scales::rescale(temp[], to = c(0, layers$max_value[i]))
  if(layer_files[i] == "placeholder.tif"){temp <- place}
  temp[is.na(grid_proj)] <- NA
  temp[is.na(temp)] <- 0
  temp <- round(temp, digits = 0)
  temp <- as.integer(temp)
  eco <- raster::stack(eco, temp)
  cat("\nIteration", i, "complete - Title:", layers$title[i], "- TIFF name:", layers$file_name[i])
}

writeRaster(eco, "output_multiband/ecosystems.tif", bandorder='BIL', datatype = 'INT2U', COMPRESS=LZW, overwrite=TRUE)
```

# PRESSURES - create multiband tif
```{r}
place <- grid_proj * 0
layers <- rio::import("layer_list.xlsx") |> dplyr::filter(theme == "Pressure")
layers$file_name[is.na(layers$file_name)] <- "placeholder.tif"
layer_files <- layers$file_name

eco <- raster::stack()

for(i in 1:length(layer_files)){
  temp <- raster(paste0("source_layers_reproject/", layer_files[i]))
  temp[is.na(temp)] <- 0 # Add zeros before rescaling?
  temp[] <- scales::rescale(temp[], to = c(0, layers$max_value[i]))
  if(layer_files[i] == "placeholder.tif"){temp <- place}
  temp[is.na(grid_proj)] <- NA
  temp[is.na(temp)] <- 0
  temp <- round(temp, digits = 0)
  temp <- as.integer(temp)
  eco <- raster::stack(eco, temp)
  cat("\nIteration", i, "complete - Title:", layers$title[i], "- TIFF name:", layers$file_name[i])
}

writeRaster(eco, "output_multiband/pressures.tif", bandorder='BIL', datatype = 'INT2U', COMPRESS=LZW, overwrite=TRUE)
```

