---
title: "R Notebook"
output: html_notebook
---

```{r}
library(raster)
library(dplyr)
library(tidyr)
library(rio)
library(scales)
library(crayon)
library(mapview)
library(RColorBrewer)
library(rgdal)
library(gdalUtils)

# Import grid
grid <- raster("grid/grid_1km_v01.1.tif")
meta <- rio::import("layer_list.xlsx")
meta$file_name[is.na(meta$file_name)] <- "placeholder.tif"
```


# SET VERSION
```{r}
# set version
version = "v01" # (1, 2, 3... relates to major releases for WIOSym, use v01.1, or v01.1.1 if you need additional version control use "detailed version" below)
d_version = ".5"  # detailed version (you can add another level e.g. 0.0  -> v01.0.0)
```

# ECOSYSTEMT COMPONENTS - check layer projection, extent, resolution, and export correction
```{r}
layers <- rio::import("layer_list.xlsx") |> dplyr::filter(theme == "Ecosystem")
layers$file_name[is.na(layers$file_name)] <- "placeholder.tif"
layer_files <- layers$file_name
layer_files <- layer_files[c(1,2)] # run for only select layers

for(i in 1:length(layer_files)){
  temp <- raster(paste0("source_layers/", layer_files[i]))
  if(identical(crs(grid), crs(temp)) == F) {temp <- projectRaster(temp, grid)}
  x1 <- identical(extent(grid), extent(temp))
  x2 <- identical(crs(grid), crs(temp))
  x3 <- identical(res(grid), res(temp))
  cat("\nLayer", i ,"checked - ")
  if(x1&x2&x3){cat(layer_files[i], "passed checks")} else{cat(red("WARNING:", layer_files[i], "FAILED checks"))}
  writeRaster(temp, paste0("source_layers_fixed/", layer_files[i]), overwrite = T)
}
```

# PRESSURES - check layer projection, extent, resolution, and export correction
```{r}
layers <- rio::import("layer_list.xlsx") |> dplyr::filter(theme == "Pressure")
layers$file_name[is.na(layers$file_name)] <- "placeholder.tif"
layer_files <- layers$file_name
#layer_files <- layer_files[c(1,2)] # run for only select layers

for(i in 1:length(layer_files)){
  temp <- raster(paste0("source_layers/", layer_files[i]))
  if(identical(crs(grid), crs(temp)) == F) {temp <- projectRaster(temp, grid)}
  x1 <- identical(extent(grid), extent(temp))
  x2 <- identical(crs(grid), crs(temp))
  x3 <- identical(res(grid), res(temp))
  cat("\nLayer", i ,"checked - ")
  if(x1&x2&x3){cat(layer_files[i], "passed checks")} else{cat(red("WARNING:", layer_files[i], "FAILED checks"))}
  writeRaster(temp, paste0("source_layers_fixed/", layer_files[i]), overwrite = T)
}
```

# Reproject all rasters
```{r}
# Import grid
grid <- raster("grid/grid_1km_v01.1.tif")
l <- list.files("source_layers_fixed")
l <- l[c(1,50)] # run for only select layers

for(i in 1:length(l)){
  temp <- raster(paste0("source_layers_fixed/", l[i]))
  temp <- projectRaster(from = temp, to = grid_proj)
  writeRaster(temp, paste0("source_layers_reproject/", l[i]), overwrite = T)
  cat("\nIteration", i, "done")
}
```

# ECOSYSTEMT COMPONENTS - create images
```{r}
place <- raster("grid/grid_1km_v01.1.tif") * 0
layers <- rio::import("layer_list.xlsx") |> dplyr::filter(theme == "Ecosystem")
#layers <- filter(layers, title == "Deep pelagic" | title == "Midwater pelagic") # run for only select layers
layers$file_name[is.na(layers$file_name)] <- "placeholder.tif"
layer_files <- layers$file_name

# create folders / paths
path_tif <- dest_path <- paste0("./output_geotif/", paste0("_", version, d_version, "/eco/")) 
dest_path
if (!dir.exists(dest_path)){
  dir.create(dest_path, recursive = TRUE)
}

path_png <- dest_path <- paste0("./output_images/", paste0("_", version, d_version, "/eco/")) 
dest_path
if (!dir.exists(dest_path)){
  dir.create(dest_path, recursive = TRUE)
}



for(i in 1:length(layers$title)){
#i=1
  temp <- raster(paste0("source_layers_fixed/", layers$file_name[i]))
  temp[is.na(temp)] <- 0 #???
  temp[] <- scales::rescale(temp[], to = c(0, layers$max_value[i]))
  if(layers$file_name[i] == "placeholder.tif"){temp <- place}
  temp[is.na(temp)] <- 0
  temp <- round(temp, digits = 0)
  temp[is.na(grid)] <- NA
  
  writeRaster(temp, paste0(path_tif, layers$title[i], " ",version, d_version,".tif"), datatype = 'INT2U', COMPRESS=LZW, overwrite=TRUE)
 
  png(paste0(path_png, layers$title[i], " ",version, d_version, ".png"), width = ncol(temp), height = nrow(temp))
  pal <- colorRampPalette(rev(c('#d53e4f','#f46d43','#fdae61','#fee08b','#ffffbf','#e6f598','#abdda4','#66c2a5','#3288bd')))
  plot(temp, maxpixels = ncell(temp), main = "", axes = FALSE, ylim=extent(temp)[3:4], 
       breaks = seq(0,100,1), col = pal(101), legend = F, box = F, colNA = "gray92")
  dev.off()
  cat("\nIteration", i, "done")
}

```

# PRESSURES - create images
```{r}
place <- raster("grid/grid_1km_v01.1.tif") * 0
layers <- rio::import("layer_list.xlsx") |> dplyr::filter(theme == "Pressure")
#layers <- filter(layers, title == "Temperature rise" | title == "Sea level rise") # run for only select layers
layers$file_name[is.na(layers$file_name)] <- "placeholder.tif"
layer_files <- layers$file_name


# create folders / paths
path_tif <- dest_path <- paste0("./output_geotif/", paste0("_", version, d_version, "/pres/")) 
dest_path
if (!dir.exists(dest_path)){
  dir.create(dest_path, recursive = TRUE)
}

path_png <- dest_path <- paste0("./output_images/", paste0("_", version, d_version, "/pres/")) 
dest_path
if (!dir.exists(dest_path)){
  dir.create(dest_path, recursive = TRUE)
}


for(i in 1:length(layers$title)){
  temp <- raster(paste0("source_layers_fixed/", layers$file_name[i]))
  temp[is.na(temp)] <- 0 #???
  temp[] <- scales::rescale(temp[], to = c(0, layers$max_value[i]))
  if(layers$file_name[i] == "placeholder.tif"){temp <- place}
  temp[is.na(temp)] <- 0
  temp <- round(temp, digits = 0)
  temp[is.na(grid)] <- NA
  
  writeRaster(temp, paste0(path_tif, layers$title[i], " ",version, d_version,".tif"), datatype = 'INT2U', COMPRESS=LZW, overwrite=TRUE)
 
  png(paste0(path_png, layers$title[i], " ",version, d_version, ".png"), width = ncol(temp), height = nrow(temp))
  pal <- colorRampPalette(rev(c('#d53e4f','#f46d43','#fdae61','#fee08b','#ffffbf','#e6f598','#abdda4','#66c2a5','#3288bd')))
  plot(temp, maxpixels = ncell(temp), main = "", axes = FALSE, ylim=extent(temp)[3:4], 
       breaks = seq(0,100,1), col = pal(101), legend = F, box = F, colNA = "gray92")
  dev.off()
  cat("\nIteration", i, "done")
}

```

# ECOSYSTEM COMPONENTS - produce metadata
```{r}
meta <- read.csv("metadata_template.csv", sep = ";")
layers <- rio::import("layer_list.xlsx") |> dplyr::filter(theme == "Ecosystem")
meta[1:nrow(layers),] <- NA

meta$Bandnumber <- 1:nrow(layers)
meta$Title <- layers$title
meta$Symphony.Category <- "Ecosystem"
meta$Symphony.Theme <- layers$subtheme

names(meta) <- gsub(x = names(meta), pattern = "\\.", replacement = " ")
write.csv2(meta, "output_multiband/ecosystems_metadata.csv", row.names = FALSE, quote = FALSE)
```

# PRESSURES - produce metadata
```{r}
meta <- read.csv("metadata_template.csv", sep = ";")
layers <- rio::import("layer_list.xlsx") |> dplyr::filter(theme == "Pressure")
meta[1:nrow(layers),] <- NA

meta$Bandnumber <- 1:nrow(layers)
meta$Title <- layers$title
meta$Symphony.Category <- "Pressure"
meta$Symphony.Theme <- layers$subtheme

names(meta) <- gsub(x = names(meta), pattern = "\\.", replacement = " ")
write.csv2(meta, "output_multiband/pressures_metadata.csv", row.names = FALSE, quote = FALSE)
```

# ECOSYSTEM COMPONENTS - create multiband tif
```{r}
place <- grid_proj * 0
layers <- rio::import("layer_list.xlsx") |> dplyr::filter(theme == "Ecosystem")
layers$file_name[is.na(layers$file_name)] <- "placeholder.tif"
layer_files <- layers$file_name

eco <- raster::stack()

for(i in 1:length(layer_files)){
  temp <- raster(paste0("source_layers_reproject/", layer_files[i]))
  temp[is.na(temp)] <- 0 # Add zeros before rescaling?
  temp[] <- scales::rescale(temp[], to = c(0, layers$max_value[i]))
  if(layer_files[i] == "placeholder.tif"){temp <- place}
  temp[is.na(grid_proj)] <- NA
  temp[is.na(temp)] <- 0
  temp <- round(temp, digits = 0)
  temp <- as.integer(temp)
  eco <- raster::stack(eco, temp)
  cat("\nIteration", i, "complete - Title:", layers$title[i], "- TIFF name:", layers$file_name[i])
}

writeRaster(eco, "output_multiband/ecosystems.tif", bandorder='BIL', datatype = 'INT2U', COMPRESS=LZW, overwrite=TRUE)
```

# PRESSURES - create multiband tif
```{r}
place <- grid_proj * 0
layers <- rio::import("layer_list.xlsx") |> dplyr::filter(theme == "Pressure")
layers$file_name[is.na(layers$file_name)] <- "placeholder.tif"
layer_files <- layers$file_name

eco <- raster::stack()

for(i in 1:length(layer_files)){
  temp <- raster(paste0("source_layers_reproject/", layer_files[i]))
  temp[is.na(temp)] <- 0 # Add zeros before rescaling?
  temp[] <- scales::rescale(temp[], to = c(0, layers$max_value[i]))
  if(layer_files[i] == "placeholder.tif"){temp <- place}
  temp[is.na(grid_proj)] <- NA
  temp[is.na(temp)] <- 0
  temp <- round(temp, digits = 0)
  temp <- as.integer(temp)
  eco <- raster::stack(eco, temp)
  cat("\nIteration", i, "complete - Title:", layers$title[i], "- TIFF name:", layers$file_name[i])
}

writeRaster(eco, "output_multiband/pressures.tif", bandorder='BIL', datatype = 'INT2U', COMPRESS=LZW, overwrite=TRUE)
```

