---
title: "R Notebook"
output: html_notebook
---

# ========================================================================================================= #
# Script by Edmond Sacre, SLU
# This script creates rasters from the Global Fishing Watch raw data
# ========================================================================================================= #

# PREPARATIONS

```{r, include = FALSE}
# Install / load packages
#x <- c("sp", "raster", "tidyr", "dplyr")
#install.packages(x, dependencies=TRUE)
```

## Install packages
```{r, include = FALSE}
library(sf) 
library(raster)
library(fasterize)
library(tidyr)
library(dplyr)
library(rgdal)
library(tidyverse)
library(leaflet)
```

## View folders/themes
```{r}
folders_raw <- read.table("./shiny_data_upload/folders.txt", sep = "\t", header = TRUE)
folders_data <- read.table("./process/templates/folders_data.txt", sep = "\t", header = TRUE) 
#folders <- folders_raw %>% add_row(folders_data)
```

## Define versions/directories
```{r}
version = "v01"
d_version = ".0"
location = "glo"
theme1 <- "act"
subtheme1 <- "fishing_com"
name1 <- "gfw_effort"
(dest1_path <- paste("./data", location, theme1, subtheme1, name1, version, sep="/"))
(work1_path <- paste(dest1_path, "proc", sep="/"))
(proc1_path <- paste(dest1_path, "proc_log", sep="/"))
script_path <- "./process/r"
(script_path_windows <- paste(getwd(), "/process/r", sep=""))
(script_name <- paste(theme1, "_", subtheme1,"_", location, "_", name1,"_s01_",version, ".R", sep=""))
```

## Create directories
```{r}
#dir.create(dest1_path, recursive = TRUE)
#dir.create(work1_path, recursive = TRUE)
#dir.create(proc1_path, recursive = TRUE)
```

## Read in WIOSym standard grid
```{r}
grid_1km <- raster("./data/reg/grid/grid/v01/grid_1km_v01.1.tif")
```

```{r}
extent(grid_1km)
```

## Define sources
```{r}
source1_dir <-  "./data_raw/glo/act/fishing_com/gfw/es2105270900/"
source1_id <- "es2105270900"
source1_metasym <- read.table(paste0(source1_dir, source1_id, "_metasym.txt"), sep = "\t")
#source1_metasym # check that metasym file is populated, and fix if not - before finalising the processing
```

# PROCESSING

## Import specific year

```{r}
# gfw_year <- 2020
# 
# base_grid <- grid_1km
# base_grid[base_grid == 1] <- 0
# 
# file_name = paste0("./data_raw/glo/act/fishing_com/gfw/es2105270900/fleet-daily-csvs-100-v2-", gfw_year)
# 
# data_list <- list.files(file_name, full.names = TRUE)
# data_file <- data_list[1]
# 
# fishing <- read.csv(data_file)

#summary(fishing$geartype)
```


```{r}
#levels(droplevels(fishing$geartype))
```

```{r}
#fishing <- filter(fishing, geartype ==  "other_purse_seines" | geartype ==  "tuna_purse_seines")
```


 [1] "dredge_fishing"     "drifting_longlines" "fishing"            "fixed_gear"         "other_purse_seines" "other_seines"      
 [7] "pole_and_line"      "pots_and_traps"     "purse_seines"       "seiners"            "set_gillnets"       "set_longlines"     
[13] "squid_jigger"       "trawlers"           "trollers"           "tuna_purse_seines"

### Categorizatin of gear types

"fixed_gear" name includes: "fixed_gear", "set_longlines", "set_gillnets", "pots_and_traps"
"other_fishing" name includes: "trollers", "pole_and_line",  


## Import all years and combine

### First choose a gear type
```{r}
gear_name <- "drifting_longlines"  #for the file name
```

### Reproject grid raster if necessary (encountered some problems using standard projection)
```{r}
grid_1km <- projectRaster(grid_1km, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs", method = "ngb")
```

```{r, include = FALSE}

for(j in 2020:2020){

  gfw_year <- j
  
  base_grid <- grid_1km
  
  base_grid[base_grid == 1] <- 0
  
  #data_list <- list.files("./data_raw/glo/act/fishing_com/gfw/es2105270900/fleet-daily-csvs-100-v2-2017", full.names = TRUE)
  
  file_name = paste0("./data_raw/glo/act/fishing_com/gfw/es2105270900/fleet-daily-csvs-100-v2-", j)
  data_list <- list.files(file_name, full.names = TRUE)
  
  #######
  
  time1 <- Sys.time()
  
  nit <- length(data_list)
  
  for(i in 1:nit) {
  
    t0 <- proc.time()
    
    data_file <- data_list[i]
    fishing <- read.csv(data_file)
    
    fishing <- filter(fishing, geartype == "drifting_longlines") # EDIT HERE
    
    fishing$cell_ll_lon_fact <- as.factor(fishing$cell_ll_lon)
    fishing$cell_ll_lat_fact <- as.factor(fishing$cell_ll_lat)
    
    fishing_sum <- fishing %>% group_by(cell_ll_lon_fact, cell_ll_lat_fact) %>% summarise(sum_fishing_hours = sum(fishing_hours))
    
    fishing_sum$x <- as.numeric(levels(fishing_sum$cell_ll_lon_fact))[fishing_sum$cell_ll_lon_fact]
    fishing_sum$y <- as.numeric(levels(fishing_sum$cell_ll_lat_fact))[fishing_sum$cell_ll_lat_fact]
    fishing_sum$z <- fishing_sum$sum_fishing_hours
    
    fishing_sum <- fishing_sum %>% ungroup() %>% select(x, y, z)
    
    fishing_sum <- filter(fishing_sum, x > 5 & x < 82)
    fishing_sum <- filter(fishing_sum, y > -43 & y < 20)
    
    if(nrow(fishing_sum) == 0) {
      
      t1 <- proc.time() - t0
      dt <- as.numeric(t1[3])
      cat("\n", "Iteration", i, "complete", " - processing time", dt, "seconds")
      
    }
    
    else{
      
      fishing_ras <- rasterFromXYZ(fishing_sum, res=c(0.01,0.01), crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
      #fishing_ras <- rasterFromXYZ(fishing_sum, res=c(0.01,0.01), crs = "+proj=cea +lat_ts=-12 +lon_0=12 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs")
      
      fishing_ras <- resample(fishing_ras, grid_1km, method = "ngb")
      
      fishing_ras <- crop(fishing_ras, grid_1km)
      
      fishing_ras[is.na(fishing_ras) == TRUE] <- 0
      
      base_grid <- fishing_ras + base_grid
      
      # remove (necessary)
      #rm(fishing)
      #rm(fishing_sum)
      #rm(fishing_ras)
      
      t1 <- proc.time() - t0
      dt <- as.numeric(t1[3])
      cat("\n", "Iteration", i, "complete", " - processing time", dt, "seconds")
      
      }
    
  }
  
  time2 <- Sys.time()
  
  total.time <- time2 - time1
  total.time
  
  
  store_base_grid <- base_grid
  base_grid[base_grid == 0] <- NA
  
  
  
  
  # EXPORT
  
  grid_1km_original <- raster("./data/reg/grid/grid/v01/grid_1km_v01.1.tif")
  
  base_grid_proj <- projectRaster(base_grid, crs = crs(grid_1km_original), method = "ngb")
  
  (dest1_path_file1 <-paste(dest1_path, "/", subtheme1, "_", gear_name, "_", gfw_year, "_1km_", version, d_version, ".tif", sep=""))
  
  writeRaster(base_grid_proj, dest1_path_file1, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)

}
```

## Import yearly maps and combine

```{r}
grid_1km_original <- raster("./data/reg/grid/grid/v01/grid_1km_v01.1.tif")
```

```{r}
gfw2012 <- raster("./data/glo/act/fishing_com/gfw_effort/v01/fishing_com_gfw_effort_2012_1km_v01.0.tif")
gfw2013 <- raster("./data/glo/act/fishing_com/gfw_effort/v01/fishing_com_gfw_effort_2013_1km_v01.0.tif")
gfw2014 <- raster("./data/glo/act/fishing_com/gfw_effort/v01/fishing_com_gfw_effort_2014_1km_v01.0.tif")
gfw2015 <- raster("./data/glo/act/fishing_com/gfw_effort/v01/fishing_com_gfw_effort_2015_1km_v01.0.tif")
gfw2016 <- raster("./data/glo/act/fishing_com/gfw_effort/v01/fishing_com_gfw_effort_2016_1km_v01.0.tif")
gfw2017 <- raster("./data/glo/act/fishing_com/gfw_effort/v01/fishing_com_gfw_effort_2017_1km_v01.0.tif")
gfw2018 <- raster("./data/glo/act/fishing_com/gfw_effort/v01/fishing_com_gfw_effort_2018_1km_v01.0.tif")
gfw2019 <- raster("./data/glo/act/fishing_com/gfw_effort/v01/fishing_com_gfw_effort_2019_1km_v01.0.tif")
gfw2020 <- raster("./data/glo/act/fishing_com/gfw_effort/v01/fishing_com_gfw_effort_2020_1km_v01.0.tif")
```

```{r}
gfw2012[is.na(gfw2012) == TRUE] <- 0
gfw2013[is.na(gfw2013) == TRUE] <- 0
gfw2014[is.na(gfw2014) == TRUE] <- 0
gfw2015[is.na(gfw2015) == TRUE] <- 0
gfw2016[is.na(gfw2016) == TRUE] <- 0
gfw2017[is.na(gfw2017) == TRUE] <- 0
gfw2018[is.na(gfw2018) == TRUE] <- 0
gfw2019[is.na(gfw2019) == TRUE] <- 0
gfw2020[is.na(gfw2020) == TRUE] <- 0
```

```{r}
gfw_sum <- gfw2012 + gfw2013 + gfw2014 + gfw2015 + gfw2016 + gfw2017 + gfw2018 + gfw2019 + gfw2020
```

```{r}
#gfw_sum[is.na(grid_1km_original) == TRUE] <- NA
gfw_sum[gfw_sum == 0] <- NA
```

```{r}
(dest1_path_file1 <-paste(dest1_path, "/", subtheme1, "_", name1, "_", "sum_1km_", version, d_version, ".tif", sep=""))
  
writeRaster(gfw_sum, dest1_path_file1, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
```
## Log sum version

```{r}
gfw_sum <- log10(gfw_sum)
```

```{r}
(dest1_path_file1 <-paste(dest1_path, "/", subtheme1, "_", name1, "_", "log10sum_1km_", version, d_version, ".tif", sep=""))
  
writeRaster(gfw_sum, dest1_path_file1, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
```


```{r}
writeRaster(grid_1km, "./grid_1km_WGS84nodefs.tif", overwrite=TRUE)
```


