# =========================================================================================================
# Script by Edmond Sacre, SLU
# =========================================================================================================

# PREPARATIONS
```{r, include = FALSE}
# Install / load packages
#x <- c("sf", "raster", "tidyr", "dplyr")
#install.packages(x, dependencies=TRUE)
```

## Install packages
```{r, include = FALSE}
library(sp)
library(sf) 
library(raster)
library(fasterize)
library(tidyr)
library(dplyr)
library(rgdal)
library(tidyverse)
library(leaflet)
```

## View folders/themes
```{r}
folders_raw <- read.table("./shiny_data_upload/folders.txt", sep = "\t", header = TRUE)
folders_data <- read.table("./process/templates/folders_data.txt", sep = "\t", header = TRUE)
#folders <- folders_raw %>% add_row(folders_data)
```

## Define versions/directories
```{r}
version = "v01"
d_version = ".0"
location = "reg"
theme1 <- "act"
subtheme1 <- "survey"
name1 <- "seismic_survey"
(dest1_path <- paste("./data", location, theme1, subtheme1, name1, version, sep="/"))
(work1_path <- paste(dest1_path, "proc", sep="/"))
(proc1_path <- paste(dest1_path, "proc_log", sep="/"))
script_path <- "./process/r"
(script_path_windows <- paste(getwd(), "/process/r", sep=""))
(script_name <- paste(theme1, "_", subtheme1,"_", location, "_", name1,"_s01_",version, ".R", sep=""))
```

## Create directories
```{r}
dir.create(dest1_path, recursive = TRUE)
dir.create(work1_path, recursive = TRUE)
dir.create(proc1_path, recursive = TRUE)
```

## Read in WIOSym standard grid
```{r}
grid_1km <- raster("./data/reg/grid/grid/v01/grid_1km_v01.1.tif")
#grid_1km_wgs <- projectRaster(grid_1km, crs = CRS("+proj=longlat +datum=WGS84 +no_defs"))
```

## Define sources
```{r}
num_sources <- 1  # write the number of sources you will use in this script (to be used later for exporting metadata)
  
source1_dir <-  "./data_raw/glo/act/survey/mapstand/es2110251252/"
source1_id <- "es2110251252"
source1_metasym <- read.table(paste0(source1_dir, source1_id, "_metasym.txt"), sep = "\t")
source1_metasym # check that metasym file is populated, and fix if not - before finalising the processing

source_list <- c(source1_id) # for multiple sources, should look like - c(source1_id, source2_id, source3_id) etc.
```

*****************************************************************
*****************************************************************
# PROCESSING

## NOTE: Some pre-processing steps were performed in ArcGIS for efficiency - converted spatial lines to raster

# Install processing specific packages
```{r, include = FALSE}
library(spatstat)
library(spatialEco)
library(mapview)
library(stars)
library(fasterize)
library(rgdal)
library(rgeos)
library(climateStability)
```

# Import data
```{r}
data_file <- "./data/reg/act/survey/seismic_survey/v01/proc/seismic_ras.tif"
seismic <- raster(data_file)
```

# Give seismic surveys with unknown date minimum raster value
```{r}
s.temp <- seismic
s.temp[s.temp == 0] <- NA
minValue(s.temp)
seismic[seismic == 0] <- 1969
```

# Minimum year is 1969 and max year is 2019
# Scale data so that the highest values are seismic surveys done more recently
# as these would be expected to have more recent impact
```{r}
seismic <- projectRaster(seismic, grid_1km)
seismic <- resample(seismic, grid_1km)
minValue(seismic)
maxValue(seismic)
seismic <- rescale0to1(seismic)
mapview(seismic, col.regions = brewer.pal(11, "Spectral"))
```

# Convert to integer (using ceiling function) and give NA cells a value of 0
```{r}
seismic <- seismic * 100
seismic <- ceiling(seismic)
seismic[seismic == 0] <- 1
seismic[is.na(seismic)] <- 0
seismic[is.na(grid_1km)] <- NA
mapview(seismic, col.regions = brewer.pal(11, "Spectral"))
```

```{r}
product1 <- seismic
```

*****************************************************************
*****************************************************************


# EXPORT

## Specify output directory and product file name

```{r}
(dest1_path_file1 <-paste0(dest1_path, "/", subtheme1, "_", name1, "_", version, d_version, ".tif"))
```

## Export metadata
```{r}
sourcesym <- data.frame(matrix(data = NA, nrow = num_sources, ncol = 1))
colnames(sourcesym) <- "id"

for(i in 1:num_sources){
  sourcesym[i,1] <- source_list[i]
}

sourcesym
```

```{r}
sourcesym_path <- paste0(dest1_path_file1, "_sourcesym.txt")
write.table(sourcesym, sourcesym_path, row.names = F, sep = "\t")
```

## Export product
```{r}
writeRaster(product1, dest1_path_file1, COMPRESS=LZW, datatype = 'INT2S', overwrite=TRUE)
```

