# =========================================================================================================
# Script by Edmond Sacre, SLU
# This script creates rasters from the from Global Fishing Watch fishing effort data
# (https://globalfishingwatch.org/)
# This script also resamples the rasters to the same resolution and extent as the standard WIOSym grid
# =========================================================================================================

# PREPARATIONS
```{r, include = FALSE}
# Install / load packages
#x <- c("sf", "raster", "tidyr", "dplyr")
#install.packages(x, dependencies=TRUE)
```

## Install packages
```{r, include = FALSE}
library(sf) 
library(raster)
library(fasterize)
library(tidyr)
library(dplyr)
library(rgdal)
library(tidyverse)
library(leaflet)
library(mapview)
library(RColorBrewer)
```

## View folders/themes
```{r}
folders_raw <- read.table("./shiny_data_upload/folders.txt", sep = "\t", header = TRUE)
folders_data <- read.table("./process/templates/folders_data.txt", sep = "\t", header = TRUE)
#folders <- folders_raw %>% add_row(folders_data)
```

## Define versions/directories
```{r}
version = "v01"
d_version = ".0"
location = "glo"   #check
theme1 <- "act"   #check
subtheme1 <- "fishing_com"   #check
name1 <- "iotc_catch"   #check
name2 <- "" # check
(dest1_path <- paste("./data", location, theme1, subtheme1, name1, version, sep="/"))
(work1_path <- paste(dest1_path, "proc", sep="/"))
(proc1_path <- paste(dest1_path, "proc_log", sep="/"))
script_path <- "./process/r"
(script_path_windows <- paste(getwd(), "/process/r", sep=""))
(script_name <- paste(theme1, "_", subtheme1,"_", location, "_", name1,"_s01_",version, ".R", sep=""))
```

## Create directories
```{r}
dir.create(dest1_path, recursive = TRUE)
dir.create(work1_path, recursive = TRUE)
dir.create(proc1_path, recursive = TRUE)
```

## Read in WIOSym standard grid
```{r}
grid_1km <- raster("./data/reg/grid/grid/v01/grid_1km_v01.1.tif")
```

## Define sources
```{r}
num_sources <- 2  # write the number of sources you will use in this script (to be used later for exporting metadata)
  
source1_dir <-  "./data_raw/reg/act/ship/iotc/es2206171253/"    #check
source1_id <- "es2206171253"    #check
source1_metasym <- read.table(paste0(source1_dir, source1_id, "_metasym.txt"), sep = "\t")
#source1_metasym # check that metasym file is populated, and fix if not - before finalising the processing

source2_dir <-  "./data_raw/glo/act/fishing_com/gfw/es2105270900/"    #check
source2_id <- "es2105270900"    #check
source2_metasym <- read.table(paste0(source2_dir, source2_id, "_metasym.txt"), sep = "\t")
#source2_metasym # check that metasym file is populated, and fix if not - before finalising the processing

source_list <- c(source1_id) # for multiple sources, should look like - c(source1_id, source2_id, source3_id) etc.
```

*****************************************************************
*****************************************************************
# PROCESSING
```{r}
library(dplyr)
library(raster)
library(mapview)
library(RColorBrewer)
```

Tmt = metric tonnes
```{r}
#data_file <- "./data_raw/reg/act/ship/iotc/es2206171253/DATA/IOTC_TCAC_GRID_1x1.csv"
data_file <- "./data_raw/reg/act/ship/iotc/es2206171253/DATA/GRIDS.csv"
iotc <- read.csv(data_file)
data_file <- "./data_raw/reg/act/ship/iotc/es2206171253/DATA/YFT/YFT_CATCHES.csv"
bet <- read.csv(data_file)
```

```{r}
#iotc$Grid <- iotc$CODE
iotc <- dplyr::select(iotc, Grid, Long, Lat)
bet <- left_join(bet, iotc, by = "Grid")
```

```{r}
bet_sum <- bet |> group_by(Year, Fleet, Grid) |> summarize(sum_catch = sum(Tmt))
bet_sum <- bet_sum |> filter(Fleet == 'FS', Year == 2018)
bet_sum <- left_join(bet_sum, iotc, by = "Grid") %>% ungroup()
bet_sum <- dplyr::select(bet_sum, Long, Lat, sum_catch)
colnames(bet_sum) <- c("x", "y", "z")
bet_sum$y <- round(bet_sum$y/2.5)*2.5
bet_sum$x <- round(bet_sum$x/2.5)*2.5
fishing_ras <- rasterFromXYZ(bet_sum, res=c(5,5), crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
mapview(-fishing_ras, col.regions = brewer.pal(11, "Spectral"))
```

```{r}
plot(fishing_ras)
```



*****************************************************************
*****************************************************************

# EXPORT

## Specify output directory and product file name

```{r}
(dest1_path_file1 <-paste0(dest1_path, "/", subtheme1, "_", name1, "_", name2, "_", version, d_version, ".tif"))
```

## Export metadata
```{r}
sourcesym <- data.frame(matrix(data = NA, nrow = num_sources, ncol = 1))
colnames(sourcesym) <- "id"

for(i in 1:num_sources){
  sourcesym[i,1] <- source_list[i]
}

sourcesym
```

```{r}
sourcesym_path <- paste0(dest1_path_file1, "_sourcesym.txt")
write.table(sourcesym, sourcesym_path, row.names = F, sep = "\t")
```


## Export product

```{r}
writeRaster(product1, dest1_path_file1, COMPRESS=LZW, overwrite=TRUE)
```

