# =========================================================================================================
# Script by Edmond Sacre, SLU
# This script creates distribution models based on occurrence records
# This script also resamples the rasters to the same resolution and extent as the standard WIOSym grid
# =========================================================================================================

# PREPARATIONS
```{r, include = FALSE}
# Install / load packages
#x <- c("sf", "raster", "tidyr", "dplyr")
#install.packages(x, dependencies=TRUE)
```

## Install packages
```{r, include = FALSE}
library(sf) 
library(raster)
library(fasterize)
library(tidyr)
library(dplyr)
library(rgdal)
library(tidyverse)
library(leaflet)
```

## View folders/themes
```{r}
folders_raw <- read.table("./shiny_data_upload/folders.txt", sep = "\t", header = TRUE)
folders_data <- read.table("./process/templates/folders_data.txt", sep = "\t", header = TRUE)
#folders <- folders_raw %>% add_row(folders_data)
```

## Define versions/directories
```{r}
version = "v01"
d_version = ".0"
location = "reg"   #check
theme1 <- "eco"   #check
subtheme1 <- "cfsih_shark"   #check
name1 <- "shark_sdm"   #check
name2 <- "" #check
(dest1_path <- paste("./data", location, theme1, subtheme1, name1, version, sep="/"))
(work1_path <- paste(dest1_path, "proc", sep="/"))
(proc1_path <- paste(dest1_path, "proc_log", sep="/"))
script_path <- "./process/r"
(script_path_windows <- paste(getwd(), "/process/r", sep=""))
(script_name <- paste(theme1, "_", subtheme1,"_", location, "_", name1,"_s01_",version, ".R", sep=""))
(dest1_path_file1 <-paste0(dest1_path, "/", subtheme1, "_", name1, "_", version, d_version, ".tif"))
```

## Create directories
```{r}
dir.create(dest1_path, recursive = TRUE)
dir.create(work1_path, recursive = TRUE)
dir.create(proc1_path, recursive = TRUE)
```

## Read in WIOSym standard grid
```{r}
grid_1km <- raster("./data/reg/grid/grid/v01/grid_1km_v01.1.tif")
```

## Define sources
```{r}
num_sources <- 2  # write the number of sources you will use in this script (to be used later for exporting metadata)
  
source1_dir <-  "./data_raw/glo/eco/spec/obis/es2202041521/"    #check
source1_id <- "es2202041521"    #check
source1_metasym <- read.table(paste0(source1_dir, source1_id, "_metasym.txt"), sep = "\t")
#source1_metasym # check that metasym file is populated, and fix if not - before finalising the processing

source2_dir <-  "./data_raw/reg/env/ocean/copernicus/es2204071326/"   #check
source2_id <- "es2105070953"    #check
source2_metasym <- read.table(paste0(source1_dir, source1_id, "_metasym.txt"), sep = "\t")
#source2_metasym # check that metasym file is populated, and fix if not - before finalising the processing

source_list <- c(source1_id, source2_id) # for multiple sources, should look like - c(source1_id, source2_id, source3_id) etc.
```

*****************************************************************
*****************************************************************
# PROCESSING

# Install processing specific packages
```{r, include = FALSE}
library(robis)
library(spatstat)
library(spatialEco)
library(raster)
library(sf)
library(sp)
library(sdm)
library(mapview)
library(RColorBrewer)
```

# Specify taxon for which to pull sightings data from OBIS
# Taxon IDs can be found at https://www.marinespecies.org/index.php
```{r}
#IF ONLY ONE TAXON GROUP IS DESIRED
#spec <- occurrence(taxonid = 105794, geometry = "POLYGON ((86.66016 23.88584, 89.12109 -47.27923,
#                   -22.14844 -48.34165, -22.50000 21.77991, 86.66016 23.88584))")

#map_leaflet(spec)
```

```{r}
# IF MULTIPLE TAXON GROUPS ARE DESIRED
spec.list <- c("105794",
               "105843",
               "105816",
               "105926",
               "217647",
               "105817", #
               "105838",
               "105839",
               "105840",
               "105818",
               "105819"
               )

g <- "POLYGON ((86.66016 23.88584, 89.12109 -47.27923, -22.14844 -48.34165, -22.50000 21.77991, 86.66016 23.88584))"
spec <- occurrence(taxonid = spec.list[1], geometry = g)
spec <- spec[,c("scientificName", "decimalLatitude", "decimalLongitude", "country", "date_year", "category")]

for(i in 2:length(spec.list)){
  spec.temp <- occurrence(taxonid = spec.list[i], geometry = g)
  spec.temp <- spec.temp[,c("scientificName", "decimalLatitude", "decimalLongitude", "country", "date_year", "category")]
  spec <- rbind(spec, spec.temp)
}

map_leaflet(spec)
```

# Convert OBIS sightings to spatial points object
```{r}
xy <- spec[, c("decimalLongitude", "decimalLatitude")]
spec_sp <- SpatialPointsDataFrame(coords = xy, data = spec, proj4string = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))

# Downsample so that only 1 sighting per grid cell is counted (removes some sampling bias)
# temp <- raster("./process/r/predictors/mean_chl_chlorophyll_a.tif")
# spec_sp$Occurrence <- 1
# 
# spec_raster <- raster::rasterize(spec_sp, temp, field = "Occurrence", fun = mean)
# spec_sp <- rasterToPoints(spec_raster, spatial = TRUE)
# 
species <- spec_sp
```

# Import predictor layers
```{r}
lst <- list.files("./data_raw/reg/env/ocean/copernicus/es2204071326/predictors/", full.names = T)
preds <- stack(lst)
species$Occurrence <- 1
species <- species[,"Occurrence"]
e <- raster::extract(preds, species, method = "simple") # Extract predictor values to points
species <- cbind(species, e)
species <- sp.na.omit(species, margin = 1) # Remove points with no predictor data
plot(preds)
```

# If desired, decrease the resolution of the predictions to speed up processing time (e.g. for testing)
```{r}
preds <- aggregate(preds, fact = 40)
```

# Run ensemble distribution modelling
```{r}
time1 <- Sys.time()
d <- sdmData(formula=Occurrence~.,train=species, predictors=preds, bg=list(n=1000))
d
m1 <- sdm(Occurrence~.,data=d,methods=c('brt','glm','svm','gam')) 
time2 <- Sys.time()
(step1.time <- difftime(time2, time1, units = "hours"))
m1
```

```{r}
# Individual models
time1 <- Sys.time()
p1 <- predict(m1,newdata=preds,filename='./process/r/_archive/p1.tif', overwrite = T, nc = 6)
time2 <- Sys.time()
(step1.time <- difftime(time2, time1, units = "hours"))
plot(p1)
```

```{r}
# Ensemble model
time1 <- Sys.time()
e1 <- ensemble(m1,newdata=preds,filename=dest1_path_file1,setting=list(method='weighted',stat='TSS'), overwrite=T, nc=6)
time2 <- Sys.time()
(step1.time <- difftime(time2, time1, units = "hours"))
#mapview(e1, col.regions = brewer.pal(11, "Spectral"))
```

```{r}
# Other analytics of interest
#rcurve(m1, id =4)
#getVarImp(m1, id = 1)
```

```{r}
product1 <- projectRaster(e1, crs = crs(grid_1km))
product1 <- resample(product1, grid_1km)
product1[is.na(grid_1km)==TRUE] <- NA

mapview(product1, col.regions = brewer.pal(11, "Spectral"))
```

*****************************************************************
*****************************************************************

# EXPORT

## Specify output directory and product file name

```{r}
(dest1_path_file1 <-paste0(dest1_path, "/", subtheme1, "_", name1, "_", version, d_version, ".tif"))
```

## Export metadata
```{r}
sourcesym <- data.frame(matrix(data = NA, nrow = num_sources, ncol = 1))
colnames(sourcesym) <- "id"

for(i in 1:num_sources){
  sourcesym[i,1] <- source_list[i]
}

sourcesym
```

```{r}
sourcesym_path <- paste0(dest1_path_file1, "_sourcesym.txt")
write.table(sourcesym, sourcesym_path, row.names = F, sep = "\t")
```

## Export product
```{r}
writeRaster(product1, dest1_path_file1, COMPRESS=LZW, overwrite=TRUE)
```

