#=========================================================================================================
# Script by Edmond Sacre, SLU
# This script creates rasters from the from the shapefiles available on the IUCN red list of threatened species
# (https://www.iucnredlist.org/resources/spatial-data-download)
# This script also resamples the rasters to the same resolution and extent as the standard WIOSym grid
#=========================================================================================================

# PREPARATIONS
```{r, include = FALSE}
# Install / load packages
#x <- c("sf", "raster", "tidyr", "dplyr")
#install.packages(x, dependencies=TRUE)
```

## Install packages
```{r, include = FALSE}
library(sf) 
library(raster)
library(fasterize)
library(tidyr)
library(dplyr)
library(rgdal)
library(tidyverse)
library(leaflet)
```

## View folders/themes
```{r}
folders_raw <- read.table("./shiny_data_upload/folders.txt", sep = "\t", header = TRUE)
folders_data <- read.table("./process/templates/folders_data.txt", sep = "\t", header = TRUE)
#folders <- folders_raw %>% add_row(folders_data)
```

## Define versions/directories
```{r}
version = "v01"
d_version = ".0"
location = "glo"
theme1 <- "eco"
subtheme1 <- "mam"
name1 <- "threatened_mammals"
(dest1_path <- paste("./data", location, theme1, subtheme1, name1, version, sep="/"))
(work1_path <- paste(dest1_path, "proc", sep="/"))
(proc1_path <- paste(dest1_path, "proc_log", sep="/"))
script_path <- "./process/r"
(script_path_windows <- paste(getwd(), "/process/r", sep=""))
(script_name <- paste(theme1, "_", subtheme1,"_", location, "_", name1,"_s01_",version, ".R", sep=""))
```

## Create directories
```{r}
dir.create(dest1_path, recursive = TRUE)
dir.create(work1_path, recursive = TRUE)
dir.create(proc1_path, recursive = TRUE)
```

## Read in WIOSym standard grid
```{r}
grid_1km <- raster("./data/reg/grid/grid/v01/grid_1km_v01.1.tif")
```

## Define sources
```{r}
num_sources <- 1  # write the number of sources you will use in this script (to be used later for exporting metadata)
  
source1_dir <-  "./data_raw/glo/eco/mam/iucn/es2105070953/"
source1_id <- "es2105070953"
source1_metasym <- read.table(paste0(source1_dir, source1_id, "_metasym.txt"), sep = "\t")
#source1_metasym # check that metasym file is populated, and fix if not - before finalising the processing

source_list <- c(source1_id) # for multiple sources, should look like - c(source1_id, source2_id, source3_id) etc.
```

# PROCESSING
## Read in raw data
```{r}
data_file <- "./data_raw/glo/eco/mam/iucn/es2105070953/MAMMALS_MARINE_ONLY/MAMMALS_MARINE_ONLY.shp"

spec <- st_read(data_file) #%>%
  #filter(category == "CR" | category == "EN" | category == "VU" | category == "NT") # Red list category
```

## Create empty grid for rasterizing polygons
```{r}
spec_grid <- raster(spec, res = 1/100)
```

## Convert the IUCN polygons into a raster, and calculate the total number of species in each cell
```{r}
spec_raster <- fasterize(spec, spec_grid, fun = "sum")
```

## Crop the raster to approximately the WIOSym extent (to speed up projecting)
```{r}
ex <- extent(1, 85,-45, 25)
spec_raster <- crop(spec_raster, ex)
```

## Project IUCN raster to same projection as WIOSym grid
```{r}
spec_raster <- projectRaster(spec_raster, crs = crs(grid_1km))
```

## Resample to WIOSym grid
```{r}
spec_raster <- resample(spec_raster, grid_1km, method = "ngb")

spec_raster[is.na(grid_1km) == TRUE] <- NA
```

## View raster
```{r}
plot(spec_raster)
```

##  View raster interactively
```{r, echo = FALSE, warning = FALSE, error = FALSE, results = "hide", fig.keep = "all"}
pal <- colorNumeric(c("#3288bd", "#ffffbf", "#d53e4f"), values(spec_raster), na.color = "transparent")

leaflet() %>% addRasterImage(spec_raster, layerId = "values", colors = pal)
```


# EXPORT
## Specify output directory and product file name
```{r}
(dest1_path_file1 <-paste0(dest1_path, "/", subtheme1, "_", name1, "_richness_1km", "_", version, d_version, ".tif"))
```

## Export metadata
```{r}
sourcesym <- data.frame(matrix(data = NA, nrow = num_sources, ncol = 1))
colnames(sourcesym) <- "id"

for(i in 1:num_sources){
  sourcesym[i,1] <- source_list[i]
}

sourcesym
```

```{r}
sourcesym_path <- paste0(dest1_path_file1, "_sourcesym.txt")
write.table(sourcesym, sourcesym_path, row.names = F, sep = "\t")
```


## Export product
```{r}
writeRaster(spec_raster, dest1_path_file1, COMPRESS=LZW, datatype = 'INT4S', overwrite=TRUE)
```

