# =========================================================================================================
# Script by Edmond Sacre, SLU
# This script creates distribution models based on occurrence records
# This script also resamples the rasters to the same resolution and extent as the standard WIOSym grid
# =========================================================================================================

# PREPARATIONS
```{r, include = FALSE}
# Install / load packages
#x <- c("sf", "raster", "tidyr", "dplyr")
#install.packages(x, dependencies=TRUE)
```

## Install packages
```{r, include = FALSE}
library(sf) 
library(raster)
library(fasterize)
library(tidyr)
library(dplyr)
library(rgdal)
library(tidyverse)
library(leaflet)
```

## View folders/themes
```{r}
folders_raw <- read.table("./shiny_data_upload/folders.txt", sep = "\t", header = TRUE)
folders_data <- read.table("./process/templates/folders_data.txt", sep = "\t", header = TRUE)
#folders <- folders_raw %>% add_row(folders_data)
```

## Define versions/directories
```{r}
version = "v01"
d_version = ".0"
location = "reg"   #check
theme1 <- "eco"   #check
subtheme1 <- "cfsih_shark"   #check
name1 <- "shark_sdm"   #check
name2 <- "allspecies" #check
(dest1_path <- paste("./data", location, theme1, subtheme1, name1, version, sep="/"))
(work1_path <- paste(dest1_path, "proc", sep="/"))
(proc1_path <- paste(dest1_path, "proc_log", sep="/"))
script_path <- "./process/r"
(script_path_windows <- paste(getwd(), "/process/r", sep=""))
(script_name <- paste(theme1, "_", subtheme1,"_", location, "_", name1,"_s01_",version, ".R", sep=""))
(dest1_path_file1 <-paste0(dest1_path, "/", subtheme1, "_", name1, "_", version, d_version, ".tif"))
```

## Create directories
```{r}
dir.create(dest1_path, recursive = TRUE)
dir.create(work1_path, recursive = TRUE)
dir.create(proc1_path, recursive = TRUE)
```

## Read in WIOSym standard grid
```{r}
grid_1km <- raster("./data/reg/grid/grid/v01/grid_1km_v01.1.tif")
```

## Define sources
```{r}
num_sources <- 2  # write the number of sources you will use in this script (to be used later for exporting metadata)
  
source1_dir <-  "./data_raw/glo/eco/spec/obis/es2202041521/"    #check
source1_id <- "es2202041521"    #check
source1_metasym <- read.table(paste0(source1_dir, source1_id, "_metasym.txt"), sep = "\t")
#source1_metasym # check that metasym file is populated, and fix if not - before finalising the processing

source2_dir <-  "./data_raw/reg/env/ocean/copernicus/es2204071326/"   #check
source2_id <- "es2105070953"    #check
source2_metasym <- read.table(paste0(source1_dir, source1_id, "_metasym.txt"), sep = "\t")
#source2_metasym # check that metasym file is populated, and fix if not - before finalising the processing



source_list <- c(source1_id, source2_id) # for multiple sources, should look like - c(source1_id, source2_id, source3_id) etc.
```

*****************************************************************
*****************************************************************
# PROCESSING

# Install processing specific packages
```{r, include = FALSE}
library(robis)
library(spatstat)
library(spatialEco)
library(raster)
library(sf)
library(sp)
library(sdm)
library(mapview)
library(RColorBrewer)
```

# Specify taxon for which to pull sightings data from OBIS
# Taxon IDs can be found at https://www.marinespecies.org/index.php
# Selachii = 368408
```{r}
# IF ONLY ONE TAXON GROUP IS DESIRED
spec <- occurrence(#taxonid = 368408,
                   scientificname = "Centrophorus squamosus",
                   geometry = "POLYGON ((86.66016 23.88584, 89.12109 -47.27923, -22.14844 -48.34165, -22.50000 21.77991, 86.66016 23.88584))",
                   startdate = as.Date("1970-01-01"))

#spec <- filter(spec, date_year > 2010) #filter by year if desired
#spec <- filter(spec, category == "CR") #filter by red list status
#spec <- filter(spec, category == "CR" | category == "EN") #filter by red list status
#spec <- filter(spec, category == "CR" | category == "EN" | category == "VU") #filter by red list status

map_leaflet(spec)
```

```{r}
levels(as.factor(spec$scientificName))
```

# Convert OBIS sightings to spatial points object
```{r}
xy <- spec[, c("decimalLongitude", "decimalLatitude")]
spec_sp <- SpatialPointsDataFrame(coords = xy, data = spec, proj4string = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))

# Downsample so that only 1 sighting per grid cell is counted (removes some sampling bias)
# temp <- grid_1km
# spec_sp$Occurrence <- 1
# spec_raster <- raster::rasterize(spec_sp, temp, field = "Occurrence", fun = mean)
# spec_sp <- rasterToPoints(spec_raster, spatial = TRUE)
species <- spec_sp
```

# Import predictor layers
```{r}
lst <- list.files("./data/reg/util/sdm_predictors/", full.names = T)
preds <- stack(lst)
species$Occurrence <- 1
species <- species[,"Occurrence"]
e <- raster::extract(preds, species, method = "simple") # Extract predictor values to points
species <- cbind(species, e)
species <- sp.na.omit(species, margin = 1) # Remove points with no predictor data
plot(preds)
```

# If desired, decrease the resolution of the predictions to speed up processing time (e.g. for testing)
```{r}
preds <- aggregate(preds, fact = 10)
```

# Run ensemble distribution modelling
```{r}
d <- sdmData(formula=Occurrence~.,train=species, predictors=preds, bg=list(n=1000))
d
m1 <- sdm(Occurrence~.,data=d,methods=c('brt','svm')) 
m1
write.sdm(d, filename = paste0(work1_path,"/", name1, "_sdmdata"), overwrite = T)
write.sdm(m1, filename = paste0(work1_path,"/", name1, "_sdmmodel"), overwrite = T)
```

```{r}
# Individual models
p1 <- predict(m1,newdata=preds,filename='./process/r/_archive/p1.tif', overwrite = T, nc = 6) 
plot(p1)
```

```{r}
# Ensemble model
e1 <- ensemble(m1,newdata=preds,filename='e1.tif',setting=list(method='weighted',stat='TSS'), overwrite=T, nc=6)
#mapview(e1, col.regions = brewer.pal(11, "Spectral"))
```

```{r}
# Other analytics of interest
#rcurve(m1, id =4)
#getVarImp(m1, id = 1)
```

```{r}
#prod <- p1[[1]]#select specific model for product

prod <- e1
product1 <- projectRaster(prod, crs = crs(grid_1km))
product1 <- resample(product1, grid_1km)
product1[is.na(grid_1km)==TRUE] <- NA

mapview(product1, col.regions = brewer.pal(11, "Spectral"))
```

```{r}
# Transformations and adjustments
# test <- product1
# test[test < 0.8] <- NA
# mapview(test, col.regions = brewer.pal(11, "Spectral"))
```

# Combine sdms for different species to get a measure of species richness
```{r}
#base <- grid_1km - 1
eco1 <- raster("./data/reg/eco/cfsih_shark/shark_sdm/v01/cfsih_shark_shark_sdm_Carcharhinus_obscurus_v01.0.tif")
eco2 <- raster("./data/reg/eco/cfsih_shark/shark_sdm/v01/cfsih_shark_shark_sdm_Carcharhinus_signatus_v01.0.tif")
eco3 <- raster("./data/reg/eco/cfsih_shark/shark_sdm/v01/cfsih_shark_shark_sdm_Centrophorus_granulosus_v01.0.tif")
eco4 <- raster("./data/reg/eco/cfsih_shark/shark_sdm/v01/cfsih_shark_shark_sdm_Centrophorus_squamosus_v01.0.tif")
eco5 <- raster("./data/reg/eco/cfsih_shark/shark_sdm/v01/cfsih_shark_shark_sdm_Isurus_oxyrinchus_v01.0.tif")
eco6 <- raster("./data/reg/eco/cfsih_shark/shark_sdm/v01/cfsih_shark_shark_sdm_Sphyrna_zygaena_v01.0.tif")
eco7 <- raster("./data/reg/eco/cfsih_shark/shark_sdm/v01/cfsih_shark_shark_sdm_Squalus_acanthias_v01.0.tif")


#eco1[eco1<0.7] <- 0
#eco2[eco2<0.7] <- 0
#eco3[eco3<0.7] <- 0

eco <- eco1+eco2+eco3+eco4+eco5+eco6+eco7
product1 <- eco

mapview(eco, col.regions = brewer.pal(11, "Spectral"))
```

*****************************************************************
*****************************************************************

# EXPORT

## Specify output directory and product file name

```{r}
#(dest1_path_file1 <- paste0(dest1_path, "/", subtheme1, "_", name1, "_", name2, "_", version, d_version, ".tif"))
(dest1_path_file1 <-paste0(dest1_path, "/", subtheme1, "_", name1, "_", version, d_version, ".tif"))
```

## Export metadata
```{r}
sourcesym <- data.frame(matrix(data = NA, nrow = num_sources, ncol = 1))
colnames(sourcesym) <- "id"

for(i in 1:num_sources){
  sourcesym[i,1] <- source_list[i]
}

sourcesym
```

```{r}
sourcesym_path <- paste0(dest1_path_file1, "_sourcesym.txt")
write.table(sourcesym, sourcesym_path, row.names = F, sep = "\t")
```

## Export product

```{r}
writeRaster(product1, dest1_path_file1, COMPRESS=LZW, overwrite=TRUE)
```

