# =========================================================================================================
# Script by Edmond Sacre, SLU
# This script creates a map of marine heatwaves as defined by NOAA, which categorizes marine heatwaves
# into categories 1 through 5. To produce this layer, we used data from the years 1985 through 2023.
# A heatwave is defined as exceeding the 90th percentile of values in a give grid cell. The classification
# is then assigned based on the degree to which the temperature exceed this threshold. We build upon this 
# by identifying the duration of the MHW (the average number of days above the threshold), the intensity
# (the average category) and the frequency (the number of days per year above this threshold).
# =========================================================================================================

# PREPARATIONS
```{r, include = FALSE}
# Install / load packages
#x <- c("sf", "raster", "tidyr", "dplyr")
#install.packages(x, dependencies=TRUE)
```

## Install packages
```{r, include = FALSE}
library(tidyr)
library(dplyr)
library(terra)
```

## View folders/themes
```{r}
#folders_raw <- read.table("./shiny_data_upload/folders.txt", sep = "\t", header = TRUE)
#folders_data <- read.table("./process/templates/folders_data.txt", sep = "\t", header = TRUE)
#folders <- folders_raw %>% add_row(folders_data)
```

## Define versions/directories
```{r}
version = "v01"
d_version = ".0"
location = "reg"   #check
theme1 <- "pres"   #check
subtheme1 <- "cc_temp"   #check
name1 <- "temp_rise_mhw"   #check
name2 <- "" #check
(dest1_path <- paste("./data", location, theme1, subtheme1, name1, version, sep="/"))
(work1_path <- paste(dest1_path, "proc", sep="/"))
(proc1_path <- paste(dest1_path, "proc_log", sep="/"))
script_path <- "./process/r"
(script_path_windows <- paste(getwd(), "/process/r", sep=""))
(script_name <- paste(theme1, "_", subtheme1,"_", location, "_", name1,"_s01_",version, ".Rmd", sep=""))
(dest1_path_file1 <-paste0(dest1_path, "/", subtheme1, "_", name1, "_", version, d_version, ".tif"))
```

## Create directories
```{r}
dir.create(dest1_path, recursive = TRUE)
dir.create(work1_path, recursive = TRUE)
dir.create(proc1_path, recursive = TRUE)
```

## Read in WIOSym standard grid
```{r}
grid_1km <- rast("./data/reg/grid/grid/v01/grid_1km_v01.1.tif")
grid_proj <- rast("./products/grid/grid_1km_v01.1_epsg4326.tif")
```

## Define sources
```{r}
num_sources <- 1  # write the number of sources you will use in this script (to be used later for exporting metadata)
  
source1_dir <-  "./data_raw/glo/pres/cc_temp/noaa/es2405150813488330/"    #check
source1_id <- "es2405150813488330"    #check
source1_metasym <- read.table(paste0(source1_dir, source1_id, "_metasym.txt"), sep = "\t")
#source1_metasym # check that metasym file is populated, and fix if not - before finalising the processing

source_list <- c(source1_id) # for multiple sources, should look like - c(source1_id, source2_id, source3_id) etc.
```

*****************************************************************
*****************************************************************
# PROCESSING

# Install processing specific packages
```{r, include = FALSE}
library(terra)
library(ncdf4)
```

# You will first need to run the "mhw_download" script to download all of the NetCDF files
# WARNING: This will take a long time to run
# Mean and sum MHW category per year
```{r}
for(j in 1985:2023){
  
  yr <- j
  l <- list.files(paste0(work1_path, "/mhw_noaa/", yr), full.names = T)
  #l <- l[1:10]
  nit <- 0
  
  for(i in 1:length(l)){
  
    fp <- l[i] 
    #r <- rast(fp)["heatwave_category"]
    n <- nc_open(fp)
    lon <- ncdf4::ncvar_get(n, n$var[["lon"]])
    lat <- ncdf4::ncvar_get(n, n$var[["lat"]])
    d <- ncdf4::ncvar_get(n, n$var[["heatwave_category"]])
    nc_close(n)
    
    r <- rast(t(d[, ncol(d):1]), ext=c(range(lon), range(lat)), crs="+proj=lonlat")
    r[r == -1] <- 0
    #r[is.na(r)] <- 0
    
    if(nit == 0){mhw <- r}
    else{mhw <- mhw + r}
    nit <- nit+1
    #cat("-", nit)
  }
  
  mhw_mean <- mhw/length(l) # Calculate average MHW category
  mhw_sum <- mhw
  writeRaster(mhw_mean, paste0(work1_path, "/mhw_mean/mhw_mean_cat_", yr, ".tif"), overwrite = T)
  writeRaster(mhw_sum, paste0(work1_path, "/mhw_sum/mhw_sum_cat_", yr, ".tif"), overwrite = T)
  plot(mhw_sum)
  cat("-", j)
  
}
```

# Number of days each year with heatwave, i.e. above 90th percentile
```{r}
for(j in 1985:2023){
  yr <- j
  l <- list.files(paste0(work1_path, "/mhw_noaa/", yr), full.names = T)
  #l <- l[1:10]
  nit <- 0
  
  for(i in 1:length(l)){
  
    fp <- l[i] 
    #r <- rast(fp)["heatwave_category"]
    n <- nc_open(fp)
    lon <- ncdf4::ncvar_get(n, n$var[["lon"]])
    lat <- ncdf4::ncvar_get(n, n$var[["lat"]])
    d <- ncdf4::ncvar_get(n, n$var[["heatwave_category"]])
    nc_close(n)
    
    r <- rast(t(d[, ncol(d):1]), ext=c(range(lon), range(lat)), crs="+proj=lonlat")
    r[r == -1] <- 0
    r[r > 1] <- 1 # Convert all above 90th percentile to 1
    
    if(nit == 0){mhw <- r}
    else{mhw <- mhw + r}
    nit <- nit+1
  }
  
  writeRaster(mhw, paste0(work1_path, "/mhw_days/mhw_days_cat_", yr, ".tif"), overwrite = T)
  #plot(mhw)
  cat("-", j)

}
```

# No. of days above category 3
```{r}
for(j in 1985:2023){
  yr <- j
  l <- list.files(paste0(work1_path, "/mhw_noaa/", yr), full.names = T)
  #l <- l[1:10]
  nit <- 0
  
  for(i in 1:length(l)){
  
    fp <- l[i] 
    #r <- rast(fp)["heatwave_category"]
    n <- nc_open(fp)
    lon <- ncdf4::ncvar_get(n, n$var[["lon"]])
    lat <- ncdf4::ncvar_get(n, n$var[["lat"]])
    d <- ncdf4::ncvar_get(n, n$var[["heatwave_category"]])
    nc_close(n)
    
    r <- rast(t(d[, ncol(d):1]), ext=c(range(lon), range(lat)), crs="+proj=lonlat")
    r[r == -1] <- 0
    r[r < 3] <- 0 # Convert all above in cat 3 and above to 1
    r[r >= 3] <- 1
    
    if(nit == 0){mhw <- r}
    else{mhw <- mhw + r}
    nit <- nit+1
  }
  
  writeRaster(mhw, paste0(work1_path, "/mhw_days_cat3plus/mhw_days_cat3plus_", yr, ".tif"), overwrite = T)
  #plot(mhw)
  cat("-", j)

}
```

# MHW duration - this will take quite a long time - approx 10 mins per year
```{r}
dur <- function(x){
  x[is.na(x)] <- 0
  x[x > 1] <- 1
  x <- rle(x)
  if(length(x$lengths[x$values == 1]) == 0){m <- 0}
  else(m <- mean(x$lengths[x$values == 1]))
  return(m)
}

for(j in 1985:2023){

  yr <- j
  l <- list.files(paste0(work1_path, "/mhw_noaa/", yr), full.names = T)
  #l <- l[1:20]
  nit <- 0
  
  for(i in 1:length(l)){
  
    fp <- l[i] 
    #r <- rast(fp)["heatwave_category"]
    n <- nc_open(fp)
    lon <- ncdf4::ncvar_get(n, n$var[["lon"]])
    lat <- ncdf4::ncvar_get(n, n$var[["lat"]])
    d <- ncdf4::ncvar_get(n, n$var[["heatwave_category"]])
    nc_close(n)
    
    r <- rast(t(d[, ncol(d):1]), ext=c(range(lon), range(lat)), crs="+proj=lonlat")
    r[r == -1] <- 0
    r <- crop(r, ext(8, 78, -40, 16))
    #r <- crop(r, ext(45, 50, -15, -10))
    
    n <- basename(fp)
    st <- nchar(n)-6
    sp <- nchar(n)-3
    n <- substr(n, st, sp)
    names(r) <- n
    
    if(nit == 0){mhw <- r}
    else{mhw <- c(mhw, r)}
    nit <- nit+1
  }
  
  base <- (mhw[[1]]*0)+1
  temp <- app(mhw, fun = dur, cores = 6)
  mhw_dur <- temp*base
  writeRaster(mhw_dur, paste0(work1_path, "/mhw_duration/mhw_mean_duration_", yr, ".tif"), overwrite = T)
  cat("-", j)

}
```

# MHW final map of intensity*duration*frequency
# Here we take the mean across all years from 1985 to 2023
```{r}
# MHW frequency
l <- list.files(paste0(work1_path, "/mhw_days/"), full.names = T)
r <- rast(l)
r <- app(r, fun = 'mean')
r <- crop(r, ext(8, 78, -40, 16))
r <- terra::project(r, grid_1km, method = "cubic")
mhw_days <- r

# MHW intensity
l <- list.files(paste0(work1_path, "/mhw_mean/"), full.names = T)
r <- rast(l)
r <- app(r, fun = 'mean')
r <- crop(r, ext(8, 78, -40, 16))
r <- terra::project(r, grid_1km, method = "cubic")
mhw_intens <- r

# MHW duration
l <- list.files(paste0(work1_path, "/mhw_duration/"), full.names = T)
r <- rast(l)
r <- app(r, fun = 'mean')
r <- crop(r, ext(8, 78, -40, 16))
r <- terra::project(r, grid_1km, method = "cubic")
mhw_dur <- r

mhw <- mhw_days*mhw_intens*mhw_dur
#q <- quantile(terra::values(mhw, na.rm = T), probs = 0.50) # Add quantile cutoff? Default to no cutoff
#mhw[mhw < q] <- 0
plot(mhw)
```

```{r}
product1 <- mhw
```

*****************************************************************
*****************************************************************

# EXPORT

## Specify output directory and product file name
```{r}
(dest1_path_file1 <-paste0(dest1_path, "/", subtheme1, "_", name1, "_", version, d_version, ".tif"))
```

## Export image of product raster(s)
```{r}
(img_path <- paste0(dest1_path, "/", subtheme1, "_", name1, "_cutoff_", version, d_version, ".png"))
grid_1km_raster <- raster::raster(grid_1km)
prod1_scale <- raster::raster(product1)
prod1_scale[prod1_scale == 0] <- NA
prod1_scale[] <- scales::rescale(prod1_scale[], to = c(1, 100))
prod1_scale[is.na(prod1_scale)] <- 0
prod1_scale[is.na(grid_1km_raster)] <- NA


png(img_path, width = ncol(prod1_scale), height = nrow(prod1_scale))
pal <- colorRampPalette(rev(c('#d53e4f','#f46d43','#fdae61','#fee08b','#ffffbf','#e6f598','#abdda4','#66c2a5','#3288bd')))
plot(prod1_scale, maxpixels = ncell(prod1_scale), main = "", axes = FALSE, ylim=raster::extent(prod1_scale)[3:4], 
     breaks = seq(0,100,1), col = pal(101), legend = F, box = F, colNA = "gray92")
dev.off()
```

## Export metadata
```{r}
sourcesym <- data.frame(matrix(data = NA, nrow = num_sources, ncol = 1))
colnames(sourcesym) <- "id"

for(i in 1:num_sources){
  sourcesym[i,1] <- source_list[i]
}

sourcesym
```

```{r}
sourcesym_path <- paste0(dest1_path_file1, "_sourcesym.txt")
write.table(sourcesym, sourcesym_path, row.names = F, sep = "\t")
```

## Export product
```{r}
writeRaster(product1, dest1_path_file1, gdal = c("COMPRESS=LZW"), overwrite=TRUE)
```

