# ========================================================================================================= #
# IMPORTANT - To get started:
Because this script is an R Markdown script, the default directory is the location of this .Rmd file.
You must open this script from within the .Rproj file associated with your wiosym database for it to work.
If you haven't used R notebooks before (.Rmd), each of the code "chunks" can be executed by clicking the green play
button at the top right of the chunk.

To run the script from the project directory, in R Studio, you need to go to: Tools > Global Options > R Markdown
and set "Evaluate chunks in directory" to "Project", though for latest version of R/Rstudio it may not be needed...
# ========================================================================================================= #
# ABOUT
# ========================================================================================================= #
## Brief description: 
AIS / Shipping and noise pressure to standard grid
### Scripts/processes that needs to be updated prior?: 
###
### Script by: 
gk, sgu, Rver 4.1.1.
### Updated: 
gk, sgu, Rver 4.1.1. New AIS data from HUB Ocean 2022-09
### Developer check: Initial, org, yymmdd, comments
### External check: Initial, org, yymmdd, comments
# ========================================================================================================= # 
# PREPARATIONS
# ========================================================================================================= #

### Install packages
```{r, include = FALSE}
# list all packages that this script is dependent on, remove any redundant ones
x <- c("sf", "raster", "rgdal", "tidyverse", "gdalUtils")
#install.packages("gdalUtils", dependencies=TRUE)
```

### Load packages
```{r, include = FALSE}
library(gdalUtils)
library(sf) 
library(raster)
library(rgdal)
library(tidyverse)
library(readxl)
```

## Set version
```{r}
version = "v01" # relates to major iteration of WIOSym (v01 for grid 2021, v02 for grid 2022)
d_version = ".2"  # detailed version (you can add another level e.g. 0.0  -> v01.0.0)
seq = "s01" # sequential order - change accordingly if process is run in several scripts
```

## Set geographic scope
```{r}
read_tsv("./shiny_data_upload/locations.txt")  # prints available locations to choose from and their abbreviations
```

```{r}
location = "reg" # choose applicable 3 letter abbreviation from "location_val" -  reg for whole WIO area
```

## Set main theme
```{r}
folders_raw <- read_tsv("./shiny_data_upload/folders.txt")
folders_data <- read_tsv("./shiny_data_upload/folders_data.txt") # additional folders for data not applicable for data_raw 
folders <- folders_raw %>% add_row(folders_data)
(unique(folders["theme_folder"]))
```

```{r}
theme <- "act" # Copy from "theme_folder" (e.g. "eco")
```

## Set Subtheme
```{r}
folders %>%
  filter(theme_folder == theme) %>% 
  dplyr::select(subtheme_folder, subtheme_name) %>% 
  print(n = Inf)
```

```{r}
subtheme <- "ship" # this script will adress many pressurs bot working from shipping activity dir
```

## Set paths
```{r}
(dest_path <- paste("./data", location, theme, subtheme, version, "", sep="/")) # path to final product
(work_path <- paste(dest_path, "proc/", sep=""))
(proc_path <- paste(dest_path, "proc_log/", sep=""))
(archive_path <- paste(dest_path, "_archive/", sep=""))
(script_path <- "./process/r/")
```

## Create directories
```{r}
dir.create(dest_path, recursive = TRUE)
dir.create(work_path, recursive = TRUE)
dir.create(proc_path, recursive = TRUE)
dir.create(archive_path, recursive = TRUE)
dir.create(script_path, recursive = TRUE)
```

## Save R Script
```{r}
name1 <- "" # "REPLACE_" if script need additional name... don't use unless needed to avoid duplication.
(script_path_windows <- paste(getwd(), "/process/r", sep=""))
(script_name <- paste(theme, "_", subtheme,"_", location, "_", name1, seq, "_", version, ".Rmd", sep=""))
```
 # copy path and name and use File/Save As in R studio to save rmd file
 
#==========================================================================================================================#
# INDATA
#==========================================================================================================================#
# Set DATA_RAW sources
IMPORTANT all data harvested from data_raw needs to have a metasym.txt file in its root directory (not file specific)
## DIR1: 
```{r}
source1_dir <- source_dir <-  "./data_raw/reg/act/ship/hub/gk2209241037/"  # REPLACE - internal path
source1 <- source_id <- "gk2209241037" # 
(read_tsv(paste(source_dir, source_id, "_metasym.txt", sep=""))) # check the metasym file
``` 

```{r}
dir(source_dir, recursive=T, pattern= ".tiff$")
```

```{r}
# Files
ship_stack_names <- list.files(source_dir, pattern = ".tiff$", recursive=T)
ship_stack <- stack(paste(source_dir, list.files(source_dir, pattern = ".tiff$", recursive=T), sep=''))


# addressing problems with stack... needs flipping and projection string
ship_stack <- flip(ship_stack, direction='y')
proj4string(ship_stack) <- CRS("+init=epsg:4326")

(names(ship_stack))

#writeRaster(ship_stack[[5]], paste(work_path, "test.tif", sep=""), overwrite=T, datatype = 'FLT4S', COMPRESS=LZW)
```

### declare number of "data_raw" directorys above
```{r}
data_raw_metasym_num <- 1 # set total number of data_raw dir used (e.g. <- 1), to help in export section
```


# DATA sources
IMPORTANT Products in data are based on information in data_raw. Check that *_sourcesym.txt exists for each file to track acccumulated IDs. 

## DATA1: Grid 1km - loading all flavours (watermask 1, 0, NA) at once since identical sourcesym apply
```{r}
data_dir <-  "./data/reg/grid/grid/v01/"  # input your data directory
dir(data_dir) # check what data is in this directory

data1_file <- "grid_1km_v01.1.tif" 
data1_file2 <- "grid_1km_0_v01.1.tif" 
data1_file3 <- "grid_1km_na_v01.1.tif" 

(grid_1km_path <- path <- paste(data_dir, data1_file, sep=""))
(grid_1km <- raster(path)) # read and check your data
(grid_1km_0_path <- path <- paste(data_dir, data1_file2, sep=""))
(grid_1km_0 <- path <- raster(path)) # read and check your data
(grid_1km_na_path <- path <- paste(data_dir, data1_file3, sep=""))
(grid_1km_na <- path <- raster(path)) # read and check your data

data1_sourcesym <- read_tsv(paste(data_dir, data1_file, "_sourcesym.txt", sep=""))
# check that source IDs exists and make sense

plot(grid_1km)
```

## DATA2: Grid 250m - loading all flavours (watermask 1, 0, NA) at once since same sourcesym apply
```{r}
data2_dir <- data_dir <-  "./data/reg/grid/grid/v01/"  # input your data directory

data2_file <- "grid_250m_v01.1.tif" # your data file
data2_file2 <- "grid_250m_0_v01.1.tif" # your data file
data2_file3 <- "grid_250m_na_v01.1.tif" # your data file

(grid_250m_path <- path <- paste(data_dir, data2_file, sep=""))
(grid_250m <- raster(path)) # read and check your data
(grid_250m_0_path <- path <- paste(data_dir, data2_file2, sep=""))
(grid_250m_0 <- path <- raster(path)) # read and check your data
(grid_250m_na_path <- path <- paste(data_dir, data2_file3, sep=""))
(grid_250m_na <- path <- raster(path)) # read and check your data

(data2_sourcesym <- read_tsv(paste(data_dir, data2_file, "_sourcesym.txt", sep="")))
# check so source IDs exists and make sense
# check that source IDs exists and make sense
#plot(grid_250m)
```

# data3: bounding box for the grid (created by SGU in GIS so no data_raw sources involved)
data3_dir <- "./data/reg/grid/scope/v01/" 
dir(data2_dir)
data3_file <- "wiosym_data_bounding_box_v01.shp" 
data3_sourcesym <- read_tsv(paste(data2_dir, data2_file, "_sourcesym.txt", sep=""))
data3_sourcesym # check so source IDs exists and make sense
data3_path <- paste(data3_dir, data3_file, sep="")
data3_path # note: this data is not read in as an object since the process in this script happens in GDAL outside R

extent_box <- st_read(data3_path)




## declare number of "data" files
```{r}
data_sourcesym_num <-  2 # set total number of data files (e.g. <- 2), to help in export section
```

#==========================================================================================================================#
# PROCESSING
#==========================================================================================================================#



## create mask for harbor areas, and interpolate stack
```{r}
stack <- ship_stack
names(stack)
r <- stack[[5]] # total shipping
summary(r) 

# reclassify to port mask - use total hours to identify potential anchor areas (e.g. most ships at 0 velocity). approximate hours for these areas are >1500 / 2 years. observed using GIS...


m <-c(-10,1500,NA, 1500,100000000000000,1)
m
rc <- reclassify(r, m)
plot(rc)
rc
rc_poly = rasterToPolygons(rc, na.rm=TRUE, dissolve=T)
rc_poly_sf <- st_as_sf(rc_poly)
rc_poly_sf = st_cast(rc_poly_sf,"POLYGON") # explode into singlepart features

path1 <- paste(work_path, "portmask_uned.shp", sep="")
path2 <- paste(work_path, "portmask.shp", sep="")


st_write(rc_poly_sf, path1, driver="ESRI Shapefile", delete_layer=T)

# MANUAL EDIT NEEDED - edit of shapefile was conducted in GIS to remove shipping areas from mask. then reinport the polygon, resaved as portmask.shp

mask <- rc_poly_edit_sf <- st_read(path2)
plot(mask)


# mask and interpolate stack

stack_interp <- stack_crop <- mask(stack, mask, inverse=T)


for(i in 1:length(names(stack_crop)))
{
  
r <- stack_crop[[i]]

# focal filter to extrapolate and fill whole grid
r1 <- focal(r, w=matrix(1,3,3), fun=mean, na.rm=T, pad=T, NAonly=T) 
r2 <- focal(r1, w=matrix(1,3,3), fun=mean,na.rm=T, pad=T, NAonly=T) 
r3 <- focal(r2, w=matrix(1,5,5), fun=mean,na.rm=T, pad=T, NAonly=T) 
r4 <- focal(r3, w=matrix(1,7,7), fun=mean,na.rm=T, pad=T, NAonly=T) 
r5 <- focal(r4, w=matrix(1,9,9), fun=mean,na.rm=T, pad=T, NAonly=T) 

r5[is.na(r5[])] <- 0 

stack_interp[[i]] <- r5

}

names(stack_interp) <- names(stack_crop)
(stack_interp)

```


# NOISE POLLUTION  - AIS into noise disturbane
Resources
https://www.frontiersin.org/articles/10.3389/fmars.2019.00606/full#supplementary-material
https://www.emsa.europa.eu/protecting-the-marine-environment/underwater-noise.html  (Sounds_report.pdf, Figure 16 - Different vessel types and overlapping of hearing range of different groups of species.)

http://resource.npl.co.uk/acoustics/techguides/seaabsorption/

Container ships
absorbtion at 31Hz, 20 C, 100m depth is approx 5 db/km, peak-level ~180db (container ship) resulting impact zone to 90db db ~ 18 km

Tankers
absorbtion at 80Hz, 20 C, 100m depth is approx 20 db/km, peak-level ~180db (coastal tanker) resulting impact zone  to 90db db ~ 5 km

Fishing/other

absorbtion at 500Hz, 20 C, 100m depth is approx 145 db/km, peak 120db? (fishing boat/jetski) - resulting impact zone ~ 1km

sailing boat
absorbtion at 2500Hz, 20 C, 100m depth is approx 1500 db/km, peak 110db? (small leasure/motorboat) resulting impact zone ~100m


https://www.antennas.ca/calc_db_watts.htm
180 10^12 kw
170 10^11
160 10^10 kw
150 10^9 kw
140 10^8 kw
130 10^7 kw
120 10^6 kw
110 10^5 kw
100 10^4 kw
90 10^3 kw

Additional input: dB of different shiptypes, approximate from Dr.  Hammar, shipping expert in sweden
f_high170dB mainly commersal traffic (assuming oil vessels are the same..)
f_med  160 dB
f_low  150 dB


## model noise propagation and calculate watt hours for all ship types in AIS data
```{r}
stack <- stack_interp

# importing weights - only cargo and tanker needs focal statistics/weights since spreading is limited on other vesseltypes
dir(work_path)


# Created weight matrix in excel based on above assumptions and importing below. 
# weight low frequenzy data to dB loss in water

w_cargo <- as.matrix(read_excel(paste(work_path, "weightmatrix_db.xlsx", sep=""), 2, col_names=F))

w_tanker <- as.matrix(read_excel(paste(work_path, "weightmatrix_db.xlsx", sep=""), 3, col_names=F))

names(stack)


# calculate cargo Watt x hour

r <- stack[[1]] # cargo
r_cargo <- focal(r, w=w_cargo, fun=sum, na.rm=T, pad=T, NAonly=F) 

r_dif <- r_cargo -r
plot(r_dif)

r_cargo_wh <- r_cargo * 100000000000000 * 0.4223 # w hours at 170db, normalised for average watt in 2.5km cell

# calculate "tanker" Watt x hour

r <- stack[[4]]  # tanker
r_tanker <- focal(r, w=w_tanker, fun=sum, na.rm=T, pad=T, NAonly=F) 
r_dif <- r_tanker -r
plot(r_dif)

r_tanker_wh <- r_tanker * 100000000000000 * 0.0784 # w hours at 170db normalized for average watt in 2.5km cell

# calculate passanger Watt x hour

r <- stack[[3]] # passenger
r_pass_wh <- r * 100000000000 * 0.04# w hours at 140db normalized for average watt in 2.5km cell


# calculate fishing Watt x hour
r <- stack[[2]] # fishing
r_fish_wh <- r * 100000000000 * 0.04# w hours at 140db normalized for average watt in 2.5km cell

# calculate other/leasure crafts (its an unknown mix but assuming they are fairly small since no specific AIS category) Watt x hour

r <- stack[[5]]-stack[[1]]-stack[[2]]-stack[[3]]-stack[[4]]  # additional boats
r_other_wh <- r * 10000000000 * 0.04 # w hours at 130db normalized for average watt in 2.5km cell


# summary of toal noise pollution in watt hours
r_sum_wh <- r_cargo_wh + r_tanker_wh + r_pass_wh + r_fish_wh + r_other_wh
r_sum_wh
plot(r_sum_wh)

```



## resample noise to 1km grid 
```{r}

# Resample to approximately 1km resolution
r <- r_sum_wh

s <- raster(extent(r), resolution=res(r)/3, crs=crs(r))

r_res <- resample(r, s, method='bilinear')
r_res
plot(r_res)

raster_path <- paste(work_path, "noise_rs_wh_geo.tif", sep="")

writeRaster(r_res, raster_path, overwrite=T, COMPRESS=LZW)




inraster <- raster(raster_path)
inraster

# write empty grid for gdalutil work
writeRaster(grid_1km_na, paste(work_path, "grid_1km_gdalutil.tif", sep= ""),  overwrite=T, COMPRESS=LZW)

# write to file
dstfile <- paste(work_path, "grid_1km_gdalutil.tif", sep= "")

crs(grid_1km)
# 250m grid transformation from high resolution water mask

sr <- crs(inraster)
tr <- crs(grid_1km)
#  "+proj=cea +lat_ts=-12 +lon_0=12 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"    

r_1km <- gdalUtils::gdalwarp(srcfile = raster_path,
                                dstfile = dstfile,
                                s_srs = sr,
                                t_srs = tr,
                                tr = c(1000, 1000),
                                tap = TRUE,
                                output_Raster = FALSE,
                                overwrite = TRUE,
                                r = "average",
                                multi = TRUE,
)


r <- r_1km <- raster(dstfile)


# Final check that all is good...
r <- merge(r, grid_1km_0)
r <- crop(r, grid_1km)
r <- mask(r, grid_1km)

plot(r)

r_path <- paste(work_path, "noise_wh_grid1km.tif", sep="")

writeRaster(r, r_path, overwrite=TRUE, datatype = 'FLT4S', COMPRESS=LZW)

```

## summary impact and normalise to 140db and 24h*365 impact

```{r}

r_sum_wh_1km <- raster(r_path)

r_sum_norm <- r_sum_wh_1km /(2*365*24*100000000000*2.7*2.7) # normalized to number of days in data set ' 24h * watt at 140db (max impact in sensitivity matrix), normalized to hours for approx 1km cell from 2.7 km cell
r_sum_norm
r_sum_norm[r_sum_norm>1] <- 1 # all values higher then the hour/year/watt normalized max will be set as "maximum impact"

r <-r_sum_norm
plot(r)



r_path <- paste(work_path, "noise_wh_norm01_grid1km.tif", sep="")

writeRaster(r, r_path, overwrite=TRUE, datatype = 'FLT4S', COMPRESS=LZW)

noise_norm01_1km <- raster(r_path)

```

# SHIPSTRIKES
```{r}

stack <- stack_interp

names(stack)

f_high <- 1 # tanker, container. highest sensitivity relates to big ships at full speed -  assuming commercial traffic have bigger impact due to size and speed
f_med <- 0.2 # fishing boats / passenger ship, lower speed, smaller. approximation needs review
f_low  <- 0.1 # other / small crafts factor approximation needs review

r_other <- stack[[5]]-stack[[1]]-stack[[2]]-stack[[3]]-stack[[4]] 

ship_strikes_h <-  ((stack[[1]] + stack[[4]])*f_high + (stack[[2]] + stack[[3]])*f_med + (r_other)*f_low)

r <- ship_strikes_h
plot(r)
```

## resample noise to 1km grid 
```{r}

# Resample to approximately 1km resolution
r <- ship_strikes_h

s <- raster(extent(r), resolution=res(r)/3, crs=crs(r))

r_res <- resample(r, s, method='bilinear')
r_res
plot(r_res)

raster_path <- paste(work_path, "shipstrike_rs_h_geo.tif", sep="")

writeRaster(r_res, raster_path, overwrite=T, COMPRESS=LZW)




inraster <- raster(raster_path)
inraster

# write empty grid for gdalutil work
writeRaster(grid_1km_na, paste(work_path, "grid_1km_gdalutil.tif", sep= ""),  overwrite=T, COMPRESS=LZW)

# write to file
dstfile <- paste(work_path, "grid_1km_gdalutil.tif", sep= "")

crs(grid_1km)
# 250m grid transformation from high resolution water mask

sr <- crs(inraster)
tr <- crs(grid_1km)
#  "+proj=cea +lat_ts=-12 +lon_0=12 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"    

r_1km <- gdalUtils::gdalwarp(srcfile = raster_path,
                                dstfile = dstfile,
                                s_srs = sr,
                                t_srs = tr,
                                tr = c(1000, 1000),
                                tap = TRUE,
                                output_Raster = FALSE,
                                overwrite = TRUE,
                                r = "average",
                                multi = TRUE,
)


r <- r_1km <- raster(dstfile)


# Final check that all is good...
r <- merge(r, grid_1km_0)
r <- crop(r, grid_1km)
r <- mask(r, grid_1km)

plot(r)

r_path <- paste(work_path, "shipstrike_h_grid1km.tif", sep="")

writeRaster(r, r_path, overwrite=TRUE, datatype = 'FLT4S', COMPRESS=LZW)

```

## normalise
```{r}
r_path <- paste(work_path, "shipstrike_h_grid1km.tif", sep="")

r <- raster(r_path) 
r
ratio <- 2.7*2.7

l <- 100 # estimated max life expectancy (years) for whales

r_norm <- r*l/(ratio*2*365*24) # normalized using ship hours/year, 1km grid and estimated life expectancy (slow growing and long living species get higher risk of getting hit!) 
r_norm
plot(r_norm)

r_norm[r_norm>1] <- 1
r_norm[r_norm<0] <- 0
r_norm
plot(r_norm)

r <- r_norm



r_path <- paste(work_path, "shipstrike_norm01_grid1km.tif", sep="")

writeRaster(r, r_path, overwrite=TRUE, datatype = 'FLT4S', COMPRESS=LZW)

shipstrike_norm01_1km <- raster(r_path)
```

# SHIP POLLUTION 
Regulation of ships at anchor: Safety and environmental implications
https://www.sciencedirect.com/science/article/pii/S0308597X22000999

```{r}
stack <- ship_stack

names(stack)

r <- stack[[5]] # going only with "global layer here due to missing information on ship pollution related to category - assuming equal impact/h
w <- focalWeight(r, 0.04, "Gauss") # simulating spread of spill related pollutants assuming about 2 days dispersal time at 0.5 knot current speed
w
r_focal_20km <- focal(r, w, fun=sum, na.rm=T, pad=T, NAonly=F) 
plot(r_focal_20km)


w <- focalWeight(r, 0.01, "Gauss") # estimating spread of local pollutant effects
r_focal_5km <- focal(r, w, fun=sum, na.rm=T, pad=T, NAonly=F) 
plot(r_focal_5km)




r_pollution <- (r_focal_20km +r_focal_5km)/2

plot(r_pollution)

```

## resample pollution to 1km grid 
```{r}

# Resample to approximately 1km resolution
r <- r_pollution

s <- raster(extent(r), resolution=res(r)/3, crs=crs(r))

r_res <- resample(r, s, method='bilinear')
r_res
plot(r_res)

raster_path <- paste(work_path, "ship_pollution_rs_h_geo.tif", sep="")

writeRaster(r_res, raster_path, overwrite=T, COMPRESS=LZW)




inraster <- raster(raster_path)
inraster

# write empty grid for gdalutil work
writeRaster(grid_1km_na, paste(work_path, "grid_1km_gdalutil.tif", sep= ""),  overwrite=T, COMPRESS=LZW)

# write to file
dstfile <- paste(work_path, "grid_1km_gdalutil.tif", sep= "")

crs(grid_1km)


sr <- crs(inraster)
tr <- crs(grid_1km)
#  "+proj=cea +lat_ts=-12 +lon_0=12 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"    

r_1km <- gdalUtils::gdalwarp(srcfile = raster_path,
                                dstfile = dstfile,
                                s_srs = sr,
                                t_srs = tr,
                                tr = c(1000, 1000),
                                tap = TRUE,
                                output_Raster = FALSE,
                                overwrite = TRUE,
                                r = "average",
                                multi = TRUE,
)


r <- r_1km <- raster(dstfile)


# Final check that all is good...
r <- merge(r, grid_1km_0)
r <- crop(r, grid_1km)
r <- mask(r, grid_1km)

plot(r)

r_path <- paste(work_path, "ship_pollution_h_grid1km.tif", sep="")

writeRaster(r, r_path, overwrite=TRUE, datatype = 'FLT4S', COMPRESS=LZW)

```

## calculate pressure from ship hours
```{r}
r_path
r<-raster(r_path)

pollution_h_1km <- r

r_norm <- r/(24*52*2/2)   # assuming maximum spill/pollution event in sensitivity matrix occurs once every week, and stays for two weeks 
r_norm
r_norm[r_norm>1] <- 1

pollution_norm01_1km <- r_norm  # pollution impact ready for export!
plot(r_norm)

r_path <- paste(work_path, "ship_pollution_norm01_grid1km.tif", sep="")

writeRaster(pollution_norm01_1km, r_path, overwrite=TRUE, datatype = 'FLT4S', COMPRESS=LZW)

```


# RECREATIONAL BOATING 
```{r}
stack <- ship_stack

names(stack)

r <- stack[[3]] 

w <- focalWeight(r, 0.014, "Gauss") # estimating spread of disturbance from tourism activities (~half impact in direct pixel, half to the surrounding areas)
w
r_focal_5km <- focal(r, w, fun=sum, na.rm=T, pad=T, NAonly=F) 
plot(r_focal_5km)


r <- boating_h <- r_focal_5km   # ready for export
plot(r)
```

## Resample recreational boating to 1km grid 
```{r}

# Resample to approximately 1km resolution
r <- boating_h

s <- raster(extent(r), resolution=res(r)/3, crs=crs(r))

r_res <- resample(r, s, method='bilinear')
r_res
plot(r_res)

raster_path <- paste(work_path, "boating_rs_h_geo.tif", sep="")

writeRaster(r_res, raster_path, overwrite=T, COMPRESS=LZW)




inraster <- raster(raster_path)
inraster

# write empty grid for gdalutil work
writeRaster(grid_1km_na, paste(work_path, "grid_1km_gdalutil.tif", sep= ""),  overwrite=T, COMPRESS=LZW)

# write to file
dstfile <- paste(work_path, "grid_1km_gdalutil.tif", sep= "")

crs(grid_1km)


sr <- crs(inraster)
tr <- crs(grid_1km)
#  "+proj=cea +lat_ts=-12 +lon_0=12 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"    

r_1km <- gdalUtils::gdalwarp(srcfile = raster_path,
                                dstfile = dstfile,
                                s_srs = sr,
                                t_srs = tr,
                                tr = c(1000, 1000),
                                tap = TRUE,
                                output_Raster = FALSE,
                                overwrite = TRUE,
                                r = "average",
                                multi = TRUE,
)


r <- r_1km <- raster(dstfile)


# Final check that all is good...
r <- merge(r, grid_1km_0)
r <- crop(r, grid_1km)
r <- mask(r, grid_1km)

plot(r)

r_path <- paste(work_path, "boating_h_grid1km.tif", sep="")

writeRaster(r, r_path, overwrite=TRUE, datatype = 'FLT4S', COMPRESS=LZW)

```

## calculate pressure from ship hours
```{r}
r_path
r <- boating_h_1km <- raster(r_path)

r_norm <- r/(8*52*2/2)   # assuming max disturbance from tourism activities in sensitivity matrix occurs when impacted 8h / week, and last for 2 weeks
r_norm
r_norm[r_norm>1] <- 1

boating_norm01_1km <- r_norm  # pollution impact ready for export!
plot(r_norm)

r_path <- paste(work_path, "boating_norm01_grid1km.tif", sep="")

writeRaster(boating_norm01_1km, r_path, overwrite=TRUE, datatype = 'FLT4S', COMPRESS=LZW)

```



#==========================================================================================================================#
# EXPORT
#==========================================================================================================================#
# Write products to root directory when checked and ready to use elsewhere, accompanied by yourfilename.ext_sourcesym.txt 
# for each file to track source IDs as they accumulate between processes / data


# SOURCESYM
### Create sourcesym file with all source IDs used (data and data_raw) 
# data raw sources (dir)
```{r}
print(paste("data_raw metasym files used = ", data_raw_metasym_num, sep=""))
(data_raw_sources <- tibble(id = c(source1))) # REPLACE / add all data_raw sources
```

## data sources (files)
```{r}
print(paste("data_sourcym files = ", data_sourcesym_num, sep=""))
data_sources <- data1_sourcesym %>% # add any "sourcesym" files applicable to this product (check indata section), keep adding sourcesym files as needed
  add_row(data2_sourcesym)%>%  
  #add_row(data3_sourcesym)%>% 
  #add_row(data4_sourcesym)%>% 
  #add_row(data5_sourcesym)%>% 
  #add_row(data6_sourcesym)%>% 
  unique() %>% 
  print()
```
## Sources combined
```{r}
(all_sources <- data_raw_sources %>% add_row(data_sources))
```
### if a product only use a subset of the total sources, copy the section above and create individual files below (e.g. product1_sources <- ...)


# PRODUCT 1: 
## read objects
```{r}
product_orig <- noise_norm01_1km # input and check your object to be written to file
product_orig2 <- shipstrike_norm01_1km # input and check your object to be written to file
product_orig3 <- pollution_norm01_1km # input and check your object to be written to file
product_orig4 <- boating_norm01_1km # input and check your object to be written to file

plot(product_orig)
plot(product_orig2)
plot(product_orig3)
plot(product_orig4)

```
## set names
```{r}
scale <- "1km" # "1km" / "250m"
unit <- "wh_year_norm01"  # add unit for product with original values (e.g. m, km, perc, presence, proportion
unit2 <- "h_year_norm01"  # add unit for product with original values (e.g. m, km, perc, presence, proportion
unit3 <- "h_year_norm01"  # add unit for product with original values (e.g. m, km, perc, presence, proportion
unit4 <- "h_year_norm01"  # add unit for product with original values (e.g. m, km, perc, presence, proportion

component_names <- read_tsv("./shiny_data_upload/component_names.txt") 
component_names_selected <- component_names %>%
  dplyr::filter(theme_folder == theme) %>% 
  dplyr::select(group, file_name, description) %>% 
  print(n = Inf)
```

```{r}
descriptive_name <- "underwater_noise"  # name of product, for pressure/ecosystem components select from name list above
descriptive_name2 <- "ship_strike"  # name of product, for pressure/ecosystem components select from name list above
descriptive_name3 <- "ship_pollution"  # name of product, for pressure/ecosystem components select from name list above
descriptive_name4 <- "boating"  # name of product, for pressure/ecosystem components select from name list above

```
## paths
```{r}
(product_orig_path <- paste(dest_path, descriptive_name, "_", unit, "_", scale, "_", version, d_version, ".tif", sep="" ))
(product_orig2_path <- paste(dest_path, descriptive_name2, "_", unit2, "_", scale, "_", version, d_version, ".tif", sep="" ))
(product_orig3_path <- paste(dest_path, descriptive_name3, "_", unit3, "_", scale, "_", version, d_version, ".tif", sep="" ))
(product_orig4_path <- paste(dest_path, descriptive_name4, "_", unit4, "_", scale, "_", version, d_version, ".tif", sep="" ))




```
## Write to file
```{r}
# check datatype to be appropriate for values (flt or int, 'INT4S' option to..)
writeRaster(product_orig, product_orig_path, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
writeRaster(product_orig2, product_orig2_path, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
writeRaster(product_orig3, product_orig3_path, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
writeRaster(product_orig4, product_orig4_path, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)


# write sourcesym file, change "all_sources" to "selected_sources" if subset only for specific file
write_tsv(all_sources, paste(product_orig_path, "_sourcesym.txt", sep=""))  
write_tsv(all_sources, paste(product_orig2_path, "_sourcesym.txt", sep="")) 
write_tsv(all_sources, paste(product_orig3_path, "_sourcesym.txt", sep="")) 
write_tsv(all_sources, paste(product_orig4_path, "_sourcesym.txt", sep="")) 
```


# SAVE SCRIPT
## 1. save your current R script (File/Save)
## 2. Save a carbon copy of the R script to proc_log (identical script to what is used to create the products exported
## Do not change the *.Rmd carbon copy in proc_log/ unless you change this version of the products (new versions are developed under process/r)

```{r}
# 2. run code below and go to File/Save As and paste link and name:
dest_path_script <- paste(getwd(), proc_path)
print(dest_path_script <- gsub(" .", "", dest_path_script)) # path
print(script_name_copy <- paste(gsub(".Rmd", "", script_name), d_version, ".Rmd", sep="")) # script name
```

#==========================================================================================================================#
# FINAL CHECK
#==========================================================================================================================#
# Developer check
## Do your own final check on the products and double check sourcesym file exists and are complete for all files in root 
## -> Sign at top of script in about section (# Approved by:)

# External check 
## To ensure repeatability and quality, make sure one colleage can run the script
## When script is proven repeatable/understandable and products/metadata look ok 
## -> Sign at top of script in about section (# Approved by:), and make any comments if needed
## There is also a work_log document in onedrive / gitHUB, check current arrangement with Swedish dev team
