# =========================================================================================================
# Script by Edmond Sacre, SLU
# This script creates rasters from the from the shapefiles available on the IUCN red list of threatened species
# (https://www.iucnredlist.org/resources/spatial-data-download)
# This script also resamples the rasters to the same resolution and extent as the standard WIOSym grid
# =========================================================================================================

# PREPARATIONS
```{r, include = FALSE}
# Install / load packages
#x <- c("sf", "raster", "tidyr", "dplyr")
#install.packages(x, dependencies=TRUE)
```

## Install packages
```{r, include = FALSE}
library(sf) 
library(raster)
library(fasterize)
library(tidyr)
library(dplyr)
library(rgdal)
library(tidyverse)
library(leaflet)
```

## View folders/themes
```{r}
folders_raw <- read.table("./shiny_data_upload/folders.txt", sep = "\t", header = TRUE)
folders_data <- read.table("./process/templates/folders_data.txt", sep = "\t", header = TRUE)
#folders <- folders_raw %>% add_row(folders_data)
```

## Define versions/directories
```{r}
version = "v01"
d_version = ".0"
location = "reg"   #check
theme1 <- "act"   #check
subtheme1 <- "fishing"   #check
name1 <- "ghost_fishing"   #check
name2 <- "" #check
(dest1_path <- paste("./data", location, theme1, subtheme1, name1, version, sep="/"))
(work1_path <- paste(dest1_path, "proc", sep="/"))
(proc1_path <- paste(dest1_path, "proc_log", sep="/"))
script_path <- "./process/r"
(script_path_windows <- paste(getwd(), "/process/r", sep=""))
(script_name <- paste(theme1, "_", subtheme1,"_", location, "_", name1,"_s01_",version, ".R", sep=""))
```

## Create directories
```{r}
dir.create(dest1_path, recursive = TRUE)
dir.create(work1_path, recursive = TRUE)
dir.create(proc1_path, recursive = TRUE)
```

## Read in WIOSym standard grid
```{r}
grid_1km <- raster("./data/reg/grid/grid/v01/grid_1km_v01.1.tif")
```

## Define sources
```{r}
num_sources <- 1  # write the number of sources you will use in this script (to be used later for exporting metadata)
  
source1_dir <-  "./data_raw/glo/act/fishing_com/gfw/es2105270900/"    #check
source1_id <- "es2105270900"    #check
source1_metasym <- read.table(paste0(source1_dir, source1_id, "_metasym.txt"), sep = "\t")
#source1_metasym # check that metasym file is populated, and fix if not - before finalising the processing

source_list <- c(source1_id) # for multiple sources, should look like - c(source1_id, source2_id, source3_id) etc.
```

*****************************************************************
*****************************************************************
# PROCESSING

```{r}
library(sp)
library(rgdal)
library(leaflet)
library(spatialEco)
#library(RKUM)
library(mapview)
```

```{r}
gg_drift_longlines <- raster(paste0(work1_path, "/fishing_com_fishing_effort_drifting_longlines_v01.0_unstandardized.tif")) * 0.20
gg_pole_and_line <- raster(paste0(work1_path, "/fishing_com_fishing_effort_pole_and_line_v01.0_unstandardized.tif")) * 0.20
gg_set_longlines <- raster(paste0(work1_path, "/fishing_com_fishing_effort_set_longlines_v01.0_unstandardized.tif")) * 0.20
gg_pots_and_traps <- raster(paste0(work1_path, "/fishing_com_fishing_effort_pots_and_traps_v01.0_unstandardized.tif")) * 0.19
gg_purse_seines <- raster(paste0(work1_path, "/fishing_com_fishing_effort_purse_seines_v01.0_unstandardized.tif")) * 0.066 
gg_tuna_purse_seines <- raster(paste0(work1_path, "/fishing_com_fishing_effort_tuna_purse_seines_v01.0_unstandardized.tif")) * 0.066
gg_trawling <- raster(paste0(work1_path, "/fishing_com_fishing_effort_trawlers_v01.0_unstandardized.tif")) * 0.12 
gg_dredging <- raster(paste0(work1_path, "/fishing_com_fishing_effort_dredge_fishing_v01.0_unstandardized.tif")) * 0.18 
gg_set_gillnets <- raster(paste0(work1_path, "/fishing_com_fishing_effort_set_gillnets_v01.0_unstandardized.tif")) * 0.084


gg_drift_longlines[is.na(gg_drift_longlines)] <- 0
gg_pole_and_line[is.na(gg_pole_and_line)] <- 0
gg_set_longlines[is.na(gg_set_longlines)] <- 0
gg_pots_and_traps[is.na(gg_pots_and_traps)] <- 0
gg_purse_seines[is.na(gg_purse_seines)] <- 0
gg_tuna_purse_seines[is.na(gg_tuna_purse_seines)] <- 0
gg_trawling[is.na(gg_trawling)] <- 0
gg_dredging[is.na(gg_dredging)] <- 0
gg_set_gillnets[is.na(gg_set_gillnets)] <- 0

ghost <- gg_drift_longlines+gg_pole_and_line+gg_set_longlines+gg_pots_and_traps+
          gg_purse_seines+gg_tuna_purse_seines+gg_trawling+gg_dredging+gg_set_gillnets
ghost[ghost < 0.1] <- NA
spec_sp <- rasterToPoints(ghost, spatial = TRUE)
spec_sp$value <- spec_sp$layer

spec_kernel <- sp.kde(x = spec_sp,
                      y = spec_sp$value,
                      newdata = as.vector(extent(grid_1km)),
                      nr=500,
                      nc=500,
                      bw = 500000, 
                      standardize = TRUE,
                      mask = FALSE)
spec_kernel <- resample(spec_kernel, grid_1km)
spec_kernel[is.na(grid_1km)==T] <- NA
```

```{r}
product1 <- spec_kernel
```

# Calculate the max values that should be used in the standarized raster
# Here I take the average of all gear loss rates above
# Therefore we set the maximum value in the raster to 15
```{r}
mean(c(0.20, 0.20, 0.20, 0.19, 0.066, 0.066, 0.12, 0.18, 0.084))
```


*****************************************************************
*****************************************************************

# EXPORT

## Specify output directory and product file name

```{r}
(dest1_path_file1 <-paste0(dest1_path, "/", subtheme1, "_", name1, "_", version, d_version, ".tif"))
```

## Export metadata
```{r}
sourcesym <- data.frame(matrix(data = NA, nrow = num_sources, ncol = 1))
colnames(sourcesym) <- "id"

for(i in 1:num_sources){
  sourcesym[i,1] <- source_list[i]
}

sourcesym
```

```{r}
sourcesym_path <- paste0(dest1_path_file1, "_sourcesym.txt")
write.table(sourcesym, sourcesym_path, row.names = F, sep = "\t")
```


## Export product

```{r}
writeRaster(product1, dest1_path_file1, COMPRESS=LZW, overwrite=TRUE)
```

