# ========================================================================================================= #
# IMPORTANT - To get started:
Because this script is an R Markdown script, the default directory is the location of this .Rmd file.
You must open this script from within the .Rproj file associated with your wiosym database for it to work.
If you haven't used R notebooks before (.Rmd), each of the code "chunks" can be executed by clicking the green play
button at the top right of the chunk.

To run the script from the project directory, in R Studio, you need to go to: Tools > Global Options > R Markdown
and set "Evaluate chunks in directory" to "Project", though for latest version of R/Rstudio it may not be needed...
# ========================================================================================================= #
# ABOUT
# ========================================================================================================= #
## Brief description: 
Script to define temperate and tropical zone from SST and bottom temp data
### Scripts/processes that needs to be updated prior?:
mean_bottom_temperature_v2.0.ipynb, sea_surface_temperature_v2.0.ipynb
### Script by: Initial, org, Rver: 
gk, sgu, r v4.1.1. (a continuation from work done by Queste B in v01 2022)
### Updated: Initial, org, Rver, comments
### Developer check: Initial, org, yymmdd, comments
### External check: Initial, org, yymmdd, comments
# ========================================================================================================= #
# INSTRUCTIONS
# ========================================================================================================= #    
This is a simplified template that is made to adjust already existing data products, for a full template to produce new data use the main temaplte
# ========================================================================================================= # 
# PREPARATIONS
# ========================================================================================================= #

### Install packages
```{r, include = FALSE}
# list all packages that this script is dependent on, remove any redundant ones
x <- c("sf", "raster", "rgdal", "tidyverse")
#install.packages(x, dependencies=TRUE)
```

### Load packages
```{r, include = FALSE}
library(sf) 
library(raster)
library(rgdal)
library(tidyverse)
```

## Set version / theme / subtheme (copy from original script or path to get right..)
```{r}
version = "v02" # relates to major iteration of WIOSym (v01 for grid 2021, v02 for grid 2022)
d_version = ".0"  # detailed version (you can add another level e.g. 0.0  -> v01.0.0)
seq = "s02" # sequential order - change accordingly if process is run in several scripts
location = "reg"
theme = "env"
subtheme = "ocean"
name1 <- "_temptrop" # add name what this script does if appropriate ("uncertainty", "norm01")
```


## Set paths
Locate the relative path (from wiosym folder) to the root directpry of the data folder you are working in, copy and paste to "dest_path", the rest will update itself
```{r}
#(dest_path <- "REPLACE") # path to final product
(dest_path <- "./data/reg/env/ocean/v02/") # example of path to product directory

(work_path <- paste(dest_path, "proc/", sep=""))
(proc_log_path <- paste(dest_path, "proc_log/", sep=""))
(archive_path <- paste(dest_path, "_archive/", sep=""))
(script_path <- "./process/r/")
```

## Create directories
```{r}
dir.create(dest_path, recursive = TRUE)
dir.create(work_path, recursive = TRUE)
dir.create(proc_log_path, recursive = TRUE)
dir.create(archive_path, recursive = TRUE)
dir.create(script_path, recursive = TRUE)
```


## Save R Script
```{r}
(script_path_windows <- paste(getwd(), "/process/r", sep=""))
(script_name <- paste(theme, "_", subtheme,"_", location, "_", seq, "_", version, name1, ".Rmd", sep=""))
```
 # copy path and name and use File/Save As in R studio to save rmd file
 
#==========================================================================================================================#
# INDATA
#==========================================================================================================================#


# Set DATA_RAW sources
IMPORTANT all data harvested from data_raw needs to have a metasym.txt file in its root directory (not file specific)
When building on a a previous script you ight already have a complete sourcesym file that you can copy and paste for the product you do in this script (filename_sourcesym.txt), its ok! 
## copy and paste your data_raw sources from previous script. 
if you add new ones make sure to ingest the metasym file (see original template if code example is needed)

## DIR1: Bathymetry uncertainty from gebco
```{r}
#source1_dir <- source_dir <-  ""  # example "./data_raw/reg/env/topo/gebco/gk2110122231/"
#dir(source_dir) # check what data is in data_raw directory
#source_file <- ""  # your data file(s)
#source1_path2 <- source_path <- paste(source_dir, source_file, sep="")
#source_path
#(gebco_tid <- raster(source_path))
#plot(gebco_tid)
```


# DATA sources
IMPORTANT Products in data are based on information in data_raw. Check that *_sourcesym.txt exists for each file to track acccumulated IDs. 
The sourcesym files can be identical within some directory in which case you can load only one, but its not always the case which is why there is one file for each file. 

## DATA1: Grid 1km - loading all flavours (watermask 1, 0, NA) at once since identical sourcesym apply
```{r}
data_dir <-  "./data/reg/grid/grid/v01/"  # input your data directory
dir(data_dir) # check what data is in this directory

data1_file <- "grid_1km_v01.1.tif" 
data1_file2 <- "grid_1km_0_v01.1.tif" 
data1_file3 <- "grid_1km_na_v01.1.tif" 

(grid_1km_path <- path <- paste(data_dir, data1_file, sep=""))
(grid_1km <- raster(path)) # read and check your data
(grid_1km_0_path <- path <- paste(data_dir, data1_file2, sep=""))
(grid_1km_0 <- path <- raster(path)) # read and check your data
(grid_1km_na_path <- path <- paste(data_dir, data1_file3, sep=""))
(grid_1km_na <- path <- raster(path)) # read and check your data
(data1_sourcesym <- read_tsv(paste(data_dir, data1_file, "_sourcesym.txt", sep="")))
# check that source IDs exists and make sense

#plot(grid_1km)
```

## DATA2: Grid 250m - loading all flavours (watermask 1, 0, NA) at once since same sourcesym apply
```{r}
data2_dir <- data_dir <-  "./data/reg/grid/grid/v01/"  # input your data directory

data2_file <- "grid_250m_v01.1.tif" # your data file
data2_file2 <- "grid_250m_0_v01.1.tif" # your data file
data2_file3 <- "grid_250m_na_v01.1.tif" # your data file

grid_250m_path <- path <- paste(data_dir, data2_file, sep="")
grid_250m <- raster(path) # read and check your data
grid_250m_0_path <- path <- paste(data_dir, data2_file2, sep="")
grid_250m_0 <- path <- raster(path) # read and check your data
grid_250m_na_path <- path <- paste(data_dir, data2_file3, sep="")
grid_250m_na <- path <- raster(path) # read and check your data
(data2_sourcesym <- read_tsv(paste(data_dir, data2_file, "_sourcesym.txt", sep="")))
# check so source IDs exists and make sense
# check that source IDs exists and make sense
# plot(grid_250m)
```

## DATA3: Sea Temperature data 
```{r}
dir(dest_path)

SBT <- raster(paste(dest_path, "grid_1km_bottomT_byq.tif", sep=""))
SST <- raster(paste(dest_path, "grid_1km_SST_byq.tif", sep=""))

# Load sourcesym file

dir(dest_path, pattern = "sourcesym.txt")

(SBT_sourcesym <- read_tsv(paste(dest_path, "grid_1km_bottomT_byq_sourcesym.txt", sep="")))
(SST_sourcesym <-  read_tsv(paste(dest_path, "sea_surface_temperature_v2.0_sourcesym.txt", sep="")))


plot(SBT)
plot(SST)
```


#==========================================================================================================================#
# PROCESSING
#==========================================================================================================================#

#Hypothesis on temp range
Temperate Tropical limits were informed by  Stuart-Smith RD, Edgar GJ, Bates AE. Thermal limits to the geographic distributions of shallow-water marine species. Nature Ecology & Evolution. 2017 Dec;1(12):1846-52. https://www.nature.com/articles/s41559-017-0353-x
As well as Oceanographer Dr. Queste B, Gothenburg University


# SST temp/trop layer
```{r}
#reclassify values 0-1

r <- SST
writeRaster(r, paste(work_path, "sst.tif", sep=""), COMPRESS=LZW, datatype = 'INT4S', overwrite=TRUE)
r <- raster(paste(work_path, "sst.tif", sep="")) # due to strange problem with value range the SBT was exported and then imported
r

minv <- 23000000
maxv <- 25500000

r[r<=minv]<- minv 
r[r>=maxv]<- maxv
r <- r-minv

sst01_trop <- r/maxValue(r)
plot(sst01_trop)
writeRaster(sst01_trop, paste(work_path, "sst01_trop.tif", sep=""), COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)

sst01_temp <- -1*(sst01_trop-1)
plot(sst01_temp)
writeRaster(sst01_temp, paste(work_path, "sst01_temp.tif", sep=""), COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
```


# SBT temp/trop layer
Sea Bottom Temperature (used for habitat deeper then 40m)
Mean Annual Sea Bottom temperatures 2011-2016 from https://doi.org/10.48670/moi-00021 were used 
```{r}
r <- SBT
writeRaster(r, paste(work_path, "sbt.tif", sep=""), COMPRESS=LZW, datatype = 'INT4S', overwrite=TRUE)
r <- raster(paste(work_path, "sbt.tif", sep="")) # due to strange problem with value range the SBT was exported and then imported
r
# temp range slightly lower then for SST based on observation of data/model differences

minv <- 21500000
maxv <- 24000000

r[r<=minv]<- minv 
r[r>=maxv]<- maxv
r <- r-minv

sbt01_trop <- r/maxValue(r)
plot(sbt01_trop)
writeRaster(sbt01_trop, paste(work_path, "sbt01_trop.tif", sep=""), COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)

sbt01_temp <- -1*(sbt01_trop-1)
plot(sbt01_temp)
writeRaster(sbt01_temp, paste(work_path, "sbt01_temp.tif", sep=""), COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
```




```

#==========================================================================================================================#
# EXPORT
#==========================================================================================================================#

# SOURCESYM
Create sourcesym file with all source IDs used (data and data_raw) 
```{r}
## data sources (files)

(data_source1 <- SST_sourcesym) 
(data_source2 <- SBT_sourcesym) 

```


# PRODUCT 1: Uncertainty
OBS if you need pointers for other naming, the full template has the instructions e.g. pointing to the component list with names, and using theme/subtheme to make the name...
```{r}

export_object1 <- sst01_temp
export_object2 <- sst01_trop

export_object3 <- sbt01_temp
export_object4 <- sbt01_trop

dir(dest_path)
product_orig1 <- "temp_limit01_sst_1km.tif" 
product_orig2 <- "trop_limit01_sst_1km.tif" 
product_orig3 <- "temp_limit01_sbt_1km.tif" 
product_orig4 <- "trop_limit01_sbt_1km.tif" 


(product_path1 <- paste(dest_path, product_orig1, sep="" ))
(product_path2 <- paste(dest_path, product_orig2, sep="" ))
(product_path3 <- paste(dest_path, product_orig3, sep="" ))
(product_path4 <- paste(dest_path, product_orig4, sep="" ))




# check datatype to be appropriate for values ('FLT4S' (normalised data) / 'INT4S' (uncertainty data) main standards to use)
writeRaster(export_object1, product_path1, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
writeRaster(export_object2, product_path2, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
writeRaster(export_object3, product_path3, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
writeRaster(export_object4, product_path4, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)


write_tsv(data_source1, paste(product_path1, "_sourcesym.txt", sep="")) 
write_tsv(data_source1, paste(product_path2, "_sourcesym.txt", sep="")) 
write_tsv(data_source2, paste(product_path3, "_sourcesym.txt", sep="")) 
write_tsv(data_source2, paste(product_path4, "_sourcesym.txt", sep="")) 

```


# SAVE SCRIPT
1. save your current R script (File/Save)
2. Save a carbon copy of the R script to proc_log (identical script to what is used to create the products exported
Do not change the *.Rmd carbon copy in proc_log/ unless you change this version of the products (new versions are developed under process/r)

```{r}
# 2. run code below and go to File/Save As and paste link and name:
#(script_name <- paste(theme, "_", subtheme,"_", location, "_", seq, "_", version, d_version, name1, ".Rmd", sep=""))
```

#==========================================================================================================================#
# FINAL CHECK
#==========================================================================================================================#
# Developer check
 Do your own final check on the products and double check sourcesym file exists and are complete for all files in root 
 -> Sign at top of script in about section (# Approved by:)

# External check 
 To ensure repeatability and quality, make sure one colleage can run the script
 When script is proven repeatable/understandable and products/metadata look ok 
 -> Sign at top of script in about section (# Approved by:), and make any comments if needed
 There is also a work_log document in onedrive / gitHUB, check current arrangement with Swedish dev team
