---
title: "wiosym metadata search"
output: html_notebook
---

NOTE: You must open this script from within the .Rproj file associated with your wiosym database for it to work.

If you haven't used R notebooks before (.Rmd), each of the code "chunks" can be executed by clicking the green play 
button at the top right of the chunk.


### List metadata files

```{r}
#meta <- list.files(pattern = "_metasym.txt", recursive = TRUE, full.names = TRUE) # edited -- to check only on /data_raw directory
meta <- list.files(path = "./data_raw/", pattern = "_metasym.txt", recursive = TRUE, full.names = TRUE)
meta
```

```{r}
meta_df <- data.frame(matrix(data = NA, nrow = length(meta), ncol = 2))
colnames(meta_df) <- c("file_path", "folder_path")

for(i in 1:length(meta)) {
  
  meta_df$file_path[i] <- meta[i]
  meta_df$folder_path[i] <- dirname(meta[i])
  
}
```


### Compile the metadata files into a master data frame

```{r}
cnames <- c("id", "file_path", "local_path", "download_date", "user", "provider_list", "provider_manual",
            "source", "citation", "copyright", "copyright_details", "location", "theme", "theme_tag",
            "subtheme", "subtheme_tag", "tags", "comments" 
            )

meta_master <- data.frame(matrix(data = NA, nrow = 0, ncol = 18))
colnames(meta_master) <- cnames

#temp <- read.table(meta[1], sep = "\t")
#meta_master[1,] <- temp[,2]

for(i in 1:length(meta)){
  
  temp <- read.table(meta[i], sep = "\t")
  df <- data.frame(matrix(data = NA, nrow = 1, ncol = 18))
  colnames(df) <- cnames
  df[1,] <- temp[,2]
  
  meta_master <- rbind(meta_master, df)
  
  
}

meta_master
```


### Check to see that all of the metadata files have data with them

```{r}

meta_df$id <- meta_master$id
meta_df$local_path <- meta_master$local_path
meta_master$full_path <- NA


for(i in 1:length(meta)){
  
  x <- list.files(meta_df$folder_path[i])
  
  #p <- as.character(p)
  p <- paste0(meta_df$local_path[i], "/", meta_df$folder_path[i])
  p <- gsub("/", "\\\\", p)
  p <- gsub(p, pattern="\\\\", replacement="\\", fixed=TRUE)
  #p <- paste0(meta_df$local_path[i], "\n", p)
  
  true <- paste0("ID:", meta_df$id[i], " - data appears to be in the folder")
  false <- paste0("ID:", meta_df$id[i], " - WARNING: NO DATA IN THE FOLDER, check ", p)
  
  meta_master$full_path[i] <- p
  
  z <- ifelse(length(x) > 1, true, false)
  
  cat(z, '\n')
  
}

```


### List files in each dataset

```{r}
for(i in 1:length(meta)) {

cat("\nFiles in", meta_df$id[i], ": \n")
cat("    ", paste0("", list.files(meta_df$folder_path[i]), sep = " ---"), "\n")

}
```

****************************************************
# Tag search


### List tags present in the existing metadata in the wiosym database
```{r}

tag_list <- c("coral")

for (i in 1:length(meta)) {
  
  x <- strsplit(meta_master$tags[i], ";")[[1]]
  tag_list <- c(tag_list,x)
  
}

tag_list <- unique(tag_list[tag_list != ""])

sort(tag_list)
```

### Search for the ID and paths of data containing specific tags
```{r}
tags <-  c("coral") #multiple tags can be searched with e.g. c("eco", "fish")

tag_search <- meta_master[grep(paste(tags, collapse = "|"), meta_master$tags), c("id", "user", "full_path", "tags")]
tag_search
```

### Print file paths for use in Windows Explorer
```{r}
cat(tag_search$full_path, sep = "\n")
```

****************************************************
### Export metadata master file
```{r}
write.csv(meta_master, "metasym_master.csv", row.names = FALSE)
```

