### IMPORTANT: Because this script is an R Markdown script, the default directory is the location of this .Rmd file.
# To run the script from the project directory, in R Studio, you need to go to: Tools > Global Options > R Markdown
# and set "Evaluate chunks in directory" to "Project"
# -----------------------------------------------------------------------------------------------------------

### ABOUT
# ========================================================================================================= #
# Brief description: 
# Script by: Initial, org
# Updated: Initial, org, comments
# Developer check: Initial, org, yymmdd
# External check: Initial, org, yymmdd, comments
# ========================================================================================================= #

### INSTRUCTIONS
# ========================================================================================================= #    
# Style
  # Write code that is easy to follow and repeat.
  # Suggested R style is tidyverse. Style guide here https://style.tidyverse.org/index.html

# Metadata
  # All indata from "data_raw" must have a unique ID with *_metasym.txt file to go along, check before you add data.
  # All indata from and outputs to "data" must have a *_sourcesym.txt file with all (accumulated) sourceIDs used to create the layer.
  # The metasym files and tracking of those in the sourcesym files ensures correct sources for products and ISO metadata. This template helps you get it right..
  
# Naming and folders
  # Predefined standard names for main/sub folders and final products in two text files that can be added to in need.
  # ..\shiny_data_upload\modify_txt_files_v01.2.xlsx and ..\process\templates\modify_txt_files_data_v01.0.xlsx
  # Main version (e.g. v01 for 2021) follows dev cycles for all WIOSYm. For version control within v01 use V01.1, V01.2 etc..
  # More code/instructions to standardize folders and names under "set destinations" section
  # Each product dir have a "proc" folder. It's the temp space where all workfiles are stored. No final products here.  
  # All final products are stored under the root folder (e.g. ..\data\reg\eco\bh_coral\v01) together with 
  # a nameofmyfile_sourcesym.txt file with all source IDs used. No other files here. move products to _archive\ dir if defunct 
# ========================================================================================================= # 
### PREPARATIONS
# ========================================================================================================= #

# Install packages
```{r, include = FALSE}
# list all packages that this script is dependent on, remove any redundant ones
x <- c("leaflet", "sf", "raster", "rgdal", "tidyverse")
#install.packages(x, dependencies=TRUE)
```

# Load packages
```{r, include = FALSE}
library(leaflet)
library(sf) 
library(raster)
library(rgdal)
library(tidyverse)
```

## Set version
```{r}
version = "v01" # relates to major iteration of WIOSym (v01 for grid 2021, v02 for grid 2022)
d_version = ".0"  # detailed version (you can add another level e.g. 0.0  -> v01.0.0)
seq = "s01" # sequential order - change accordingly if process is run in several scripts
```

## Set geographic scope
```{r}
read_tsv("./shiny_data_upload/locations.txt")  # prints available locations to choose from and their abbreviations
```
```{r}
location = "reg" # choose applicable 3 letter abbreviation from "location_val" -  reg for whole WIO area
```

## Set main theme
```{r}
folders_raw <- read_tsv("./shiny_data_upload/folders.txt")
folders_data <- read_tsv("./process/templates/folders_data.txt") # additional folders for data not applicable for data_raw 
folders <- folders_raw %>% add_row(folders_data)
(unique(folders["theme_folder"]))
```
```{r}
theme <- "REPLACE" # Copy from "theme_folder" (e.g. "eco")
```

## Set Subtheme
```{r}
folders %>%
  filter(theme_folder == theme) %>% 
  dplyr::select(subtheme_folder, subtheme_name) %>% 
  print(n = Inf)
```
```{r}
subtheme <- "REPLACE" # copy from "subtheme_folder"
```

## Set paths
```{r}
(dest_path <- paste("./data", location, theme, subtheme, version, sep="/")) # path to final product
(work_path <- paste(dest_path, "proc", sep="/"))
(proc_path <- paste(dest_path, "proc_log", sep="/"))
(archive_path <- paste(dest_path, "_archive", sep="/"))
(script_path <- "./process/r")
```

## Create directories
```{r}
dir.create(dest_path, recursive = TRUE)
dir.create(work_path, recursive = TRUE)
dir.create(proc_path, recursive = TRUE)
dir.create(archive_path, recursive = TRUE)
dir.create(script_path, recursive = TRUE)
```

## Save R Script
```{r}
name1 <- "" # "REPLACE_" if script need additional name... don't use unless needed to avoid duplication.
(script_path_windows <- paste(getwd(), "/process/r", sep=""))
(script_name <- paste(theme, "_", subtheme,"_", location, "_", name1, seq, "_", version, ".Rmd", sep=""))
```
 # copy path and name and use File/Save As in R studio to save rmd file
 
#==========================================================================================================================#
### INDATA
#==========================================================================================================================#

## Set DATA_RAW sources
# IMPORTANT all data harvested from data_raw needs to have a metasym.txt file in its root directory (not file specific)
# Data_raw DIR1
```{r}
source1_dir <- source_dir <-  "./data_raw/.../uniqueid/"  # REPLACE - internal path
source1 <- source_id <- "uniqueid" # REPLACE
(read_tsv(paste(source_dir, source_id, "_metasym.txt", sep=""))) # check the metasym file
``` 
```{r}
dir(source_dir, recursive=F)
```

# Files DIR1
```{r}
object_name <- st_read(paste(source_dir, "name.ext", sep="")) # REPLACE object_name with descriptive short name and "name.ext" - copy internal path to file in source dir
# keep adding files from source 1 as needed
```

## Data_raw DIR2
```{r}
# ... keep adding sources/files, copy above
```

# declare number of "data_raw" directorys
```{r}
data_raw_metasym_num <-  # REPLACE - set total number of data_raw dir used (e.g. <- 1), to help in export section
```




## Set "DATA" sources
# IMPORTANT Products in data are based on information in data_raw. Check that *_sourcesym.txt exists for each file to track acccumulated IDs. 

# DATA 1: WIOSym standard grid 1km
```{r}
data_dir <-  "./data/reg/grid/grid/v01/"  # input your data directory
dir(data_dir) # check what data is in this directory
```
```{r}
data_file <- "grid_1km_v01.1.tif" # REPLACE - your data file
(data1_sourcesym <- read_tsv(paste(data_dir, data_file, "_sourcesym.txt", sep="")))
# check that source IDs exists and make sense
```
```{r}
(grid_1km <- raster(paste(data_dir, data_file, sep=""))) # read and check your data
```
```{r}
plot(grid_1km)
```

# DATA 2: ...  
```{r}
data_dir <-  "./data/.../v01/"  # REPLACE - input your data directory
dir(data_dir) # check what data is in this directory
```
```{r}
data_file <- "REPLACE" # your data file (must be under the root of data_dir, do not use files from proc/)
(data2_sourcesym <- read_tsv(paste(data_dir, data_file, "_sourcesym.txt", sep=""))) # REPLACE - check/change data number (e.g. data3_sourcesym)
# check that source IDs exists and make sense
```
```{r}
(object_name <- raster(paste(data_dir, data_file, sep=""))) # REPLACE "object_name" - read in your data object
```
```{r}
plot(object_name) # REPLACE
```


# DATA3: ...
# keep adding as needed, copy from data 2 above

# declare number of "data" files
```{r}
data_sourcesym_num <-  # REPLACE - set total number of data files (e.g. <- 2), to help in export section
```

#==========================================================================================================================#
### PROCESSING
#==========================================================================================================================#


*****************************************************************
*****************************************************************
## Here you can perform any processing to generate the data layer
## Ensure the code is clearly described with comments (#)
*****************************************************************
*****************************************************************


#==========================================================================================================================#
### EXPORT
#==========================================================================================================================#
# Write products to root directory, accompanied by yourfilename_sourcesym.txt to track source IDs
# Temp files (../proc/) generated in "PROCESS Section will lack metadata (don't use from elsewhere). Save files of value to root with correct naming & *_sourcesym.txt. 
# Check outputs - If needed, check your outputs in /proc/ before writing to file (Make note of manual check elsewhere (e.g. qgis). perhaps integrate leaflet here?
# Note on looped products with same sources, make sure to save *_sourcesym.txt for each individual product.

# SOURCESYM
# Create sourcesym file WITH all source IDs used (data and data_raw) 

# data raw sources (dir)
```{r}
print(paste("data_raw metasym files used = ", data_raw_metasym_num, sep=""))
(data_raw_sources <- tibble(id = c(source1_id, ...))) # REPLACE - keep adding all data_raw souces e.g. c( source1_id, source2_id, ...)
```

# data sources (files)
```{r}
print(paste("data_sourcym files = ", data_sourcesym_num, sep=""))
data_sources <- data1_sourcesym %>% # add any "sourcesym" files applicable to this product (check indata section), keep adding sourcesym files as needed
  add_row(data2_sourcesym)%>%  
  #add_row(data3_sourcesym)%>% 
  #add_row(data4_sourcesym)%>% 
  #add_row(data5_sourcesym)%>% 
  #add_row(data6_sourcesym)%>% 
  unique() %>% 
  print()
```
# Sources combined
```{r}
(all_sources <- data_raw_sources %>% add_row(data_sources))
```
# if a product only use a subset of the total sources, copy the section above and create individual files below (e.g. product1_sources <- ...)


# PRODUCT 1: 
# read objects
```{r}
product_orig <- REPLACE # input and check your object to be written to file
product_norm01 <- REPLACE # input and check your object to be written to file
product_uncertainty <- REPLACE # input and check your object to be written to file
plot(c(product_orig, product_norm01, product_uncertainty)) # use leaflet here instead?
```
# set names
```{r}
scale <- "REPLACE" # "1km" / "250m"
unit <- "REPLACE"  # add SI unit for product with original values (use short abbreviation e.g. meter = m, kilometer = km, percent = perc) 
(component_names <- read_tsv("./shiny_data_upload/component_names.txt")) 
```
```{r}
descriptive_name <- "REPLACE"  # name of product, for pressure/ecosystem components select from name list above
```
# paths
```{r}
(product_orig_path <- paste(dest_path,"_", descriptive_name, "_orig_", unit, "_", scale, "_", version, d_version, ".tif", sep="" ))
(product_norm01_path <- paste(dest_path,"_", descriptive_name, "_norm01_", scale, "_", version, d_version, ".tif", sep="" ))
(product_uncertainty_path <- paste(dest_path,"_", descriptive_name, "_uncertainty_", scale, "_", version, d_version, ".tif", sep="" ))
```
# Write to file
```{r}
writeRaster(product_orig, product_orig_path, COMPRESS=LZW, datatype = raster_data_type, overwrite=TRUE)
writeRaster(product_norm01, product_norm01_path, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
writeRaster(product_uncertainty, product_uncertainty_path, COMPRESS=LZW, datatype = 'INT1U', overwrite=TRUE)
write_tsv(all_sources, paste(product_orig_path, "_sourcesym.txt", sep=""))  # writes unique sourcesym file for all products, change "all_sources" to "product1_sources" if subset only
```


### SAVE SCRIPT
# 1. save your current R script (File/Save)
# 2. Save a carbon copy of the R script to proc_log (identical script to what is used to create the products exported
# Do not change the *.Rmd carbon copy in proc_log/ unless you change this version of the products (new versions are developed under process/r)

```{r}
# 2. run code below and go to File/Save As and paste link and name:
dest_path_script <- paste(getwd(), proc_path)
print(dest_path_script <- gsub(" .", "", dest_path_script)) # path
print(script_name_copy <- paste(gsub(".Rmd", "", script_name), d_version, ".Rmd", sep="")) # script name
```

#==========================================================================================================================#
### FINAL CHECK
#==========================================================================================================================#
# Developer check
# Do your own final check on the products and double check sourcesym file exists and are complete for all files in root 
# -> Sign at top of script in about section (# Approved by:)

# External check 
# To ensure repeatability and quality, make sure one colleage can run the script
# When script is proven repeatable/understandable and products/metadata look ok 
# -> Sign at top of script in about section (# Approved by:), and make any comments if needed
# There is also a work_log document in onedrive / gitHUB, check current arrangement with Swedish dev team

