# IMPORTANT: Because this script is an R Markdown script, the default directory is the location of this .Rmd file.
# To run the script from the project directory, in R Studio, you need to go to: Tools > Global Options > R Markdown
# and set "Evaluate chunks in directory" to "Project"

# =========================================================================================================
# Script by name, org
# This script creates rasters from the from the shapefiles available on the IUCN red list of threatened species
# (https://www.iucnredlist.org/resources/spatial-data-download)
# This script also resamples the rasters to the same resolution and extent as the standard WIOSym grid
# =========================================================================================================

# PREPARATIONS
```{r, include = FALSE}
# Install / load packages
#x <- c("sf", "raster", "tidyr", "dplyr")
#install.packages(x, dependencies=TRUE)
```

## Install packages
```{r, include = FALSE}
library(sf) 
library(raster)
library(fasterize)
library(tidyr)
library(dplyr)
library(rgdal)
library(tidyverse)
library(leaflet)
```

## View folders/themes
```{r}
folders_raw <- read.table("./shiny_data_upload/folders.txt", sep = "\t", header = TRUE)
folders_data <- read.table("./process/templates/folders_data.txt", sep = "\t", header = TRUE)
(folders <- folders_raw %>% add_row(folders_data))
```
```{r}
(ed_is_fich <- "small fish")
```

## Define versions/directories
```{r}
version = "v01"
d_version = ".0"
location = "glo"
theme1 <- "eco"
subtheme1 <- "mam"
name1 <- "threatened_mammals"
(dest1_path <- paste("./data", location, theme1, subtheme1, name1, version, sep="/"))
(work1_path <- paste(dest1_path, "proc", sep="/"))
(proc1_path <- paste(dest1_path, "proc_log", sep="/"))
script_path <- "./process/r"
(script_path_windows <- paste(getwd(), "/process/r", sep=""))
(script_name <- paste(theme1, "_", subtheme1,"_", location, "_", name1,"_s01_",version, ".R", sep=""))
```

## Create directories
```{r}
dir.create(dest1_path, recursive = TRUE)
dir.create(work1_path, recursive = TRUE)
dir.create(proc1_path, recursive = TRUE)
```

## Read in WIOSym standard grid
```{r}
grid_1km <- raster("./data/reg/grid/grid/v01/grid_1km_v01.1.tif")
plot(grid_1km)
```

## Define sources
```{r}
num_sources <- 1  # write the number of sources you will use in this script (to be used later for exporting metadata)
  
source1_dir <-  "./data_raw/glo/eco/mam/iucn/es2105070953/"
source1_id <- "es2105070953"
source1_metasym <- read.table(paste0(source1_dir, source1_id, "_metasym.txt"), sep = "\t")
#source1_metasym # check that metasym file is populated, and fix if not - before finalising the processing

source_list <- c(source1_id) # for multiple sources, should look like - c(source1_id, source2_id, source3_id) etc.
```

# PROCESSING



*****************************************************************
*****************************************************************
## Here you can perform any processing to generate the data layer
## Ensure the code is clearly described with comments (#)
*****************************************************************
*****************************************************************



# EXPORT

## Specify output directory and product file name

```{r}
(dest1_path_file1 <-paste0(dest1_path, "/", subtheme1, "_", name1, "_richness_1km", "_", version, d_version, ".tif"))
```

## Export metadata
```{r}
sourcesym <- data.frame(matrix(data = NA, nrow = num_sources, ncol = 1))
colnames(sourcesym) <- "id"

for(i in 1:num_sources){
  sourcesym[i,1] <- source_list[i]
}

sourcesym
```

```{r}
sourcesym_path <- paste0(dest1_path_file1, "_sourcesym.txt")
write.table(sourcesym, sourcesym_path, row.names = F, sep = "\t")
```


## Export product

```{r}
writeRaster(spec_raster, dest1_path_file1, COMPRESS=LZW, datatype = 'INT4S', overwrite=TRUE)
```

