# ========================================================================================================= #
# IMPORTANT - To get started:
Because this script is an R Markdown script, the default directory is the location of this .Rmd file.
You must open this script from within the .Rproj file associated with your wiosym database for it to work.
If you haven't used R notebooks before (.Rmd), each of the code "chunks" can be executed by clicking the green play
button at the top right of the chunk.

To run the script from the project directory, in R Studio, you need to go to: Tools > Global Options > R Markdown
and set "Evaluate chunks in directory" to "Project", though for latest version of R/Rstudio it may not be needed...
# ========================================================================================================= #
# ABOUT
# ========================================================================================================= #
## Brief description: AIS / Shipping and noise pressure to standard grid
### Sripts/processes that needs to be updated prior?: currently none, but future ersion should atleast include depth
###
### Script by: gk, sgu, Rver 4.1.1.
### Updated: Initial, org, Rver, comments
### Developer check: Initial, org, yymmdd, comments
### External check: Initial, org, yymmdd, comments
# ========================================================================================================= # 
# PREPARATIONS
# ========================================================================================================= #

### Install packages
```{r, include = FALSE}
# list all packages that this script is dependent on, remove any redundant ones
x <- c("sf", "raster", "rgdal", "tidyverse", "gdalUtils")
#install.packages("gdalUtils", dependencies=TRUE)
```

### Load packages
```{r, include = FALSE}
library(gdalUtils)
library(sf) 
library(raster)
library(rgdal)
library(tidyverse)
```

## Set version
```{r}
version = "v01" # relates to major iteration of WIOSym (v01 for grid 2021, v02 for grid 2022)
d_version = ".1"  # detailed version (you can add another level e.g. 0.0  -> v01.0.0)
seq = "s01" # sequential order - change accordingly if process is run in several scripts
```

## Set geographic scope
```{r}
read_tsv("./shiny_data_upload/locations.txt")  # prints available locations to choose from and their abbreviations
```

```{r}
location = "reg" # choose applicable 3 letter abbreviation from "location_val" -  reg for whole WIO area
```

## Set main theme
```{r}
folders_raw <- read_tsv("./shiny_data_upload/folders.txt")
folders_data <- read_tsv("./shiny_data_upload/folders_data.txt") # additional folders for data not applicable for data_raw 
folders <- folders_raw %>% add_row(folders_data)
(unique(folders["theme_folder"]))
```

```{r}
theme <- "act" # Copy from "theme_folder" (e.g. "eco")
```

## Set Subtheme
```{r}
folders %>%
  filter(theme_folder == theme) %>% 
  dplyr::select(subtheme_folder, subtheme_name) %>% 
  print(n = Inf)
```

```{r}
subtheme <- "ship" # this script will adress many pressurs bot working from shipping activity dir
```

## Set paths
```{r}
(dest_path <- paste("./data", location, theme, subtheme, version, "", sep="/")) # path to final product
(work_path <- paste(dest_path, "proc/", sep=""))
(proc_path <- paste(dest_path, "proc_log/", sep=""))
(archive_path <- paste(dest_path, "_archive/", sep=""))
(script_path <- "./process/r/")
```

## Create directories
```{r}
dir.create(dest_path, recursive = TRUE)
dir.create(work_path, recursive = TRUE)
dir.create(proc_path, recursive = TRUE)
dir.create(archive_path, recursive = TRUE)
dir.create(script_path, recursive = TRUE)
```

## Save R Script
```{r}
name1 <- "" # "REPLACE_" if script need additional name... don't use unless needed to avoid duplication.
(script_path_windows <- paste(getwd(), "/process/r", sep=""))
(script_name <- paste(theme, "_", subtheme,"_", location, "_", name1, seq, "_", version, ".Rmd", sep=""))
```
 # copy path and name and use File/Save As in R studio to save rmd file
 
#==========================================================================================================================#
# INDATA
#==========================================================================================================================#
# Set DATA_RAW sources
IMPORTANT all data harvested from data_raw needs to have a metasym.txt file in its root directory (not file specific)
## DIR1: 
```{r}
source1_dir <- source_dir <-  "./data_raw/glo/act/ship/wb/gk2110121713/"  # REPLACE - internal path
source1 <- source_id <- "gk2110121713" # 
(read_tsv(paste(source_dir, source_id, "_metasym.txt", sep=""))) # check the metasym file
``` 

```{r}
dir(source_dir, recursive=T, pattern= ".tif$")
```

```{r}
# Files
ship_stack_names <- list.files(source_dir, pattern = ".tif$", recursive=T)
ship_stack <- stack(paste(source_dir, list.files(source_dir, pattern = ".tif$", recursive=T), sep=''))
ship_stack
names(ship_stack)
# add files from dir as needed
```

### declare number of "data_raw" directorys above
```{r}
data_raw_metasym_num <- 1 # set total number of data_raw dir used (e.g. <- 1), to help in export section
```


# DATA sources
IMPORTANT Products in data are based on information in data_raw. Check that *_sourcesym.txt exists for each file to track acccumulated IDs. 

## DATA1: Grid 1km - loading all flavours (watermask 1, 0, NA) at once since identical sourcesym apply
```{r}
data_dir <-  "./data/reg/grid/grid/v01/"  # input your data directory
dir(data_dir) # check what data is in this directory

data1_file <- "grid_1km_v01.1.tif" 
data1_file2 <- "grid_1km_0_v01.1.tif" 
data1_file3 <- "grid_1km_na_v01.1.tif" 

(grid_1km_path <- path <- paste(data_dir, data1_file, sep=""))
(grid_1km <- raster(path)) # read and check your data
(grid_1km_0_path <- path <- paste(data_dir, data1_file2, sep=""))
(grid_1km_0 <- path <- raster(path)) # read and check your data
(grid_1km_na_path <- path <- paste(data_dir, data1_file3, sep=""))
(grid_1km_na <- path <- raster(path)) # read and check your data

data1_sourcesym <- read_tsv(paste(data_dir, data1_file, "_sourcesym.txt", sep=""))
# check that source IDs exists and make sense

plot(grid_1km)
```

## DATA2: Grid 250m - loading all flavours (watermask 1, 0, NA) at once since same sourcesym apply
```{r}
data2_dir <- data_dir <-  "./data/reg/grid/grid/v01/"  # input your data directory

data2_file <- "grid_250m_v01.1.tif" # your data file
data2_file2 <- "grid_250m_0_v01.1.tif" # your data file
data2_file3 <- "grid_250m_na_v01.1.tif" # your data file

(grid_250m_path <- path <- paste(data_dir, data2_file, sep=""))
(grid_250m <- raster(path)) # read and check your data
(grid_250m_0_path <- path <- paste(data_dir, data2_file2, sep=""))
(grid_250m_0 <- path <- raster(path)) # read and check your data
(grid_250m_na_path <- path <- paste(data_dir, data2_file3, sep=""))
(grid_250m_na <- path <- raster(path)) # read and check your data

(data2_sourcesym <- read_tsv(paste(data_dir, data2_file, "_sourcesym.txt", sep="")))
# check so source IDs exists and make sense
# check that source IDs exists and make sense
#plot(grid_250m)
```

# data3: bounding box for the grid (created by SGU in GIS so no data_raw sources involved)
data3_dir <- "./data/reg/grid/scope/v01/" 
dir(data2_dir)
data3_file <- "wiosym_data_bounding_box_v01.shp" 
data3_sourcesym <- read_tsv(paste(data2_dir, data2_file, "_sourcesym.txt", sep=""))
data3_sourcesym # check so source IDs exists and make sense
data3_path <- paste(data3_dir, data3_file, sep="")
data3_path # note: this data is not read in as an object since the process in this script happens in GDAL outside R

extent_box <- st_read(data3_path)




## declare number of "data" files
```{r}
data_sourcesym_num <-  2 # set total number of data files (e.g. <- 2), to help in export section
```

#==========================================================================================================================#
# PROCESSING
#==========================================================================================================================#

## Map to 1km grid 
it may be better to map to first map to 250m grid next iteration, check the source resolution, 0.005 deg should be about 500m.. and do some thinking..
```{r}
#ship_sum <- calc(ship_stack, sum) # total sum of all cells (should be 100% for substrate components, but isnt yet)

# divide by total sum and print to file
stack <- ship_stack
stack <- crop(ship_stack, extent_box)  # need to crop to even larger extent to avoid edge effects! or fill edges with NA...

for (i in c(1:6)) {
(indata_path <- paste(source1_dir, ship_stack_names[i], sep=""))

(outdata_path <- paste(work_path, names(stack)[i], "_grid1km_.tif", sep=""))
writeRaster(grid_1km_na, outdata_path, overwrite=TRUE, COMPRESS=LZW)

sr <- "+proj=cea +lat_ts=-12 +lon_0=12 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"    

raster_warp <- gdalUtils::gdalwarp(srcfile = indata_path,
                                dstfile = outdata_path,
                                s_srs = "+proj=longlat +datum=WGS84 +no_defs",
                                t_srs = sr,
                                tr = c(1000, 1000),
                                tap = TRUE,
                                output_Raster = TRUE,
                                overwrite = TRUE,
                                r = "average",
                                multi = TRUE,
                                srcnodata = "NA"
)


r <- raster(outdata_path) 

r <- crop(r, extent(grid_1km))
r <- mask(r, grid_1km)

(r_path <- paste(work_path, names(stack)[i], "_grid1km_lzw.tif", sep=""))
writeRaster(r, r_path, overwrite=TRUE, datatype = 'FLT4S', COMPRESS=LZW)
}

```


## turn AIS into noise disturbane
https://www.frontiersin.org/articles/10.3389/fmars.2019.00606/full#supplementary-material
more work needed...
```{r}
#ship_sum <- calc(ship_stack, sum) # total sum of all cells (should be 100% for substrate components, but isnt yet)

stack <- stack(paste(work_path, list.files(work_path, pattern = "_grid1km_lzw.tif$", recursive=T), sep=''))
  
plot(stack[[1]])
names(stack)

# adding 0 to nodata areas...
stack_merge <- merge(stack, grid_1km_0)
stack <- stack_merge


# factor for dB of different shiptypes, approximate from Dr. Hammar...

f_high <- 100000 # Watt at 170dB mainly commersal traffic (assuming oil vessels are the same..)
f_med <- 10000 # W at 160 dB
f_low  <- 1000 # W at 150 dB

noise <-  (stack[[1]] + stack[[5]])*f_high + (stack[[2]] + stack[[6]])*f_med + (stack[[3]] + stack[[4]])*f_low  

r <- noise

#r <- reclassify(r, c(-5, 0.5, 0))
#plot(r)

w <- focalWeight(grid_1km, 10000, "Gauss")
r_focal <- focal(r, w, na.rm=T) 

w <- focalWeight(grid_1km, 1000, "Gauss")
r_focal_1km <- focal(r, w, na.rm=T) 

noise_com <- (stack[[1]])*f_high
w <- focalWeight(grid_1km, 100000, "Gauss")
r_focal_100km <- focal(noise_com, w, na.rm=T) 




r_comb <- r_focal + r_focal_1km + r_focal_100km
#r <- r_comb
#r <- reclassify(r, c(50, 100000000000000, 50))

r <- r_comb

r <- merge(r, grid_1km_0)
r <- crop(r, grid_1km)
r <- mask(r, grid_1km)
r

#r <- reclassify(r, c(10, 50, 10))

noise_focal <- r   # noise impact ready for export!
plot(r)
```



## turn AIS into shipstrikes
more work needed...
```{r}
#ship_sum <- calc(ship_stack, sum) # total sum of all cells (should be 100% for substrate components, but isnt yet)

#stack <- stack(paste(work_path, list.files(work_path, pattern = "_grid1km_lzw.tif$", recursive=T), sep=''))
#plot(stack[[1]])
names(stack)

f_high <- 10 # assumption that needs more careful consideration. assuming commersal traffic have bigger impact due to size and speed
f_med <- 5 # W at 160 dB
f_low  <- 3 # W at 150 dB
ship_strikes <-  ((stack[[1]] + stack[[5]])*f_high + (stack[[2]] + stack[[6]])*f_med + (stack[[3]] + stack[[4]])*f_low)

r <- ship_strikes

#r <- reclassify(r, c(-5, 0.5, 0))
#plot(r)
w <- focalWeight(r, 1000, "Gauss") # "blurring" data to account for uncertainty
r_focal <- focal(r, w, na.rm=T) 

r_focal
plot(r_focal)

r <- r_focal

#r <- merge(r, grid_1km_0)
r <- crop(r, grid_1km)
r <- mask(r, grid_1km)
r

#r <- reclassify(r, c(10, 50, 10))

shipstrike_focal <- r   # noise impact ready for export!
plor(r)
```

## turn AIS into "pollution"
more work needed...
```{r}
names(stack)

r <- stack[[3]] # going only with "global layer here..., more thinking needed
w <- focalWeight(r, 50000, "Gauss") # simulating spread of pollutants
r_focal_50km <- focal(r, w, na.rm=T) 

w <- focalWeight(r, 2000, "Gauss") # simulating spread of pollutants
r_focal_2km <- focal(r, w, na.rm=T) 

r <- r_focal_50km + r_focal_2km

plot(r)

r <- merge(r, grid_1km_0)
r <- crop(r, grid_1km)
r <- mask(r, grid_1km)

r

#r <- reclassify(r, c(10, 50, 10))

pollution_focal <- r   # pollution impact ready for export!
plot(r)
```


## turn AIS into "recreational_boating"
more work needed...
```{r}
names(stack)

r <- stack[[4]] 
w <- focalWeight(r, 5000, "Gauss") # simulating spread of disturbance
r_focal_5km <- focal(r, w, na.rm=T) 

w <- focalWeight(r, 1000, "Gauss") # simulating spread of disturbance
r_focal_1km <- focal(r, w, na.rm=T) 


r <- r_focal_5km + r_focal_1km

plot(r)

r <- merge(r, grid_1km_0)
r <- crop(r, grid_1km)
r <- mask(r, grid_1km)

r

#r <- reclassify(r, c(10, 50, 10))

recreational_boating <- r   # ready for export
plot(r)
```


#==========================================================================================================================#
# EXPORT
#==========================================================================================================================#
# Write products to root directory when checked and ready to use elsewhere, accompanied by yourfilename.ext_sourcesym.txt 
# for each file to track source IDs as they accumulate between processes / data


# SOURCESYM
### Create sourcesym file with all source IDs used (data and data_raw) 
# data raw sources (dir)
```{r}
print(paste("data_raw metasym files used = ", data_raw_metasym_num, sep=""))
(data_raw_sources <- tibble(id = c(source1))) # REPLACE / add all data_raw sources
```

## data sources (files)
```{r}
print(paste("data_sourcym files = ", data_sourcesym_num, sep=""))
data_sources <- data1_sourcesym %>% # add any "sourcesym" files applicable to this product (check indata section), keep adding sourcesym files as needed
  add_row(data2_sourcesym)%>%  
  #add_row(data3_sourcesym)%>% 
  #add_row(data4_sourcesym)%>% 
  #add_row(data5_sourcesym)%>% 
  #add_row(data6_sourcesym)%>% 
  unique() %>% 
  print()
```
## Sources combined
```{r}
(all_sources <- data_raw_sources %>% add_row(data_sources))
```
### if a product only use a subset of the total sources, copy the section above and create individual files below (e.g. product1_sources <- ...)


# PRODUCT 1: 
## read objects
```{r}
product_orig <- noise_focal # input and check your object to be written to file
product_orig2 <- shipstrike_focal # input and check your object to be written to file
product_orig3 <- pollution_focal # input and check your object to be written to file
product_orig4 <- recreational_boating # input and check your object to be written to file

plot(product_orig)
```
## set names
```{r}
scale <- "1km" # "1km" / "250m"
unit <- "relative"  # add unit for product with original values (e.g. m, km, perc, presence, proportion
component_names <- read_tsv("./shiny_data_upload/component_names.txt") 
component_names_selected <- component_names %>%
  dplyr::filter(theme_folder == theme) %>% 
  dplyr::select(group, file_name, description) %>% 
  print(n = Inf)
```

```{r}
descriptive_name <- "underwater_noise"  # name of product, for pressure/ecosystem components select from name list above
descriptive_name2 <- "ship_strike"  # name of product, for pressure/ecosystem components select from name list above
descriptive_name3 <- "ship_pollution"  # name of product, for pressure/ecosystem components select from name list above
descriptive_name4 <- "recreational_boating"  # name of product, for pressure/ecosystem components select from name list above

```
## paths
```{r}
(product_orig_path <- paste(dest_path, descriptive_name, "_", "relative_intensity_w", "_", scale, "_", version, d_version, ".tif", sep="" ))
(product_orig2_path <- paste(dest_path, descriptive_name2, "_", unit, "_", scale, "_", version, d_version, ".tif", sep="" ))
(product_orig3_path <- paste(dest_path, descriptive_name3, "_", unit, "_", scale, "_", version, d_version, ".tif", sep="" ))
(product_orig4_path <- paste(dest_path, descriptive_name4, "_", unit, "_", scale, "_", version, d_version, ".tif", sep="" ))




```
## Write to file
```{r}
# check datatype to be appropriate for values (flt or int, 'INT4S' option to..)
writeRaster(product_orig, product_orig_path, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
writeRaster(product_orig2, product_orig2_path, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
writeRaster(product_orig3, product_orig3_path, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)
writeRaster(product_orig4, product_orig4_path, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)


# write sourcesym file, change "all_sources" to "selected_sources" if subset only for specific file
write_tsv(all_sources, paste(product_orig_path, "_sourcesym.txt", sep=""))  
write_tsv(all_sources, paste(product_orig2_path, "_sourcesym.txt", sep="")) 
write_tsv(all_sources, paste(product_orig3_path, "_sourcesym.txt", sep="")) 
write_tsv(all_sources, paste(product_orig4_path, "_sourcesym.txt", sep="")) 
```


# SAVE SCRIPT
## 1. save your current R script (File/Save)
## 2. Save a carbon copy of the R script to proc_log (identical script to what is used to create the products exported
## Do not change the *.Rmd carbon copy in proc_log/ unless you change this version of the products (new versions are developed under process/r)

```{r}
# 2. run code below and go to File/Save As and paste link and name:
dest_path_script <- paste(getwd(), proc_path)
print(dest_path_script <- gsub(" .", "", dest_path_script)) # path
print(script_name_copy <- paste(gsub(".Rmd", "", script_name), d_version, ".Rmd", sep="")) # script name
```

#==========================================================================================================================#
# FINAL CHECK
#==========================================================================================================================#
# Developer check
## Do your own final check on the products and double check sourcesym file exists and are complete for all files in root 
## -> Sign at top of script in about section (# Approved by:)

# External check 
## To ensure repeatability and quality, make sure one colleage can run the script
## When script is proven repeatable/understandable and products/metadata look ok 
## -> Sign at top of script in about section (# Approved by:), and make any comments if needed
## There is also a work_log document in onedrive / gitHUB, check current arrangement with Swedish dev team
