# ========================================================================================================= #
# IMPORTANT - To get started:
Because this script is an R Markdown script, the default directory is the location of this .Rmd file.
You must open this script from within the .Rproj file associated with your wiosym database for it to work.
If you haven't used R notebooks before (.Rmd), each of the code "chunks" can be executed by clicking the green play
button at the top right of the chunk.

To run the script from the project directory, in R Studio, you need to go to: Tools > Global Options > R Markdown
and set "Evaluate chunks in directory" to "Project", though for latest version of R/Rstudio it may not be needed...
# ========================================================================================================= #
# ABOUT
# ========================================================================================================= #
## Brief description: cript to make sure rasters are mapped to the grid with 0 values...
### Sripts/processes that needs to be updated prior?:
###
### Script by: gk, sgu, r4.1.1
### Updated: Initial, org, Rver, comments
### Developer check: Initial, org, yymmdd, comments
### External check: Initial, org, yymmdd, comments

### Install packages
```{r, include = FALSE}
# list all packages that this script is dependent on, remove any redundant ones
x <- c("leaflet", "sf", "raster", "rgdal", "tidyverse")
#install.packages(x, dependencies=TRUE)
```

### Load packages
```{r, include = FALSE}
library(sf) 
library(raster)
library(rgdal)
library(tidyverse)
```

# DATA sources
IMPORTANT Products in data are based on information in data_raw. Check that *_sourcesym.txt exists for each file to track acccumulated IDs. 

## DATA1: Grid 1km - loading all flavours (watermask 1, 0, NA) at once since identical sourcesym apply
```{r}
data_dir <-  "./data/reg/grid/grid/v01/"  # input your data directory
dir(data_dir) # check what data is in this directory

data1_file <- "grid_1km_v01.1.tif" 
data1_file2 <- "grid_1km_0_v01.1.tif" 
data1_file3 <- "grid_1km_na_v01.1.tif" 

(grid_1km_path <- path <- paste(data_dir, data1_file, sep=""))
(grid_1km <- raster(path)) # read and check your data
(grid_1km_0_path <- path <- paste(data_dir, data1_file2, sep=""))
(grid_1km_0 <- path <- raster(path)) # read and check your data
(grid_1km_na_path <- path <- paste(data_dir, data1_file3, sep=""))
(grid_1km_na <- path <- raster(path)) # read and check your data

data1_sourcesym <- read_tsv(paste(data_dir, data1_file, "_sourcesym.txt", sep=""))
# check that source IDs exists and make sense

plot(grid_1km)
```

## DATA2: Grid 250m - loading all flavours (watermask 1, 0, NA) at once since same sourcesym apply
```{r}
data2_dir <- data_dir <-  "./data/reg/grid/grid/v01/"  # input your data directory

data2_file <- "grid_250m_v01.1.tif" # your data file
data2_file2 <- "grid_250m_0_v01.1.tif" # your data file
data2_file3 <- "grid_250m_na_v01.1.tif" # your data file

(grid_250m_path <- path <- paste(data_dir, data2_file, sep=""))
(grid_250m <- raster(path)) # read and check your data
(grid_250m_0_path <- path <- paste(data_dir, data2_file2, sep=""))
(grid_250m_0 <- path <- raster(path)) # read and check your data
(grid_250m_na_path <- path <- paste(data_dir, data2_file3, sep=""))
(grid_250m_na <- path <- raster(path)) # read and check your data

(data2_sourcesym <- read_tsv(paste(data_dir, data2_file, "_sourcesym.txt", sep="")))
# check so source IDs exists and make sense
# check that source IDs exists and make sense
#plot(grid_250m)
```


## DATA 3: bounding box shapefile
```{r}
data3_dir <- data_dir <-  "./data/reg/grid/scope/v01/"  # input your data directory

data3_file <- "wiosym_grid_bounding_box_v01.shp" # your data file

data3_sourcesym <- read_tsv(paste(data_dir, data3_file, "_sourcesym.txt", sep=""))
#data3_sourcesym # check so source IDs exists and make sense - ok for these
data3_path <- data_path <- paste(data_dir, data3_file, sep="")

grid_poly <- st_read(data_path) # read in your data object if appropriate and perhaps change name to more informative...
#grid_poly
#plot(grid_poly)
```

# INRASTER TO FIX
```{r}
data4_dir <- data_dir <-  "./data/reg/act/lb_other/v01/proc/"  # input your data directory
dir(data_dir)
data4_file <- data_file <- "Airport_20kmbuffer_raster1.tif" # your data file
data4_path <- paste(data4_dir, data_file, sep="")
data_dir <-  "./data/reg/pres/lb_other/v01/"
data_outpath <- paste(data4_dir, "air_traffic_landingimpact_v01.0_qc.tif", sep="")
r_orig <- raster(data4_path) #
r_orig
plot(r_orig)
```





## declare number of "data" files
```{r}
data_sourcesym_num <-  4 # set total number of data files (e.g. <- 2), to help in export section
```

#==========================================================================================================================#
# PROCESSING / Export - manuel nödlösning under tidspress... måste fixas om det ska bli något att använda för alla sen...
#==========================================================================================================================#
```{r}
r <- r_orig

#r <- reclassify(r, c(-5, 0.5, 0))
plot(r)
w <- focalWeight(r, 5000, "Gauss")
r_focal <- focal(r, w,) 

r <- r_focal
#r_comb <- r + r_focal
#r <- r_comb
#r <- reclassify(r, c(50, 100000000000000, 50))



r <- merge(r, grid_1km_0)
r <- crop(r, grid_1km)
r <- mask(r, grid_1km)
r

r <- reclassify(r, c(10, 50, 10))

writeRaster(r, data_outpath, COMPRESS=LZW, datatype = 'FLT4S', overwrite=TRUE)


```
