# =========================================================================================================
# Script by Edmond Sacre, SLU
# =========================================================================================================

# PREPARATIONS
```{r, include = FALSE}
# Install / load packages
#x <- c("sf", "raster", "tidyr", "dplyr")
#install.packages(x, dependencies=TRUE)
```

## Install packages
```{r, include = FALSE}
library(tidyr)
library(dplyr)
library(terra)
```

## View folders/themes
```{r}
#folders_raw <- read.table("./shiny_data_upload/folders.txt", sep = "\t", header = TRUE)
#folders_data <- read.table("./process/templates/folders_data.txt", sep = "\t", header = TRUE)
#folders <- folders_raw %>% add_row(folders_data)
```

## Define versions/directories
```{r}
version = "v01"
d_version = ".0"
location = "reg"   #check
theme1 <- "pres"   #check
subtheme1 <- "cc_sealev"   #check
name1 <- "sea_level_rise"   #check
name2 <- "" #check
(dest1_path <- paste("./data", location, theme1, subtheme1, name1, version, sep="/"))
(work1_path <- paste(dest1_path, "proc", sep="/"))
(proc1_path <- paste(dest1_path, "proc_log", sep="/"))
script_path <- "./process/r"
(script_path_windows <- paste(getwd(), "/process/r", sep=""))
(script_name <- paste(theme1, "_", subtheme1,"_", location, "_", name1,"_s01_",version, ".Rmd", sep=""))
(dest1_path_file1 <-paste0(dest1_path, "/", subtheme1, "_", name1, "_", version, d_version, ".tif"))
```

## Create directories
```{r}
dir.create(dest1_path, recursive = TRUE)
dir.create(work1_path, recursive = TRUE)
dir.create(proc1_path, recursive = TRUE)
```

## Read in WIOSym standard grid
```{r}
grid_1km <- rast("./data/reg/grid/grid/v01/grid_1km_v01.1.tif")
grid_proj <- rast("./products/grid/grid_1km_v01.1_epsg4326.tif")
```

## Define sources
```{r}
num_sources <- 1  # write the number of sources you will use in this script (to be used later for exporting metadata)
  
source1_dir <-  "./data_raw/glo/pres/cc_sealev/cmems/es2405271350189472/"    #check
source1_id <- "es2405271350189472"    #check
source1_metasym <- read.table(paste0(source1_dir, source1_id, "_metasym.txt"), sep = "\t")
#source1_metasym # check that metasym file is populated, and fix if not - before finalising the processing

source_list <- c(source1_id) # for multiple sources, should look like - c(source1_id, source2_id, source3_id) etc.
```

*****************************************************************
*****************************************************************
# PROCESSING

# Install processing specific packages
```{r, include = FALSE}
library(terra)
library(ncdf4)
```

# Import sea level rise data from Copernicus
```{r}
fp <- paste0(source1_dir, "omi_climate_sl_global_regional_trends_19930101_P20230403.nc")
sea <- rast(fp)
sea <- crop(sea, ext(8, 78, -40, 16))
plot(sea)
```

```{r}
fp <- paste0(dest1_path, "/resint.R")
source(fp)

sea <- resint(sea, grid_1km, nit = 100, win_size = 10000, method = "cubicspline")
plot(sea)
terra::minmax(sea) # Show max value of mean rise of 14.34 mm per year
```

# Transform data based on quantile cutoff to accentuate extreme areas
```{r}
sea_quant <- sea
q <- quantile(terra::values(sea, na.rm = T), probs = 0.75) # Add quantile cutoff? Default to no cutoff
sea_quant[sea_quant < q] <- 0
plot(sea_quant)
```

```{r}
product1 <- sea_quant
```


*****************************************************************
*****************************************************************
# EXPORT

## Specify output directory and product file name
```{r}
(dest1_path_file1 <-paste0(dest1_path, "/", subtheme1, "_", name1, "_", version, d_version, ".tif"))
```

## Export image of product raster(s)
```{r}
(img_path <- paste0(dest1_path, "/", subtheme1, "_", name1, "_", version, d_version, ".png"))
grid_1km_raster <- raster::raster(grid_1km)
prod1_scale <- raster::raster(product1)
prod1_scale[prod1_scale == 0] <- NA
prod1_scale[] <- scales::rescale(prod1_scale[], to = c(1, 100))
prod1_scale[is.na(prod1_scale)] <- 0
prod1_scale[is.na(grid_1km_raster)] <- NA


png(img_path, width = ncol(prod1_scale), height = nrow(prod1_scale))
pal <- colorRampPalette(rev(c('#d53e4f','#f46d43','#fdae61','#fee08b','#ffffbf','#e6f598','#abdda4','#66c2a5','#3288bd')))
plot(prod1_scale, maxpixels = ncell(prod1_scale), main = "", axes = FALSE, ylim=raster::extent(prod1_scale)[3:4], 
     breaks = seq(0,100,1), col = pal(101), legend = F, box = F, colNA = "gray92")
dev.off()
```

## Export metadata
```{r}
sourcesym <- data.frame(matrix(data = NA, nrow = num_sources, ncol = 1))
colnames(sourcesym) <- "id"

for(i in 1:num_sources){
  sourcesym[i,1] <- source_list[i]
}

sourcesym
```

```{r}
sourcesym_path <- paste0(dest1_path_file1, "_sourcesym.txt")
write.table(sourcesym, sourcesym_path, row.names = F, sep = "\t")
```

## Export product
```{r}
writeRaster(product1, dest1_path_file1, gdal = c("COMPRESS=LZW"), overwrite=TRUE)
```

